### 2013-09-02 report/report.tex
Modified gitignore and added report texfile

\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}

\usepackage{amsmath,amssymb,amsfonts}

\title{Local Adaptive Optimization of Timestep}
\author{Malte Stær Nissen\\University of Copenhagen\\\texttt{malte.nissen@gmail.com}}

\begin{document}
\maketitle

Supervisor: Sune Darkner\\\texttt{darkner@diku.dk}

\section{Problem}

\subsection{Why?}
In order to minimize the risk of grids exploding/imploding caused by too large timesteps and avoid the need to compute the smaller timesteps on the entire grid.

\subsection{How?}

\subsection{Learning goals}
Evaluate local adaptive optimization of timesteps to minimize the risk of unwanted deformations in grids.

\subsection{Risk assessment}

\end{document}

### 2013-09-03 report/report.tex
Added more compiler-files to gitignore (latex) and updated project contract

@@
 \documentclass[11pt,a4paper]{article}
-
+\usepackage[english]{babel}
 \usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
 \usepackage[T1]{fontenc}
+\usepackage{lmodern}
 
@@
 
-\title{Local Adaptive Optimization of Timestep}
-\author{Malte Stær Nissen\\University of Copenhagen\\\texttt{malte.nissen@gmail.com}}
+\title{Project Contract\\Local Adaptive Optimization of Timestep}
+\author{Malte Stær Nissen\\University of Copenhagen\\\texttt{malte.nissen@gmail.com}\\Supervisor: Sune Darkner\\\texttt{darkner@diku.dk}}
 
@@
 
-Supervisor: Sune Darkner\\\texttt{darkner@diku.dk}
+\section{Learning Goals}
+Course specific learning goals:
+\begin{enumerate}
+	\item Formulate an operational project plan.
+	\item Search the relevant literature and write a literature review setting own work in perspective.
+	\item Be able to formulate what plagiarism is, and demonstrate proper citation and reference style.
+	\item Solve a selected problem of fair Computational and Mathematical Modelling / eScience content and difficulty.
+	\item Produce a thorough experiment plan that clearly demonstrates the quality of and highlights boundaries for the solution. 
+	\item Produce a scientific text of fair quality both textually and scientifically.
+	\item The student must be able to make an oral presentation of own work.
+\end{enumerate}
+Project specific learning goals:
+\begin{enumerate}
+	\item Evaluate local adaptive optimization of timesteps to minimize the risk of unwanted deformations in grids.
+\end{enumerate}
 
-\section{Problem}
+\section{Problem Description}
 
-\subsection{Why?}
-In order to minimize the risk of grids exploding/imploding caused by too large timesteps and avoid the need to compute the smaller timesteps on the entire grid.
+\subsection{Problem Statement}
 
-\subsection{How?}
+\subsection{Motivation}
 
-\subsection{Learning goals}
-Evaluate local adaptive optimization of timesteps to minimize the risk of unwanted deformations in grids.
+\subsection{Proposed solution}
+
+\subsection{Success criteria}
 

### 2013-09-04 report/report.tex
First draft of project contract

@@
 \subsection{Motivation}
+When running a simulation of large hyperelastic materials we wish to make
+large enough steps for the simulation to be performed within a reasonable
+timeframe, but we still want the simulation to be accurate and to avoid
+the mesh from curling up or exploding due to too large determinants of the
+moving mesh vertices. In some hyperelastic materials most of the
+discretized points are able to be simulated using larger timesteps whereas
+smaller areas might need smaller timesteps. These situations are the ones
+we wish to accomodate in this paper.
 
 \subsection{Proposed solution}
+We wish to develop a method for locally adapting the timestep size in
+order to avoid curl ups/explosions while still maintaining the original
+timestep size of the rest of the simulation. In other words for each of
+the original timesteps we perform further simulations of problematic local
+areas of the grid with smaller timesteps.
 
 \subsection{Success criteria}
+Avoid curl ups/explosions of the grid while maintaining the original
+stepsize by using local subsimulations (with smaller stepsizes)
 
 \subsection{Risk assessment}
+Low. The technique has been performed before in various shapes and it
+should therefore be do-able.
 

### 2013-09-05 report/report.tex
Project learning goals added

@@
 \begin{enumerate}
-	\item Evaluate local adaptive optimization of timesteps to minimize the risk of unwanted deformations in grids.
+    \item Implement local adaptive optimization of timesteps (LAOT)
+	\item Evaluate LAOT to minimize the risk of unwanted deformations in grids.
+    \item Discuss how LAOT is performed in practice using existing 2nd order optimization methods.
+    \item Study of convergence on implemented methods.
 \end{enumerate}

### 2013-09-05 report/report.tex
Project contract done

@@
 \subsection{Problem Statement}
+Study the application of locally adaptive timestepping in simulation of hyperelastic materials.
 
@@
 \subsection{Success criteria}
-Avoid curl ups/explosions of the grid while maintaining the original
-stepsize by using local subsimulations (with smaller stepsizes)
+To gain insight into the application of LAOT.
 
 \subsection{Risk assessment}
-Low. The technique has been performed before in various shapes and it
+Low to none. The technique has been performed before in various shapes and it
 should therefore be do-able.

### 2013-09-13 report/report.tex
Modified list of papers, report and gitignore

--- 2013-09-04 report/report.tex
@@
 \usepackage{lmodern}
+\usepackage{todonotes}
 
@@
 
-\title{Project Contract\\Local Adaptive Optimization of Timestep}
-\author{Malte Stær Nissen\\University of Copenhagen\\\texttt{malte.nissen@gmail.com}\\Supervisor: Sune Darkner\\\texttt{darkner@diku.dk}}
+\title{Local Adaptive Optimization of Timestep}
+\author{Malte Stær Nissen\\University of Copenhagen\\\texttt{malte.nissen@gmail.com} \and Supervisor: Sune Darkner\\University of Copenhagen\\\texttt{darkner@diku.dk}}
 
@@
 
-\section{Learning Goals}
-Course specific learning goals:
-\begin{enumerate}
-	\item Formulate an operational project plan.
-	\item Search the relevant literature and write a literature review setting own work in perspective.
-	\item Be able to formulate what plagiarism is, and demonstrate proper citation and reference style.
-	\item Solve a selected problem of fair Computational and Mathematical Modelling / eScience content and difficulty.
-	\item Produce a thorough experiment plan that clearly demonstrates the quality of and highlights boundaries for the solution. 
-	\item Produce a scientific text of fair quality both textually and scientifically.
-	\item The student must be able to make an oral presentation of own work.
-\end{enumerate}
-Project specific learning goals:
-\begin{enumerate}
-	\item Evaluate local adaptive optimization of timesteps to minimize the risk of unwanted deformations in grids.
-\end{enumerate}
+\tableofcontents
 
-\section{Problem Description}
+\clearpage
 
-\subsection{Problem Statement}
+\section{Introduction}
+Simulations of large (order of millions of vertices) hyperelastic materials are
+very time consuming. When working with materials of varying density of vertices,
+the highest densities often make the most strict bounds on size of the timestep
+of the simulation in order to keep the materials (and grids) stable. In a
+``standard'' simulation with a global timestep size we need to compute possibly
+unneccessarily many computations on the coarser parts of the material caused by this bound. When
+furthermore having materials with a large amount of somewhat equally distributed
+vertices and smaller patches of detailed areas with higher density of vertices,
+it would be an advantage to be able to perform large timesteps for the majority of
+the material and smaller steps for the local patches when vertices cause
+instability. We call this concept local adaptive optimization of timestep, which
+we will study further in this report.
 
-\subsection{Motivation}
-When running a simulation of large hyperelastic materials we wish to make
-large enough steps for the simulation to be performed within a reasonable
-timeframe, but we still want the simulation to be accurate and to avoid
-the mesh from curling up or exploding due to too large determinants of the
-moving mesh vertices. In some hyperelastic materials most of the
-discretized points are able to be simulated using larger timesteps whereas
-smaller areas might need smaller timesteps. These situations are the ones
-we wish to accomodate in this paper.
+\section{Previous work}
+As according to \cite{Gander:2013} local time stepping was first studiet
+in the community of ordinary differential equations (ODE) with Rice
+\cite{rice:1960} developing the split Runge-Kutta methods (multirate
+Runge-Kutta methods). The concept behind these methods is the splitting of
+the ODEs into multiple (two components in \cite{rice:1960}) components of
+which each component needs to be integrated in different scales. A strategy
+is chosen as proposed in \cite{Kvaernoe:1999} of either computing the coarser
+part first (``Fastest first strategy'') followed by the finer part or vice
+versa (``Slowest first strategy''). In order to perform these sequential
+computations either interpolation or extrapolation of the first computation
+is performed to compute the second part in each time step and the different
+time step sizes can be adapted to fit the model in each step as well. See
+\cite{Kvaernoe:1999} and \cite{Gear:1984} (similar strategy for linear
+multistep methods) for more detailed descriptions of the two strategies.
 
-\subsection{Proposed solution}
-We wish to develop a method for locally adapting the timestep size in
-order to avoid curl ups/explosions while still maintaining the original
-timestep size of the rest of the simulation. In other words for each of
-the original timesteps we perform further simulations of problematic local
-areas of the grid with smaller timesteps.
+In the partial differential equations (PDE) community, the adaptive
+timestepping area was explored and developed with the goal of simulating
+very specific known problems such as the (hyperbolic) wave equations
+and (parabolic) heat equations instead of more general applications. We
+will adhere from describing the methods further in this report, but a
+short description and comparison of the specific methods can be found in
+\cite{Gander:2013}. Generally the methods developed at first all make use
+of the same basic concepts for making the local adaptive timestepping:
+Interpolation, extrapolation, prediction and correction, which we are going to
+use in the work of this report as well.
 
-\subsection{Success criteria}
-Avoid curl ups/explosions of the grid while maintaining the original
-stepsize by using local subsimulations (with smaller stepsizes)
+Recently newer and faster methods with local time stepping have been
+developed on the basis of nonlinear PDEs and the work performed in this
+field in the 80s and 90s. First the multiresolution (MR) schemes for creating
+space-adaptive discretizations and refinements of these were developed,
+see \cite{Berger:1984}. Then multiple MR based schemes were developed.
+Dominques et al. \cite{Domingues:2008} is one of the more recent of these MR
+schemes. They describe a local scale-dependent time stepping
+for a space-adaptive multiresolution scheme using the finite volume method
+in order to obtain speed-up using larger time-steps without violating a
+defined stability constraints, which is essentially the same motivation as
+this report. The method is based on an explicit Runge-Kutta method of second
+order. As expected the timestep size is imposed by a stability condition of
+the explicit Runge-Kutta on the finest scale, which increases with the scale
+of the mesh and hence we are able to increase the timestep as well without
+violating the stability condition.
 
-\subsection{Risk assessment}
-Low. The technique has been performed before in various shapes and it
-should therefore be do-able.
+\section{Method}
+
+\section{Experiments and results}
+
+\section{Discussion}
+
+\section{Conclusion}
+
+\cite{Keshav:2007}
+\bibliography{bibliography}
+\bibliographystyle{plain}
 

### 2013-09-16 report/report.tex
Previous work section finished

@@
 
-\title{Local Adaptive Optimization of Timestep}
+\title{Local Adaptive Optimization of Time step}
 \author{Malte Stær Nissen\\University of Copenhagen\\\texttt{malte.nissen@gmail.com} \and Supervisor: Sune Darkner\\University of Copenhagen\\\texttt{darkner@diku.dk}}
@@
 \section{Introduction}
-Simulations of large (order of millions of vertices) hyperelastic materials are
-very time consuming. When working with materials of varying density of vertices,
-the highest densities often make the most strict bounds on size of the timestep
-of the simulation in order to keep the materials (and grids) stable. In a
-``standard'' simulation with a global timestep size we need to compute possibly
-unneccessarily many computations on the coarser parts of the material caused by this bound. When
-furthermore having materials with a large amount of somewhat equally distributed
-vertices and smaller patches of detailed areas with higher density of vertices,
-it would be an advantage to be able to perform large timesteps for the majority of
-the material and smaller steps for the local patches when vertices cause
-instability. We call this concept local adaptive optimization of timestep, which
-we will study further in this report.
+Simulations of large (order of millions of vertices) hyper elastic materials
+are very time consuming. When working with materials of varying density of
+vertices, the highest densities often make the most strict bounds on size of
+the time step of the simulation in order to keep the materials (and grids)
+stable. In a ``standard'' simulation with a global time step size we need
+to compute possibly unnecessarily many computations on the coarser parts of
+the material caused by this bound. When furthermore having materials with a
+large amount of somewhat equally distributed vertices and smaller patches of
+detailed areas with higher density of vertices, it would be an advantage to be
+able to perform large time steps for the majority of the material and smaller
+steps for the local patches when vertices cause instability. We call this
+concept local adaptive optimization of time step, which we will study further
+in this report.
 
 \section{Previous work}
-As according to \cite{Gander:2013} local time stepping was first studiet
+As according to \cite{Gander:2013} local time stepping was first studied
 in the community of ordinary differential equations (ODE) with Rice
-\cite{rice:1960} developing the split Runge-Kutta methods (multirate
+\cite{rice:1960} developing the split Runge-Kutta methods (multi rate
 Runge-Kutta methods). The concept behind these methods is the splitting of
@@
 time step sizes can be adapted to fit the model in each step as well. See
-\cite{Kvaernoe:1999} and \cite{Gear:1984} (similar strategy for linear
-multistep methods) for more detailed descriptions of the two strategies.
+\cite{Kvaernoe:1999} and \cite{Gear:1984} (similar strategy for linear multi
+step methods) for more detailed descriptions of the two strategies.
 
-In the partial differential equations (PDE) community, the adaptive
-timestepping area was explored and developed with the goal of simulating
+In the partial differential equations (PDE) community, the adaptive time
+stepping area was explored and developed with the goal of simulating
 very specific known problems such as the (hyperbolic) wave equations
@@
 \cite{Gander:2013}. Generally the methods developed at first all make use
-of the same basic concepts for making the local adaptive timestepping:
+of the same basic concepts for making the local adaptive time stepping:
 Interpolation, extrapolation, prediction and correction, which we are going to
@@
 
-Recently newer and faster methods with local time stepping have been
-developed on the basis of nonlinear PDEs and the work performed in this
-field in the 80s and 90s. First the multiresolution (MR) schemes for creating
-space-adaptive discretizations and refinements of these were developed,
-see \cite{Berger:1984}. Then multiple MR based schemes were developed.
-Dominques et al. \cite{Domingues:2008} is one of the more recent of these MR
-schemes. They describe a local scale-dependent time stepping
-for a space-adaptive multiresolution scheme using the finite volume method
-in order to obtain speed-up using larger time-steps without violating a
-defined stability constraints, which is essentially the same motivation as
-this report. The method is based on an explicit Runge-Kutta method of second
-order. As expected the timestep size is imposed by a stability condition of
-the explicit Runge-Kutta on the finest scale, which increases with the scale
-of the mesh and hence we are able to increase the timestep as well without
-violating the stability condition.
+Recently newer and faster methods with local time stepping have been developed
+on the basis of nonlinear PDEs and the work performed in this field in
+the 80s and 90s. First the multi resolution (MR) schemes for creating
+space-adaptive discretizations and refinements of these were developed, see
+\cite{Berger:1984}. Then multiple MR based schemes were developed. Domingues
+et al. \cite{Domingues:2008} is one of the more recent of these MR schemes.
+They describe a local scale-dependent time stepping for a space-adaptive multi
+resolution scheme using the finite volume method in order to obtain speed-up
+using larger time-steps without violating a defined stability constraints,
+which is essentially the same motivation as this report. The method is based
+on an explicit Runge-Kutta method of second order. As expected the time step
+size is imposed by a stability condition of the explicit Runge-Kutta on the
+finest scale, which increases with the scale of the mesh and hence we are able
+to increase the time step as well without violating the stability condition.
 

### 2013-09-16 report/experiments.tex
Base src code and notes of experiments added

\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage{lmodern}

\usepackage{amsmath,amssymb,amsfonts}

\title{}
\author{Malte Stær Nissen}

\begin{document}
\maketitle

Til triangulering:
qmesh,
delaunay,

Store trekanter med små trekanter i midten.

Fast forhold mellem tidsskridt til at starte med.

Derefter detekt når advektionen går galt.

$\frac{\alpha t}{2} < \Delta t$

linære beregninger

Samtidig timestepping uden genberegning på store skridt

Forsøg at se hvor meget vi kan presse citronen.

Benyt eget lille framework og modificer det

Evt. benyt lokale timesteps med genberegninger løbende

\end{document}

### 2013-09-16 src/main.m

@@ -1,0 +1,175 @@
+% Clear workspace
+clear all;
+close all;
+
+% Load mesh
+
+load('data.mat');
+
+area_show = false;
+momentum_show = false;
+
+% Set simulation time
+t_start = 0;
+fps = 30;
+dt = 1 / fps;
+t_end = 20
+
+% Set simulation constants
+%G = 79.3;
+%E = 207;
+% E = Young's modulus
+%lambda = G * (E - 2*G) / (3 * G - E);
+%mu = G;
+lambda = 398;
+mu = 265;
+
+Nle = zeros(length(cntT),2,3);
+invDe0 = zeros(2,2,cntT);
+
+% Pre-compute inv(Ge0)
+% and Ne * le
+for i = 1:cntT
+    % Positions (x,y) of (i,j,k)
+    x = X(T(i,:));
+    y = Y(T(i,:));
+
+    Ge0_ji = [x(2); y(2)] - [x(1); y(1)];
+    Ge0_ki = [x(3); y(3)] - [x(1); y(1)];
+
+    invDe0(:,:,i) = inv([Ge0_ji Ge0_ki]);
+
+    Ge0_hat_ij = ([-y(1) x(1)] - [-y(2) x(2)]);
+    Ge0_hat_jk = ([-y(2) x(2)] - [-y(3) x(3)]);
+    Ge0_hat_ki = ([-y(3) x(3)] - [-y(1) x(1)]);
+
+    Nle(i,:,1) = Ge0_hat_ij;
+    Nle(i,:,2) = Ge0_hat_jk;
+    Nle(i,:,3) = Ge0_hat_ki;
+
+end
+
+% Pre-allocate P matrix [ P1 P2 ... Plength(t) ]
+P = zeros(2,2,cntT);
+v = zeros(cntV,2);
+m = ones(cntV,1) * 10;
+
+xn = [X Y];
+
+% Find left side
+val = min(X);
+ind = find(X == val);
+qind = find(X ~= val);
+
+valMax = max(X);
+rind = find(X == valMax);
+
+TR = TriRep(T,X,Y);
+boundary = freeBoundary(TR);
+
+iters = 1:1000;
+Volume = zeros(length(iters),1);
+Momentum = zeros(length(iters),1);
+
+
+fi = zeros(2,cntV);
+% Loop over time
+%for t = t_start : dt : t_end
+for step = iters
+    parfor(e = 1:cntT,cntT)
[…]

### 2013-09-16 src/main2.m

@@ -1,0 +1,121 @@
+% Clear workspace
+clear all;
+close all;
+
+% Load mesh
+
+load('data.mat');
+
+% Set simulation time
+t_start = 0;
+fps = 100;
+dt = 1 / fps;
+t_end = 20
+
+% Set simulation constants
+lambda = 398;
+% E = Young's modulus
+%E = 
+mu = 265;
+%lambda = mu * (E - 2*mu) / (3 * mu - E);
+
+Nle = zeros(length(cntT),2,3);
+invDe0 = zeros(2,2,cntT);
+
+% Pre-compute inv(Ge0)
+% and Ne * le
+for i = 1:cntT
+    % Positions (x,y) of (i,j,k)
+    x = X(T(i,:));
+    y = Y(T(i,:));
+
+    Ge0_ji = [x(2); y(2)] - [x(1); y(1)];
+    Ge0_ki = [x(3); y(3)] - [x(1); y(1)];
+
+    invDe0(:,:,i) = inv([Ge0_ji Ge0_ki]);
+
+    Ge0_hat_ij = ([-y(1) x(1)] - [-y(2) x(2)]);
+    Ge0_hat_jk = ([-y(2) x(2)] - [-y(3) x(3)]);
+    Ge0_hat_ki = ([-y(3) x(3)] - [-y(1) x(1)]);
+
+    Nle(i,:,1) = Ge0_hat_ij;
+    Nle(i,:,2) = Ge0_hat_jk;
+    Nle(i,:,3) = Ge0_hat_ki;
+
+end
+
+% Pre-allocate P matrix [ P1 P2 ... Plength(t) ]
+P = zeros(2,2,cntT);
+v = zeros(cntV,2);
+m = ones(cntV,1) * 10;
+
+xn = [X Y];
+
+% Find left side
+val = min(X);
+ind = find(X == val);
+
+
+% Loop over time
+%for t = t_start : dt : t_end
+for step = 1:2000
+    parfor(e = 1:cntT,cntT)
+        % Positions (x,y) of (i,j,k)
+        x = xn(T(e,:),1);
+        y = xn(T(e,:),2);
+
+        ge_ji = [x(2); y(2)] - [x(1); y(1)];
+        ge_ki = [x(3); y(3)] - [x(1); y(1)];
+
+        De = [ge_ji ge_ki];
+
+        Fe = De * invDe0(:,:,e);
+
[…]

### 2013-09-19 src/dualform.m
Added notes for 1d spring experiments

@@ -1,0 +1,6 @@
+function h=dualform(p, varargin)
+    h = ones(size(p,1),1);
+
+    h(all(p >= -2,2) & all(p <= 2,2)) = 0.6;
+        fh=@(p) 0.05+0.3*dcircle(p,0,0,0.5);
+end

### 2013-09-19 src/main.m

@@ -151,7 +151,7 @@ for step = iters
 %    ylabel('y');
 %    set(fi1, 'PaperPosition', [-0.2 -0.25 6 5.5])
 %    set(fi1, 'PaperSize', [5.4 5.2]);
-%    saveas(fi1,['../images/push_frame_' num2str(step)],'pdf');
+%    saveas(fi1,['images/push_frame_' num2str(step)],'pdf');
 %    end
 end
 
@@ -163,7 +163,7 @@ ylabel('Volume');
 title('Volume');
 set(fi2, 'PaperPosition', [-0.2 -0.25 6 5.5])
 set(fi2, 'PaperSize', [5.4 5.2]);
-saveas(fi2, '../images/push_volume','pdf');
+saveas(fi2, 'images/push_volume','pdf');
 
 fi3 = figure(3);
 plot(1:step,Momentum(1:step));
@@ -172,4 +172,4 @@ xlabel('frame');
 set(fi3, 'PaperPosition', [-0.2 -0.25 6 5.5])
 set(fi3, 'PaperSize', [5.4 5.2]);
 ylabel('Sum of length of velocity');
-saveas(fi3, '../images/push_momentum','pdf');
+saveas(fi3, 'images/push_momentum','pdf');

### 2013-09-19 src/simpleMesh.m

@@ -1,0 +1,31 @@
+clear all;
+close all;
+%fd=@(p) sqrt(sum(p.^2,2))-1;
+
+fd=@(p) drectangle0(p,-2,2,-2,2);
+fh=@(p) 0.05+0.25*dcircle(p,0,0,0.05);
+figure;
+%[p,t]=distmesh2d(fd,@huniform,0.5,[-2,-2;2,2],[]);
+[p,t]=distmesh2d(fd,fh,0.05,[-1,-1;1,1],[-1,-1;-1,1;1,-1;1,1]);
+%p2 = [];
+
+%[X,Y] = meshgrid(-2:0.5:2,-2:0.5:2);
+%X2 = X(1:2:end,1:2:end) + 0.25;
+%Y2 = Y(1:2:end,1:2:end) + 0.25;
+
+%p = [X(:),Y(:)];
+
+%{
+
+p2 = p(all(p >= -1,2) & all(p <= 1,2),:);
+p2(:,1) = p2(:,1) + 0.2;
+p2(:,2) = p2(:,2) + 0.2;
+figure;
+[p2,t2]=distmesh2d(fd,@huniform,0.3,[-2,-2;2,2],[]);
+%}
+p3 = [p;p2];
+t3 = delaunay(p3(:,1),p3(:,2));
+
+figure;
+triplot(t3, p3(:,1), p3(:,2));
+axis equal;

### 2013-09-30 report/experiments.tex
Added notes for 1d spring experiments
Added experiments plan and modified spring system

@@
 
-\title{}
+\title{Experimental plan}
 \author{Malte Stær Nissen}
@@
 
-Til triangulering:
-qmesh,
-delaunay,
+The experimental plan is overall split in two parts: 1 dimensional experiments and 2
+dimensional experiments.
 
-Store trekanter med små trekanter i midten.
+\subsection{1 dimensional experiments}
+We are working with the problem of simulating a spring mass particle system in 1
+dimension. Our hypothesis is that we are able to apply adaptive timestepping to
+the simulation and thereby decrease the number of computations needed to
+simulate the system.
 
-Fast forhold mellem tidsskridt til at starte med.
+We first experiment with a uniform 1D mesh where we manually choose different
+timestep sizes to test for possible side-effects in as simple a simulation as
+possible.
 
-Derefter detekt når advektionen går galt.
+We then perform the adaptive timestepping on a 1D mesh having coarse and finer
+parts.
 
-$\frac{\alpha t}{2} < \Delta t$
+\subsection{2 dimensional experiments}
 
-linære beregninger
+In 2 dimensions we are working with the problem of simulating hyperelastic
+materials using adaptive timestepping as for the 1 dimensional experiments.
 
-Samtidig timestepping uden genberegning på store skridt
+We split the 2D experiments in a test on a somewhat uniform 2D mesh having
+manually chosen some triangles to be simumlated with smaller timesteps than
+others. After these experiments we move on to performing 2D experiments on
+space-adaptive meshes having both coarse and fine parts.
 
-Forsøg at se hvor meget vi kan presse citronen.
+\subsection{Comparative experiments}
 
-Benyt eget lille framework og modificer det
+Finally we are going to compare the performance of the adaptive timestepping
+with regular timestepping schemes in 2 dimensions.
 
-Evt. benyt lokale timesteps med genberegninger løbende
+%Lav 1 dimensionelle problemer først med fjedrer
+%
+%Til triangulering:
+%qmesh,
+%delaunay,
+%
+%Store trekanter med små trekanter i midten.
+%
+%Fast forhold mellem tidsskridt til at starte med.
+%
+%Derefter detekt når advektionen går galt.
+%Når arealet er lig 0
+%Krydsprodukt til udregningerne.
+%
+%$\frac{\alpha t}{2} < \Delta t$
+%
+%linære beregninger
+%
+%Samtidig timestepping uden genberegning på store skridt
+%
+%Forsøg at se hvor meget vi kan presse citronen.
+%
+%Benyt eget lille framework og modificer det
+%
+%
+%Evt. benyt lokale timesteps med genberegninger løbende
 

### 2013-09-30 src/1D/spring.m

@@ -3,7 +3,7 @@ close all;
 
 % Create grid of particles at positions
 pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
-pP = [-2:1:2];
+%pP = [-2:1:2];
 if size(pP,2) > size(pP,1)
     pP = pP';
 end
@@ -12,7 +12,7 @@ cntP = size(pP,1);
 % Set up initial measurements for each particle
 pV = zeros(cntP,1);         % Initial velocities
 pF = zeros(cntP,1);         % Force at time t
-pM = ones(cntP,1) * 500;    % Mass of each particle
+pM = ones(cntP,1) * 30;    % Mass of each particle
 
 % Fix the left-most point
 % Find left side
@@ -26,10 +26,10 @@ sp = [sp(1:end-1),sp(2:end)];
 cntS = size(sp,1);
 
 % Set up initial measurements for each spring
-spC = ones(cntS,1) * 0.01;   % Damping
-spK = ones(cntS,1) * 0.1;   % Stiffness
+spC = ones(cntS,1) * 0.5;   % Damping
+spK = ones(cntS,1) * 0.5;   % Stiffness
 spX0 = pP(sp);
-spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.9;  % Rest length (x0)
+spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.8;  % Rest length (x0)
 spD = ones(cntS,1) * 0.1;   % Viscous drag
 
 
@@ -42,7 +42,7 @@ distance_show = false;
 momentum_show = false;
 
 t_start = 0;
-t_stop = 100;
+t_stop = 500;
 fps_lowest = 10;
 dt_base = 1 / fps_lowest;
 dt = dt_base;
@@ -59,6 +59,10 @@ explicit = true;
 h = 1;
 
 
+positions = [];
+lengths = [];
+times = [];
+dx = zeros(cntS,1);
 while time < t_stop
 
     pF = zeros(cntP,1);
@@ -75,6 +79,10 @@ while time < t_stop
     f = f + spC .* dv .* dx ./ l;
     f = f .* (-dx ./ l);
 
+%    if time == 0
+%        f(end) = 50;
+%    end
+
 %    pFfixed = pF(pFixed);
     pF(sp(:,1)) = pF(sp(:,1)) + f;
     pF(sp(:,2)) = pF(sp(:,2)) - f;
[…]

### 2013-09-30 src/1D/springWfor.m

@@ -30,7 +30,6 @@ spC = ones(cntS,1) * 0.5;   % Damping
 spK = ones(cntS,1) * 0.5;   % Stiffness
 spX0 = pP(sp);
 spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.8;
-%spX0 = abs(spX0(:,1) - spX0(:,2));  % Rest length (x0)
 spD = ones(cntS,1) * 0.1;   % Viscous drag
 
 
@@ -124,9 +123,8 @@ figure(2);
 plot(times,positions);
 
 figure(3);
-figure('Renderer','OpenGL');
 y_values = zeros(size(positions,1),1);
-plot(positions(:,1),y_values,'x')
+plot(positions(:,1),y_values,'x');
 set(gca,'NextPlot','replaceChildren');
 for i=1:length(times)
     plot(positions(:,i),y_values,'xr');

### 2013-10-02 src/1D/displaySim.m
Modularized 1D code and started working on time adaptation

@@ -1,0 +1,45 @@
+function displaySim(data, fps)
+    movie = false;
+    animation = false;
+
+    figure();
+    cntD = size(data,1);
+    for i = 1:cntD
+        i0 = i-1;
+        plot1 = i;
+        plot2 = plot1+3;
+        subplot(2, cntD, plot1);
+        plot(data{i,3},data{i,2});
+        title(['Spring lengths for ' num2str(fps(i)) ' fps']);
+        xlabel('Time');
+        ylabel('Spring length');
+        axis([0 max(data{i,3}) 0.2 0.55]);
+        subplot(2, cntD, plot2);
+        plot(data{i,3},data{i,1});
+        title(['Simulation trace ' num2str(fps(i)) ' fps']);
+        xlabel('Time');
+        ylabel('Position');
+        axis([0 max(data{i,3}) -2 7]);
+    end
+
+%    if animation
+%        figure();
+%        y_values = zeros(size(positions,1),1);
+%        plot(positions(:,1),y_values,'x');
+%        set(gca,'NextPlot','replaceChildren');
+%        for i=1:length(times)
+%            plot(positions(:,i),y_values,'xr');
+%            hold on;
+%            title(['Time frame: ' num2str(times(i))]);
+%            %    drawnow;
+%            hold off;
+%            %    pause(0.1);
+%            F(i) = getframe;
+%        end
+%    end
+%
+%    if movie
+%        figure();
+%        movie(F,1,fps);
+%    end
+end

### 2013-10-02 src/1D/run.m

@@ -1,0 +1,12 @@
+clear all;
+close all;
+t_stop = 120;
+fpses = [10, 2, 1];
+results = cell(length(fpses),3);
+for i = 1:length(fpses)
+    [results{i,1}, results{i,2}, results{i,3}] = springSim(fpses(i),t_stop);
+end
+
+displaySim(results, fpses);
+
+diffSim(results, fpses);

### 2013-10-02 src/1D/spring.m

@@ -3,7 +3,7 @@ close all;
 
 % Create grid of particles at positions
 pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
-%pP = [-2:1:2];
+pP = [-2:1:2];
 if size(pP,2) > size(pP,1)
     pP = pP';
 end
@@ -29,7 +29,7 @@ cntS = size(sp,1);
 spC = ones(cntS,1) * 0.5;   % Damping
 spK = ones(cntS,1) * 0.5;   % Stiffness
 spX0 = pP(sp);
-spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.8;  % Rest length (x0)
+spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.8;
 spD = ones(cntS,1) * 0.1;   % Viscous drag
 
 
@@ -43,7 +43,7 @@ momentum_show = false;
 
 t_start = 0;
 t_stop = 500;
-fps_lowest = 10;
+fps_lowest = 30;
 dt_base = 1 / fps_lowest;
 dt = dt_base;
 
@@ -58,49 +58,39 @@ time = 0;
 explicit = true;
 h = 1;
 
-
 positions = [];
 lengths = [];
 times = [];
 dx = zeros(cntS,1);
-while time < t_stop
-
+while time <= t_stop
+    f = zeros(cntS,1);
     pF = zeros(cntP,1);
-    %% Force calculation
-    % Calculate spring lengths
-    pP_temp = pP(sp);
-    dx = pP_temp(:,1) - pP_temp(:,2);       % start - end: L
-    l = abs(pP_temp(:,1) - pP_temp(:,2));   % length l
+    for s = 1:cntS
+        from = sp(s,1);
+        to = sp(s,2);
 
-    pV_temp = pV(sp);                       % velocities
-    dv = pV_temp(:,1) - pV_temp(:,2);       % Difference in velocities
+        dx(s) = pP(from) - pP(to);
+        l = abs(dx(s));
+        spL(s) = dx(s);
 
-    f = spK .* (l - spX0);
-    f = f + spC .* dv .* dx ./ l;
-    f = f .* (-dx ./ l);
+        dv = pV(from) - pV(to);
+        f = spK(1) * (l - spX0(s));
+        f = f + spC(1) * dv * dx(s) / l;
+        f = f * (- dx(s) / l);
+        pF(from) = pF(from) + f;
+        pF(to) = pF(to) - f;
+    end
 
[…]

### 2013-10-02 src/1D/springForces.m

@@ -1,0 +1,31 @@
+function [F, dx] = springForces(positions, springs, dt, v, K, x0, C)
+    % Initialize internal variables
+    cntS = size(springs,1);
+    cntP = size(positions,1);
+    Fs  = zeros(cntS,1);
+
+    % Initialize output 
+    F   = zeros(cntP,2);
+    dx = zeros(cntS,1);
+
+    % Calculate force F of each spring
+    for s = 1:cntS
+        % Endpoints of spring
+        from = springs(s,1);
+        to = springs(s,2);
+
+        % Lengh of spring
+        dx(s) = springLength(positions, springs(s,:));
+        l = abs(dx(s));
+
+        % Force calculation
+        Fs = K(s) * (l - x0(s));                        % Spring force
+        Fs = Fs + C(s) * (v(from) - v(to)) * dx(s) / l; % Dampening
+        Fs = Fs * (- dx(s) / l);                        % Direction
+
+        % Apply forces to endpoints
+        F(from,1) = F(from,1) + Fs;
+        F(to,2) = F(to,2) - Fs;
+    end
+
+end

### 2013-10-02 src/1D/springLength.m

@@ -1,0 +1,8 @@
+function [l] = springLength(positions, spring)
+    % Get indices
+    left  = positions(spring(:,1));
+    right = positions(spring(:,2));
+
+    % Compute length (left - right)
+    l = left - right;
+end

### 2013-10-02 src/1D/springSim.m

@@ -1,0 +1,113 @@
+function [positions, lengths, times] = springSim(fps_lowest, t_stop)
+    % Create grid of particles at positions
+    pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
+    %pP = [-2:1:2];
+    if size(pP,2) > size(pP,1)
+        pP = pP';
+    end
+
+    % Initialize variables
+    cntP = size(pP,1);
+
+    % Set up initial measurements for each particle
+    pV = zeros(cntP,1);         % Initial velocities
+    pF = zeros(cntP,1);         % Force at time t
+    pM = ones(cntP,1) * 30;    % Mass of each particle
+
+    % Fix the left-most point
+    % Find left side
+    val = min(pP);
+    pFixed = find(pP == val);
+
+    % Register all springs
+    % Springs between all inner particles (indices)
+    sp = [1:cntP]';
+    sp = [sp(1:end-1),sp(2:end)];
+    cntS = size(sp,1);
+
+    % Set up initial measurements for each spring
+    spC = ones(cntS,1) * 0.5;   % Damping
+    spK = ones(cntS,1) * 0.5;   % Stiffness
+    spX0 = pP(sp);
+    spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.8;
+    spD = ones(cntS,1) * 0.1;   % Viscous drag
+
+    distance_show = false;
+    momentum_show = false;
+
+    t_start = 0;
+    if t_stop == 0
+        t_stop = 500;
+    end
+    if fps_lowest == 0
+        fps_lowest = 30;
+    end
+    disp(fps_lowest);
+    dt_base = 1 / fps_lowest;
+    dt = dt_base;
+
+    valMax = max(pP);
+    rind = find(pP == valMax);
+
+    % External force:
+    f_ext = 0;
+
+    % Initial time
+    time = 0;
+    explicit = true;
+    h = 1;
+
+    positions = [];
+    lengths = [];
+    times = [];
+
+    % subdivide particles
+    pSub = [];
+    pSt = [];
+    pSdt = [];
+
+    dx = zeros(cntS,1);
+    pF = zeros(cntP,2);
+    while time <= t_stop + 1e-5;
+
+        % Check for subdivision
+        if dt < dt_base && length(pSub) > 0
+
+            % Find springs to re-calculate force for
+            springs = sp(ismember(sp,pSub));
+        end
+
+        % Force calculation
[…]

### 2013-10-02 src/1D/springStep.m

@@ -1,0 +1,10 @@
+function [dp, dv] = springStep(p, m, v, F, dt)
+    % Calculate derivative of velocity
+    dv = F ./ m;
+
+    % Position change
+    dp = v * dt;
+
+    % Velocity change
+    dv = dv * dt;
+end

### 2013-10-02 src/1D/springWfor.m

@@ -1,139 +1,0 @@
-clear all;
-close all;
-
-% Create grid of particles at positions
-pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
-pP = [-2:1:2];
-if size(pP,2) > size(pP,1)
-    pP = pP';
-end
-cntP = size(pP,1);
-
-% Set up initial measurements for each particle
-pV = zeros(cntP,1);         % Initial velocities
-pF = zeros(cntP,1);         % Force at time t
-pM = ones(cntP,1) * 30;    % Mass of each particle
-
-% Fix the left-most point
-% Find left side
-val = min(pP);
-pFixed = find(pP == val);
-
-% Register all springs
-% Springs between all inner particles (indices)
-sp = [1:cntP]';
-sp = [sp(1:end-1),sp(2:end)];
-cntS = size(sp,1);
-
-% Set up initial measurements for each spring
-spC = ones(cntS,1) * 0.5;   % Damping
-spK = ones(cntS,1) * 0.5;   % Stiffness
-spX0 = pP(sp);
-spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.8;
-spD = ones(cntS,1) * 0.1;   % Viscous drag
-
-
-% Plot initial particles
-figure(1);
-plot(pP,ones(1,cntP),'xr');
-title('Initial particles');
-
-distance_show = false;
-momentum_show = false;
-
-t_start = 0;
-t_stop = 500;
-fps_lowest = 30;
-dt_base = 1 / fps_lowest;
-dt = dt_base;
-
-valMax = max(pP);
-rind = find(pP == valMax);
-
-% External force:
-f_ext = 0;
-
-% Initial time
-time = 0;
-explicit = true;
-h = 1;
-
-positions = [];
-lengths = [];
-times = [];
-dx = zeros(cntS,1);
-while time < t_stop
-    f = zeros(cntS,1);
-    pF = zeros(cntP,1);
-    for s = 1:cntS
-        from = sp(s,1);
-        to = sp(s,2);
-
-        dx(s) = pP(from) - pP(to);
-        l = abs(dx(s));
-        spL(s) = dx(s);
-
-        dv = pV(from) - pV(to);
-        f = spK(1) * (l - spX0(s));
-        f = f + spC(1) * dv * dx(s) / l;
-        f = f * (- dx(s) / l);
-        pF(from) = pF(from) + f;
[…]

### 2013-10-06 report/presentation_content.tex
Initialized new simulation (atomic) and presentation of experiments

\usepackage{pslatex}
\usepackage{listing}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multirow}

\usepackage{etoolbox}% http://ctan.org/pkg/etoolbox
\AtBeginEnvironment{figure}{\setcounter{subfigure}{0}}% Resets subfigure counter at start of figure environment

\usepackage{tikz}
\usetikzlibrary{shapes,arrows}

\tikzset{onslide/.code args={<#1>#2}{%
  \only<#1>{\pgfkeysalso{#2}} % \pgfkeysalso doesn't change the path
}}

\usepackage{scalefnt}

\usetheme[unit=ics,dk]{Frederiksberg}
\useinnertheme{MLHacks}

\AtBeginSection[]
{
	\begin{frame}<beamer>
		\frametitle{Emne}
		\tableofcontents[currentsection,hideothersubsections]
		\ghostframe
	\end{frame}
}

\title{Local Adaptive Optimization of Time Stepi - Experimental plan}
\subtitle{CMM Project Course}
\author{Malte Stær Nissen}
\institute{Datalogisk Institut}
\date{\today}

\begin{document}

\tikzstyle{decision} = [diamond, draw, fill=blue!20, 
    text width=7em, text badly centered, node distance=1.3cm, inner sep=0pt]
\tikzstyle{block} = [rectangle, draw, fill=gray!20, 
    text width=12em, text centered, minimum height=3em]
\tikzstyle{line} = [draw, -latex']
\tikzstyle{cloud} = [draw, ellipse,fill=red!20, node distance=2.3cm, minimum height=2em]

\frame[plain]{\titlepage}

\begin{frame}{Content}
	\setbeamercovered{highly dynamic}
	\begin{center}
	\tableofcontents[pausesections,hideothersubsections]
	\end{center}
\end{frame}

\section{}
\begin{frame}{Short introduction}
	\begin{itemize}
		\item Simulations of materials
		\item Heavy computations
        \item Reduce number of calculations
	\end{itemize}
\end{frame}

\section{}
\begin{frame}{Experimental plan overview}
\end{frame}

\section{Tests og resultater}

\begin{frame}{Eksempel på original maske}
		\includegraphics[height=7cm]{graphics/sa_breast_pectoral.png}
\end{frame}

\begin{frame}{Middelværdier for histogrammerne}

	{\scalefont{0.8}
	\begin{table}
		\centering
		\begin{tabular}{l l| l | l | l | l | l |}
			\cline{2-7}
			 & \multicolumn{6}{|c|}{Middelværdi} \\ \hline
			\multicolumn{1}{| l |}{Metode}		& PSE	&	PSP	&	BSE	&	BSP 	&	TSE	& TSP \\ \hline \hline
			\multicolumn{1}{| l |}{Karssemeijer $M_K$} & 0.8073 & 0.9931 & 0.8940 & 0.9603 & \multirow{2}{*}{0.9235} & \multirow{2}{*}{0.9845} \\ \cline{1-5} 
			\multicolumn{1}{| l |}{Canny  $M_C$}	& 0.7231 & 0.9982 & 0.9055 & 0.9469 & & \\ \hline
			\multicolumn{1}{| l |}{Blackbox $M_B$} & 0.7023 & 0.9542 & 0.8300 & 0.9616 & 0.9088 & 0.9745 \\ \hline
		\end{tabular}
		\caption{Oversigt over middelværdierne for testparametrene}
		\label{tbl:test_mean_values}
	\end{table}
	}
\end{frame}

\subsection{Præcisionstests}
\begin{frame}{Præcision af segmenteringen}
	\begin{itemize}
		\item Ens masker giver ens segmenteringspunkter $A$, $B$ og $C$
		\item Modellering af pektoralmusklen ved lige linje
		\item Maskerne ``modificeres'' efterfølgende af vores program
		\item Test på præcisionen af vores segmenteringer ved afvigelse mellem fundne $A_{gt}$, $B_{gt}$ og $C_{gt}$ på ground truth maskerne og $A$, $B$ og $C$ fundet vha. egen implementation
	\end{itemize}
\end{frame}

\begin{frame}{Histogrammer for præcision på $A$}
	\begin{figure}
		\begin{subfigure}{0.32\linewidth}
			\includegraphics[width=\linewidth,keepaspectratio]{graphics/karssemeijer_a_difference.png}
			\caption{$M_K, A$}
		\end{subfigure}
		\begin{subfigure}{0.32\linewidth}
			\includegraphics[width=\linewidth,keepaspectratio]{graphics/canny_a_difference.png}
			\caption{$M_C, A$}
		\end{subfigure}
		\begin{subfigure}{0.32\linewidth}
			\includegraphics[width=\linewidth,keepaspectratio]{graphics/blackbox_a_difference.png}
			\caption{$M_B, A$}
		\end{subfigure}
	\end{figure}
\end{frame}

\begin{frame}{Histogrammer for præcision på $B$ og $C$}
	\begin{figure}
		\begin{subfigure}{0.32\linewidth}
			\includegraphics[width=\linewidth,keepaspectratio]{graphics/karssemeijer_b_difference.png}
			\caption{$M_K, B$}
		\end{subfigure}
		\begin{subfigure}{0.32\linewidth}
			\includegraphics[width=\linewidth,keepaspectratio]{graphics/canny_b_difference.png}
			\caption{$M_C, B$}
		\end{subfigure}
		\begin{subfigure}{0.32\linewidth}
			\includegraphics[width=\linewidth,keepaspectratio]{graphics/blackbox_b_difference.png}
			\caption{$M_B, B$}
		\end{subfigure}
		\begin{subfigure}{0.32\linewidth}
			\includegraphics[width=\linewidth,keepaspectratio]{graphics/karssemeijer_c_difference.png}
			\caption{$M_K, C$}
		\end{subfigure}
		\begin{subfigure}{0.32\linewidth}
			\includegraphics[width=\linewidth,keepaspectratio]{graphics/canny_c_difference.png}
			\caption{$M_C, C$}
		\end{subfigure}
		\begin{subfigure}{0.32\linewidth}
			\includegraphics[width=\linewidth,keepaspectratio]{graphics/blackbox_c_difference.png}
			\caption{$M_B, C$}
		\end{subfigure}
	\end{figure}
\end{frame}

\begin{frame}{Middelværdier}
	\begin{table}
		\begin{tabular}{l | r | r | r |}
			\cline{2-4}
			{} & \multicolumn{3}{| c |}{ Middelværdi } \\ \hline
			\multicolumn{1}{| l |}{Metode}			&	$A$ (cm)		&	$B$ (cm)		& 	$C$ (cm) \\ \hline \hline
			\multicolumn{1}{| l |}{Karssemeijer $M_K$}	&	1.7202	&	2.622  	&	3.9531 \\ \hline
			\multicolumn{1}{| l |}{Canny $M_C$}		&	1.4559	&	2.300  	&	3.3958 \\ \hline
			\multicolumn{1}{| l |}{Blackbox $M_B$}		&	2.1541	&	3.8948	&	5.1714 \\ \hline
		\end{tabular}
		\caption{Middelværdier for afvigelsen på punkterne $A$, $B$ og $C$}
	\end{table}
\end{frame}
\section{Diskussion}
\subsection{Indre begrænsninger}
\begin{frame}{Systemernes begrænsninger}
	\begin{enumerate}
		\item For små mammografier i forhold til brystet
		\item Manglende skin-air grænse/brystvæv
		\item Manglende pektoralmuskel
	\end{enumerate}
\end{frame}
\subsection{Problemer ved overførsel til nyt system}
\begin{frame}{Problemer ved overførsel til nyt system}
	\begin{enumerate}
		\item Projektionsretning (ML, MLO, CC, LM)
		\item Intensitetsværdier
		\item Mammografidimensioner
	\end{enumerate}
\end{frame}

\begin{frame}{Detaljeret overblik}
	\input{flowchart/overall_detail}
\end{frame}

\end{document}

### 2013-10-06 report/report.tex
Initialized new simulation (atomic) and presentation of experiments

@@
 \title{Local Adaptive Optimization of Time step}
-\author{Malte Stær Nissen\\University of Copenhagen\\\texttt{malte.nissen@gmail.com} \and Supervisor: Sune Darkner\\University of Copenhagen\\\texttt{darkner@diku.dk}}
+\author{Malte Stær Nissen\\
+        University of Copenhagen\\
+        \texttt{malte.nissen@gmail.com} \and
+        Supervisor: Sune Darkner\\University of Copenhagen\\\texttt{darkner@diku.dk}}
 
@@
 
+\begin{align}
+F_k &= -k \\
+F_m &= -F
+\end{align}
+
 \section{Experiments and results}

### 2013-10-06 src/1D/atomic/atomicSim.m

@@ -1,0 +1,87 @@
+function [positions, lengths, times, switches, active] = springSim(fps, t_stop)
+    %% Initialize particles
+    % Create grid of particles at positions
+    pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
+    rFactor = 2;
+    stepBound = 65;
+
+    % Initialize variables
+    cntP = size(pP,1);
+    positions = [];
+    lengths = [];
+    times = [];
+    switches = [];
+    active = [];
+
+    % Set up initial measurements for each particle
+    pV = zeros(cntP,1);         % Initial velocities
+    pF = zeros(cntP,1);         % Force at time t
+    pM = ones(cntP,1) * 30;     % Mass of each particle
+
+    %% Register springs
+    % Springs between all inner particles (indices)
+    sp = [1:cntP]';
+    sp = [sp(1:end-1),sp(2:end)];
+    cntS = size(sp,1);
+
+    % Set up initial measurements for each spring
+    spC = ones(cntS,1) * 0.5;   % Damping
+    spK = ones(cntS,1) * 0.5;   % Stiffness
+    spX0 = pP(sp);
+    spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.8;
+    spD = ones(cntS,1) * 0.1;   % Viscous drag
+    spL = springLength(pP,sp);  % Length at last update
+    spSteps = zeros(cntS,1);    % Number of steps since last update
+
+    % Initial time
+    dt_base = 1 / fps;
+    dt = dt_base;
+    time = 0;
+
+    dx = zeros(cntS,1);
+    pF = zeros(cntP,2);
+
+    % Loop simulation over time
+    while time <= t_stop + dt/2;
+
+        % Find active particles pA and active springs spA
+        [pA spA spL spSteps_new] = registerActive(pP, sp, spL, pV, time, dt,...
+                                                 rFactor, spSteps, stepBound);
+        active(end+1) = length(spA);
+
+        % Force calculation of active springs/particles
+        if (length(spA) > 0)
+            [pF dx] = springForces(pP, sp(spA,:), spL, pV,...
+                                   spK(spA), spX0(spA), spC(spA));
+        end
+
+        % Position and velocity update
+        [dp dv] = springStep(pP, pM, pV, sum(pF,2), dt, pA, sp, spSteps);
+
+        % Detect collisions / switches in position
+        l = springLength(pP + dp, sp);
[…]

### 2013-10-06 src/1D/atomic/displaySim.m

@@ -1,0 +1,56 @@
+function displaySim(data, fps)
+    movie = false;
+    animation = false;
+
+    figure();
+    cntD = size(data,1);
+    for i = 1:cntD
+        i0 = i-1;
+        plot1 = i;
+        plot2 = plot1+cntD;
+        subplot(2, cntD, plot1);
+        plot(data{i,3},data{i,2});
+        title(['Spring lengths for ' num2str(fps(i)) ' fps']);
+        xlabel('Time');
+        ylabel('Spring length');
+        axis([0 max(data{i,3}) -0.3 0.55]);
+        subplot(2, cntD, plot2);
+        plot(data{i,3},data{i,1});
+        title(['Simulation trace ' num2str(fps(i)) ' fps']);
+        xlabel('Time');
+        ylabel('Position');
+        hold on;
+        for j = 1:size(data{i,4},1)
+            time = data{i,4}(j,1);
+            ind1 = data{i,4}(j,2);
+            ind2 = data{i,4}(j,3);
+            plot(time,data{i,1}(ind1, data{i,3} == time),'xr');
+            plot(time,data{i,1}(ind2, data{i,3} == time),'xr');
+        end
+        updPoints = find(data{i,5});
+        plot(data{i,3}(updPoints),data{i,1}(:,updPoints),'-xr');
+        axis([0 max(data{i,3}) -2 7]);
+        disp(length(updPoints) / max(data{i,3}));
+    end
+
+%    if animation
+%        figure();
+%        y_values = zeros(size(positions,1),1);
+%        plot(positions(:,1),y_values,'x');
+%        set(gca,'NextPlot','replaceChildren');
+%        for i=1:length(times)
+%            plot(positions(:,i),y_values,'xr');
+%            hold on;
+%            title(['Time frame: ' num2str(times(i))]);
+%            %    drawnow;
+%            hold off;
+%            %    pause(0.1);
+%            F(i) = getframe;
+%        end
+%    end
+%
+%    if movie
+%        figure();
+%        movie(F,1,fps);
+%    end
+end

### 2013-10-06 src/1D/atomic/registerActive.m

@@ -1,0 +1,34 @@
+function [pMask sMask lUpdate, steps] = registerActive(positions, springs,...
+lUpdate, velocities, time, dt, factor, steps, stepBound)
+    if nargin < 9
+        stepBound = NaN;
+    end
+
+    % New positions using existing velocity
+    positionsFull = positions + velocities * dt;
+
+    % New length
+    l = springLength(positionsFull, springs);
+
+    % Ratio between last updated length and next
+    ratio = lUpdate ./ l;
+
+    % Find indices of violating springs and points
+    sMask = ratio > factor | ratio < 1/factor;
+    if time == 0
+        sMask = ones(size(springs,1),1);
+%    else
+%        sMask = zeros(size(springs,1),1);
+    end
+    if ~isnan(stepBound)
+        sMask = sMask | steps >= stepBound;
+    end
+    % Temporary override to debug the rest of the simulator
+    sMask = find(sMask);
+    pMask = unique(springs(sMask,:));
+
+    % Update last updated lengths and number of steps since update
+    lUpdate(sMask) = l(sMask);
+    steps = steps+1;
+    steps(sMask) = 1;
+end

### 2013-10-06 src/1D/atomic/run.m

@@ -1,0 +1,13 @@
+clear all;
+close all;
+t_stop = 500;
+fpses = [10,10];
+results = cell(length(fpses),5);
+for i = 1:length(fpses)
+    [results{i,1}, results{i,2}, results{i,3}, results{i,4}, results{i,5}] = atomicSim(fpses(i),t_stop);
+%    [results{i,1}, results{i,2}, results{i,3}] = springSim(fpses(i),t_stop);
+end
+
+displaySim(results, fpses);
+
+%diffSim(results, fpses);

### 2013-10-06 src/1D/atomic/springForces.m

@@ -1,0 +1,49 @@
+function [F, dx] = springForces(positions, springs, lUpdate, v, K, x0, C)
+    element = true;
+    % Initialize internal variables
+    cntS = size(springs,1);
+    cntP = size(positions,1);
+    Fs  = zeros(cntS,1);
+
+    % Initialize output 
+    F   = zeros(cntP,2);
+    dx = zeros(cntS,1);
+
+    % Calculate force F of each spring
+    if element
+        for s = 1:cntS
+            % Endpoints of spring
+            from = springs(s,1);
+            to = springs(s,2);
+
+            % Lengh of spring
+            dx(s) = springLength(positions, springs(s,:));
+            l = abs(dx(s));
+
+            % Force calculation
+            Fs = K(s) * (l - x0(s));                        % Spring force
+            Fs = Fs + C(s) * (v(from) - v(to)) * dx(s) / l; % Dampening
+            Fs = Fs * (- dx(s) / l);                        % Direction
+
+            % Apply forces to endpoints
+            F(from,1) = F(from,1) + Fs;
+            F(to,2) = F(to,2) - Fs;
+        end
+    else
+        % Matrix alternative to spring (element) wise computation
+        from = springs(:,1);
+        to   = springs(:,2);
+
+        dx = springLength(positions, springs);
+        l  = abs(dx);
+
+        Fs = K .* (l - x0);
+        Fs = Fs + C .* (v(from) - v(to)) .* dx ./ l;
+        Fs = Fs .* (-dx ./ l);
+
+        F(from,1) = F(from,1) + Fs;
+        F(to,2)   = F(to,2) - Fs;
+
+    end
+
+end

### 2013-10-06 src/1D/atomic/springStep.m

@@ -1,0 +1,14 @@
+function [dp, dv] = springStep(p, m, v, F, dt, activeP, springs, steps)
+    cntA = size(activeP,1);
+    pSprings = zeros(cntA,1);
+    for i = 1:cntA
+        pSprings(i) = find(springs(:,1) == activeP(i) |...
+                           springs(:,2) == activeP(i),1);
+    end
+
+    % Position change
+    dp = v * dt;
+
+    % Velocity change
+    dv = F(activeP) ./ m(activeP) .* dt .* steps(pSprings);
+end

### 2013-10-06 src/1D/run.m

@@ -1,12 +1,0 @@
-clear all;
-close all;
-t_stop = 120;
-fpses = [10, 2, 1];
-results = cell(length(fpses),3);
-for i = 1:length(fpses)
-    [results{i,1}, results{i,2}, results{i,3}] = springSim(fpses(i),t_stop);
-end
-
-displaySim(results, fpses);
-
-diffSim(results, fpses);

### 2013-10-06 src/1D/spring.m

@@ -1,139 +1,0 @@
-clear all;
-close all;
-
-% Create grid of particles at positions
-pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
-pP = [-2:1:2];
-if size(pP,2) > size(pP,1)
-    pP = pP';
-end
-cntP = size(pP,1);
-
-% Set up initial measurements for each particle
-pV = zeros(cntP,1);         % Initial velocities
-pF = zeros(cntP,1);         % Force at time t
-pM = ones(cntP,1) * 30;    % Mass of each particle
-
-% Fix the left-most point
-% Find left side
-val = min(pP);
-pFixed = find(pP == val);
-
-% Register all springs
-% Springs between all inner particles (indices)
-sp = [1:cntP]';
-sp = [sp(1:end-1),sp(2:end)];
-cntS = size(sp,1);
-
-% Set up initial measurements for each spring
-spC = ones(cntS,1) * 0.5;   % Damping
-spK = ones(cntS,1) * 0.5;   % Stiffness
-spX0 = pP(sp);
-spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.8;
-spD = ones(cntS,1) * 0.1;   % Viscous drag
-
-
-% Plot initial particles
-figure(1);
-plot(pP,ones(1,cntP),'xr');
-title('Initial particles');
-
-distance_show = false;
-momentum_show = false;
-
-t_start = 0;
-t_stop = 500;
-fps_lowest = 30;
-dt_base = 1 / fps_lowest;
-dt = dt_base;
-
-valMax = max(pP);
-rind = find(pP == valMax);
-
-% External force:
-f_ext = 0;
-
-% Initial time
-time = 0;
-explicit = true;
-h = 1;
-
-positions = [];
-lengths = [];
-times = [];
-dx = zeros(cntS,1);
-while time <= t_stop
-    f = zeros(cntS,1);
-    pF = zeros(cntP,1);
-    for s = 1:cntS
-        from = sp(s,1);
-        to = sp(s,2);
-
-        dx(s) = pP(from) - pP(to);
-        l = abs(dx(s));
-        spL(s) = dx(s);
-
-        dv = pV(from) - pV(to);
-        f = spK(1) * (l - spX0(s));
-        f = f + spC(1) * dv * dx(s) / l;
-        f = f * (- dx(s) / l);
-        pF(from) = pF(from) + f;
[…]

### 2013-10-06 src/1D/switch/displaySim.m

@@ -13,12 +13,20 @@ function displaySim(data, fps)
         title(['Spring lengths for ' num2str(fps(i)) ' fps']);
         xlabel('Time');
         ylabel('Spring length');
-        axis([0 max(data{i,3}) 0.2 0.55]);
+        axis([0 max(data{i,3}) -0.3 0.55]);
         subplot(2, cntD, plot2);
         plot(data{i,3},data{i,1});
         title(['Simulation trace ' num2str(fps(i)) ' fps']);
         xlabel('Time');
         ylabel('Position');
+        hold on;
+        for j = 1:size(data{i,4},1)
+            time = data{i,4}(j,1);
+            ind1 = data{i,4}(j,2);
+            ind2 = data{i,4}(j,3);
+            plot(time,data{i,1}(ind1, data{i,3} == time),'xr');
+            plot(time,data{i,1}(ind2, data{i,3} == time),'xr');
+        end
         axis([0 max(data{i,3}) -2 7]);
     end
 

### 2013-10-06 src/1D/switch/run.m

@@ -1,0 +1,12 @@
+clear all;
+close all;
+t_stop = 200;
+fpses = [10, 2, 1];
+results = cell(length(fpses),4);
+for i = 1:length(fpses)
+    [results{i,1}, results{i,2}, results{i,3}, results{i,4}] = springSim(fpses(i),t_stop);
+end
+
+displaySim(results, fpses);
+
+%diffSim(results, fpses);

### 2013-10-06 src/1D/switch/springForces.m

@@ -1,4 +1,4 @@
-function [F, dx] = springForces(positions, springs, dt, v, K, x0, C)
+function [F, dx] = springForces(positions, springs, v, K, x0, C)
     % Initialize internal variables
     cntS = size(springs,1);
     cntP = size(positions,1);

### 2013-10-06 src/1D/switch/springSim.m

@@ -1,4 +1,4 @@
-function [positions, lengths, times] = springSim(fps_lowest, t_stop)
+function [positions, lengths, times, switches] = springSim(fps_lowest, t_stop)
     % Create grid of particles at positions
     pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
     %pP = [-2:1:2];
@@ -60,6 +60,7 @@ function [positions, lengths, times] = springSim(fps_lowest, t_stop)
     positions = [];
     lengths = [];
     times = [];
+    switches = [];
 
     % subdivide particles
     pSub = [];
@@ -78,7 +79,7 @@ function [positions, lengths, times] = springSim(fps_lowest, t_stop)
         end
 
         % Force calculation
-        [pF dx] = springForces(pP, sp, dt, pV, spK, spX0, spC);
+        [pF dx] = springForces(pP, sp, pV, spK, spX0, spC);
 
         % Position and velocity update
         [dp dv] = springStep(pP, pM, pV, sum(pF,2), dt);
@@ -86,25 +87,29 @@ function [positions, lengths, times] = springSim(fps_lowest, t_stop)
         % Detect collisions / switches in position
         l = springLength(pP + dp, sp);
         l2 = springLength(pP, sp);
-        mask = l .* l2 < 0 & l+l2 < 0;
+        mask = (l .* l2 < 0) & (l+l2 < 0);
         if any(mask)
-            disp('Switch!');
+            dt = 1/30;
+            disp('Switch in next iteration!');
             disp(time);
             pSub = unique(sp(mask,:));
+            cnt = size(pSub,1);
+            switches(end+1:end+cnt,:) = [repmat(time,[cnt 1]), pSub];
             dt = dt/10;
+            pV(pSub) = 0;
+            [pF dx] = springForces(pP, sp, pV, spK, spX0, spC);
 
-            [pF dx] = springForces(pP, sp, dt, pV, spK, spX0, spC);
             [dp dv] = springStep(pP, pM, pV, sum(pF,2), dt);
         end
-
-        % Subdivide
-        pP = pP + dp;
-        pV = pV + dv;
 
         % Registration
         positions(:,end+1) = pP;
         lengths(:,end+1) = -dx;
         times(end+1) = time;
+
+        % Subdivide
+        pP = pP + dp;
+        pV = pV + dv;
 
         % Time update
         time = time+dt;

### 2013-10-06 report/presentation_content.tex
Presentation nearly done

@@
 
-\usetheme[unit=ics,dk]{Frederiksberg}
+\usetheme[unit=ics]{Frederiksberg}
 \useinnertheme{MLHacks}
@@
 	\begin{frame}<beamer>
-		\frametitle{Emne}
+		\frametitle{Agenda}
 		\tableofcontents[currentsection,hideothersubsections]
@@
 
-\title{Local Adaptive Optimization of Time Stepi - Experimental plan}
-\subtitle{CMM Project Course}
+\title{Local Adaptive Optimization of Time Step}
+\subtitle{Experimental plan}
 \author{Malte Stær Nissen}
@@
 
-\begin{frame}{Content}
+\begin{frame}{Agenda}
 	\setbeamercovered{highly dynamic}
@@
 
-\section{}
-\begin{frame}{Short introduction}
+\section{The project in short}
+\begin{frame}{Short introduction to the project}
     \begin{itemize}
 		\item Simulations of materials
-		\item Heavy computations
-        \item Reduce number of calculations
+		\item Huge amount of computations for large materials
+        \item Reduce avg. number of calculations/time
 	\end{itemize}
@@
 
-\section{}
-\begin{frame}{Experimental plan overview}
+\begin{frame}{Illustration of motivation}
+    \begin{figure}
+        \center
+        \includegraphics[width=\linewidth,keepaspectratio]{../images/spring_motivation.png}
+        \caption{Simulations of mass spring particle system}
+        \label{fig:motivation}
+    \end{figure}
 \end{frame}
 
-\section{Tests og resultater}
-
-\begin{frame}{Eksempel på original maske}
-		\includegraphics[height=7cm]{graphics/sa_breast_pectoral.png}
-\end{frame}
-
-\begin{frame}{Middelværdier for histogrammerne}
-
-	{\scalefont{0.8}
-	\begin{table}
-		\centering
-		\begin{tabular}{l l| l | l | l | l | l |}
-			\cline{2-7}
-			 & \multicolumn{6}{|c|}{Middelværdi} \\ \hline
-			\multicolumn{1}{| l |}{Metode}		& PSE	&	PSP	&	BSE	&	BSP 	&	TSE	& TSP \\ \hline \hline
-			\multicolumn{1}{| l |}{Karssemeijer $M_K$} & 0.8073 & 0.9931 & 0.8940 & 0.9603 & \multirow{2}{*}{0.9235} & \multirow{2}{*}{0.9845} \\ \cline{1-5} 
-			\multicolumn{1}{| l |}{Canny  $M_C$}	& 0.7231 & 0.9982 & 0.9055 & 0.9469 & & \\ \hline
-			\multicolumn{1}{| l |}{Blackbox $M_B$} & 0.7023 & 0.9542 & 0.8300 & 0.9616 & 0.9088 & 0.9745 \\ \hline
-		\end{tabular}
-		\caption{Oversigt over middelværdierne for testparametrene}
-		\label{tbl:test_mean_values}
-	\end{table}
-	}
-\end{frame}
-
-\subsection{Præcisionstests}
-\begin{frame}{Præcision af segmenteringen}
+\section{Experimental plan}
+\begin{frame}{Overview of experiments}
     \begin{itemize}
-		\item Ens masker giver ens segmenteringspunkter $A$, $B$ og $C$
-		\item Modellering af pektoralmusklen ved lige linje
-		\item Maskerne ``modificeres'' efterfølgende af vores program
-		\item Test på præcisionen af vores segmenteringer ved afvigelse mellem fundne $A_{gt}$, $B_{gt}$ og $C_{gt}$ på ground truth maskerne og $A$, $B$ og $C$ fundet vha. egen implementation
+        \item 1D experiments (simple)
+        \item 2D experiments (more complex)
+        \item Apply methods for uniform meshes 
+        \item Apply methods for multi-scale meshes
+        \item Compare method with regular timestepping
     \end{itemize}
@@
 
-\begin{frame}{Histogrammer for præcision på $A$}
-	\begin{figure}
-		\begin{subfigure}{0.32\linewidth}
-			\includegraphics[width=\linewidth,keepaspectratio]{graphics/karssemeijer_a_difference.png}
-			\caption{$M_K, A$}
-		\end{subfigure}
-		\begin{subfigure}{0.32\linewidth}
-			\includegraphics[width=\linewidth,keepaspectratio]{graphics/canny_a_difference.png}
-			\caption{$M_C, A$}
-		\end{subfigure}
-		\begin{subfigure}{0.32\linewidth}
-			\includegraphics[width=\linewidth,keepaspectratio]{graphics/blackbox_a_difference.png}
-			\caption{$M_B, A$}
-		\end{subfigure}
-	\end{figure}
+\subsection{1D experiments}
+\begin{frame}{1D experiments}
+    \begin{itemize}
+        \item Application: Mass spring particle system
+        \item Hook's law
+        \item Simplicity for ease of development
+    \end{itemize}
 \end{frame}
 
-\begin{frame}{Histogrammer for præcision på $B$ og $C$}
-	\begin{figure}
-		\begin{subfigure}{0.32\linewidth}
-			\includegraphics[width=\linewidth,keepaspectratio]{graphics/karssemeijer_b_difference.png}
-			\caption{$M_K, B$}
-		\end{subfigure}
-		\begin{subfigure}{0.32\linewidth}
-			\includegraphics[width=\linewidth,keepaspectratio]{graphics/canny_b_difference.png}
-			\caption{$M_C, B$}
-		\end{subfigure}
-		\begin{subfigure}{0.32\linewidth}
-			\includegraphics[width=\linewidth,keepaspectratio]{graphics/blackbox_b_difference.png}
-			\caption{$M_B, B$}
-		\end{subfigure}
-		\begin{subfigure}{0.32\linewidth}
-			\includegraphics[width=\linewidth,keepaspectratio]{graphics/karssemeijer_c_difference.png}
-			\caption{$M_K, C$}
-		\end{subfigure}
-		\begin{subfigure}{0.32\linewidth}
-			\includegraphics[width=\linewidth,keepaspectratio]{graphics/canny_c_difference.png}
-			\caption{$M_C, C$}
-		\end{subfigure}
-		\begin{subfigure}{0.32\linewidth}
-			\includegraphics[width=\linewidth,keepaspectratio]{graphics/blackbox_c_difference.png}
-			\caption{$M_B, C$}
-		\end{subfigure}
-	\end{figure}
+\subsection{2D experiments}
+\begin{frame}{2D experiments}
+    \begin{itemize}
+        \item Extend 1D scheme
+        \item Application: Hyperelastic materials
+        \item Stress and strain
+    \end{itemize}
 \end{frame}
 
-\begin{frame}{Middelværdier}
-	\begin{table}
-		\begin{tabular}{l | r | r | r |}
-			\cline{2-4}
-			{} & \multicolumn{3}{| c |}{ Middelværdi } \\ \hline
-			\multicolumn{1}{| l |}{Metode}			&	$A$ (cm)		&	$B$ (cm)		& 	$C$ (cm) \\ \hline \hline
-			\multicolumn{1}{| l |}{Karssemeijer $M_K$}	&	1.7202	&	2.622  	&	3.9531 \\ \hline
-			\multicolumn{1}{| l |}{Canny $M_C$}		&	1.4559	&	2.300  	&	3.3958 \\ \hline
-			\multicolumn{1}{| l |}{Blackbox $M_B$}		&	2.1541	&	3.8948	&	5.1714 \\ \hline
-		\end{tabular}
-		\caption{Middelværdier for afvigelsen på punkterne $A$, $B$ og $C$}
-	\end{table}
-\end{frame}
-\section{Diskussion}
-\subsection{Indre begrænsninger}
-\begin{frame}{Systemernes begrænsninger}
-	\begin{enumerate}
-		\item For små mammografier i forhold til brystet
-		\item Manglende skin-air grænse/brystvæv
-		\item Manglende pektoralmuskel
-	\end{enumerate}
-\end{frame}
-\subsection{Problemer ved overførsel til nyt system}
-\begin{frame}{Problemer ved overførsel til nyt system}
-	\begin{enumerate}
-		\item Projektionsretning (ML, MLO, CC, LM)
-		\item Intensitetsværdier
-		\item Mammografidimensioner
-	\end{enumerate}
-\end{frame}
-
-\begin{frame}{Detaljeret overblik}
-	\input{flowchart/overall_detail}
+\subsection{Performance experiments}
+\begin{frame}{Performance experiments}
+    \begin{itemize}
+        \item Total simulation time
+    \end{itemize}
 \end{frame}

### 2013-10-06 src/1D/atomic/displaySim.m
Modified switch to give motivational images

@@ -1,6 +1,7 @@
 function displaySim(data, fps)
     movie = false;
     animation = false;
+    timesteps = false;
 
     figure();
     cntD = size(data,1);
@@ -27,10 +28,12 @@ function displaySim(data, fps)
             plot(time,data{i,1}(ind1, data{i,3} == time),'xr');
             plot(time,data{i,1}(ind2, data{i,3} == time),'xr');
         end
-        updPoints = find(data{i,5});
-        plot(data{i,3}(updPoints),data{i,1}(:,updPoints),'-xr');
+        if timesteps
+            updPoints = find(data{i,5});
+            plot(data{i,3}(updPoints),data{i,1}(:,updPoints),'-xr');
+            disp(length(updPoints) / max(data{i,3}));
+        end
         axis([0 max(data{i,3}) -2 7]);
-        disp(length(updPoints) / max(data{i,3}));
     end
 
 %    if animation

### 2013-10-06 src/1D/switch/displaySim.m

@@ -2,7 +2,7 @@ function displaySim(data, fps)
     movie = false;
     animation = false;
 
-    figure();
+    fi = figure();
     cntD = size(data,1);
     for i = 1:cntD
         i0 = i-1;
@@ -29,6 +29,8 @@ function displaySim(data, fps)
         end
         axis([0 max(data{i,3}) -2 7]);
     end
+
+    saveas(fi, '../../../images/spring_motivation.png');
 
 %    if animation
 %        figure();

### 2013-10-06 src/1D/switch/run.m

@@ -1,6 +1,6 @@
 clear all;
 close all;
-t_stop = 200;
+t_stop = 500;
 fpses = [10, 2, 1];
 results = cell(length(fpses),4);
 for i = 1:length(fpses)

### 2013-10-06 src/1D/switch/springSim.m

@@ -89,14 +89,14 @@ function [positions, lengths, times, switches] = springSim(fps_lowest, t_stop)
         l2 = springLength(pP, sp);
         mask = (l .* l2 < 0) & (l+l2 < 0);
         if any(mask)
-            dt = 1/30;
+%            dt = 1/30;
             disp('Switch in next iteration!');
             disp(time);
             pSub = unique(sp(mask,:));
             cnt = size(pSub,1);
             switches(end+1:end+cnt,:) = [repmat(time,[cnt 1]), pSub];
-            dt = dt/10;
-            pV(pSub) = 0;
+%            dt = dt/10;
+%            pV(pSub) = 0;
             [pF dx] = springForces(pP, sp, pV, spK, spX0, spC);
 
             [dp dv] = springStep(pP, pM, pV, sum(pF,2), dt);

### 2013-10-06 report/presentation_content.tex
Presentation of experimental plan done

@@
 \usetikzlibrary{shapes,arrows}
+\newcommand{\itemv}[1]{\item{#1}}
 
@@
 \begin{frame}{Short introduction to the project}
+    \pause\begin{itemize}[<+->]
+		\itemv Materials
         \begin{itemize}
-		\item Simulations of materials
-		\item Huge amount of computations for large materials
-        \item Reduce avg. number of calculations/time
+            \itemv In real life: Continuous
+            \itemv In simulations: Discrete representations using
+                particles/grids
+        \end{itemize}
+		\itemv Large/detailed materials = huge amount of particles/vertices
+        \itemv Varying grid size/scale
+        \itemv Simulation of time
+        \itemv Reduce avg. number of calculations/time
 	\end{itemize}
@@
         \center
-        \includegraphics[width=\linewidth,keepaspectratio]{../images/spring_motivation.png}
+        \includegraphics[width=\linewidth,height=6cm]{../images/spring_motivation.png}
         \caption{Simulations of mass spring particle system}
@@
 \begin{frame}{Overview of experiments}
-    \begin{itemize}
-        \item 1D experiments (simple)
-        \item 2D experiments (more complex)
-        \item Apply methods for uniform meshes 
-        \item Apply methods for multi-scale meshes
-        \item Compare method with regular timestepping
+    \pause\begin{itemize}[<+->]
+        \itemv 1D experiments (simple)
+        \itemv 2D experiments (more complex)
+        \itemv Use experiments to develop schemes
+        \itemv Apply methods for uniform meshes 
+        \itemv Apply methods for multi-scale meshes
+        \itemv Compare method with regular timestepping
     \end{itemize}
@@
 \begin{frame}{1D experiments}
-    \begin{itemize}
-        \item Application: Mass spring particle system
-        \item Hook's law
-        \item Simplicity for ease of development
+    \pause\begin{itemize}[<+->]
+        \itemv 1D for simplicity and ease of development
+        \itemv Develop a scheme from experimenting
+        \itemv Application: Mass spring particle system (clothes)
+        \itemv Hook's law
+        \itemv Uniform and multi-scale meshes and simulations
     \end{itemize}
@@
 \begin{frame}{2D experiments}
-    \begin{itemize}
-        \item Extend 1D scheme
-        \item Application: Hyperelastic materials
-        \item Stress and strain
+    \pause\begin{itemize}[<+->]
+        \itemv Extend 1D scheme to 2D
+        \itemv Gain in complexity
+        \itemv Application: Hyperelastic materials
+        \itemv Stress and strain
+        \itemv Uniform and multi-scale meshes and simulations
     \end{itemize}
@@
 \begin{frame}{Performance experiments}
-    \begin{itemize}
-        \item Total simulation time
+    \pause\begin{itemize}[<+->]
+        \itemv Total simulation time
+        \itemv Compare our 2D scheme with regular time stepping
+        \itemv Only sequential tests
+        \itemv Only in Matlab
     \end{itemize}

### 2013-10-07 src/1D/atomic/atomicSim.m
Atomic simulation without bugs

@@ -1,9 +1,7 @@
-function [positions, lengths, times, switches, active] = springSim(fps, t_stop)
+function [positions, lengths, times, switches, active] = springSim(fps, t_stop, stepBound, rFactor)
     %% Initialize particles
     % Create grid of particles at positions
     pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
-    rFactor = 2;
-    stepBound = 65;
 
     % Initialize variables
     cntP = size(pP,1);
@@ -50,14 +48,11 @@ function [positions, lengths, times, switches, active] = springSim(fps, t_stop)
         active(end+1) = length(spA);
 
         % Force calculation of active springs/particles
-        if (length(spA) > 0)
-            [pF dx] = springForces(pP, sp(spA,:), spL, pV,...
+        [pF dx] = springForces(pP, sp(spA,:), spL, pV,...
                                    spK(spA), spX0(spA), spC(spA));
-        end
 
         % Position and velocity update
         [dp dv] = springStep(pP, pM, pV, sum(pF,2), dt, pA, sp, spSteps);
-
         % Detect collisions / switches in position
         l = springLength(pP + dp, sp);
         l2 = springLength(pP, sp);

### 2013-10-07 src/1D/atomic/displaySim.m

@@ -12,13 +12,13 @@ function displaySim(data, fps)
         subplot(2, cntD, plot1);
         plot(data{i,3},data{i,2});
         title(['Spring lengths for ' num2str(fps(i)) ' fps']);
-        xlabel('Time');
+        xlabel('Time (s)');
         ylabel('Spring length');
         axis([0 max(data{i,3}) -0.3 0.55]);
         subplot(2, cntD, plot2);
         plot(data{i,3},data{i,1});
         title(['Simulation trace ' num2str(fps(i)) ' fps']);
-        xlabel('Time');
+        xlabel('Time (s)');
         ylabel('Position');
         hold on;
         for j = 1:size(data{i,4},1)

### 2013-10-07 src/1D/atomic/registerActive.m

@@ -11,20 +11,18 @@ lUpdate, velocities, time, dt, factor, steps, stepBound)
     l = springLength(positionsFull, springs);
 
     % Ratio between last updated length and next
-    ratio = lUpdate ./ l;
+    ratio = l ./ lUpdate;
 
     % Find indices of violating springs and points
-    sMask = ratio > factor | ratio < 1/factor;
+    sMask = ratio > factor | ratio <= 1/factor;
     if time == 0
         sMask = ones(size(springs,1),1);
-%    else
-%        sMask = zeros(size(springs,1),1);
     end
     if ~isnan(stepBound)
         sMask = sMask | steps >= stepBound;
     end
     % Temporary override to debug the rest of the simulator
-    sMask = find(sMask);
+    sMask = find(sMask > 0);
     pMask = unique(springs(sMask,:));
 
     % Update last updated lengths and number of steps since update

### 2013-10-07 src/1D/atomic/run.m

@@ -1,11 +1,18 @@
+% Clear environment
 clear all;
 close all;
+clc;
+
+% Set up simulation parameters
 t_stop = 500;
-fpses = [10,10];
+fpses = [10,2];
+stepBounds = [10, 8];
+rFactors = [2 2];
 results = cell(length(fpses),5);
+
+% Perform simulations
 for i = 1:length(fpses)
-    [results{i,1}, results{i,2}, results{i,3}, results{i,4}, results{i,5}] = atomicSim(fpses(i),t_stop);
-%    [results{i,1}, results{i,2}, results{i,3}] = springSim(fpses(i),t_stop);
+    [results{i,1}, results{i,2}, results{i,3}, results{i,4}, results{i,5}] = atomicSim(fpses(i),t_stop,stepBounds(i), rFactors(i));
 end
 
 displaySim(results, fpses);

### 2013-10-07 src/1D/atomic/springForces.m

@@ -1,5 +1,5 @@
 function [F, dx] = springForces(positions, springs, lUpdate, v, K, x0, C)
-    element = true;
+    element = false;
     % Initialize internal variables
     cntS = size(springs,1);
     cntP = size(positions,1);

### 2013-10-08 report/report.tex
Added notes/start of method and playing around with the simulator for new methods

@@
 \usepackage{amsmath,amssymb,amsfonts}
+
+\providecommand{\abs}[1]{\left \lvert #1 \right \rvert}
 
@@
 Simulations of large (order of millions of vertices) hyper elastic materials
-are very time consuming. When working with materials of varying density of
-vertices, the highest densities often make the most strict bounds on size of
-the time step of the simulation in order to keep the materials (and grids)
-stable. In a ``standard'' simulation with a global time step size we need
-to compute possibly unnecessarily many computations on the coarser parts of
-the material caused by this bound. When furthermore having materials with a
-large amount of somewhat equally distributed vertices and smaller patches of
-detailed areas with higher density of vertices, it would be an advantage to be
-able to perform large time steps for the majority of the material and smaller
-steps for the local patches when vertices cause instability. We call this
-concept local adaptive optimization of time step, which we will study further
-in this report.
+are very time consuming. When working with materials of varying density
+of vertices, the highest densities often make the most strict bounds on
+size of the time step of the simulation in order to keep the materials (and
+grids) stable and the discretization correct. In a ``standard'' simulation
+with a global time step size we need to compute possibly unnecessarily many
+computations on the coarser parts of the material caused by this bound.
+When furthermore having materials with a large amount of somewhat equally
+distributed vertices and smaller patches of detailed areas with higher density
+of vertices, it would be an advantage to be able to perform large time steps
+for the majority of the material and smaller steps for the local patches when
+vertices cause instability. We call this concept local adaptive optimization
+of time step, which we will study further in this report.
 
@@
 \section{Method}
+We will in this report consider the 1 dimensional problem of simulating a mass
+spring system, which is usually used for simulating cloth and hair etc. A mass
+spring system is a system consisting of a set of vertices $V$ each having a
+specified mass $m$, force $F$ and velocity $u$ and a set of springs $S$ with two
+endpoints $p_f \in V$ and $p_t \in V \setminus p_f$. Each spring has a damping
+factor $d$, stiffness constant $K$ and rest length $l_0$ describing the state of
+the spring.
 
-\begin{align}
-F_k &= -k \\
-F_m &= -F
-\end{align}
+Things to write about (notes):
+
+Damping force $F_d$:
+\begin{align*}
+    \Delta l &= l_{left} - l_{right} \\
+    F_d      &= - C_d v(i) =  - C_d \frac{\left(v_{left} - v_{right}\right)
+\Delta l}{\abs{\Delta l}}
+\end{align*}
+
+Spring force $F_s$:
+\begin{align*}
+    F_s = kx = -k \left( l - l_0 \right)
+\end{align*}
+
+Explicit Euler method:
+\begin{align*}
+    x(t+h) &= x(t) + h \frac{\partial x(t)}{\partial t}
+\end{align*}
+
+Midpoint method:
+\begin{align*}
+    x(t+h) = x(t) + h \left( f \left(x + \frac{h}{2} f \left(t \right) \right) \right)
+\end{align*}
 
 \section{Experiments and results}
+
+\subsection{Initial though}
+Normal timestepping with switch detection and refinement
+
+\subsection{Stepping with interpolation and lockout of re-calculations}
 

### 2013-10-08 src/1D/atomic/atomicSim.m

@@ -1,7 +1,8 @@
 function [positions, lengths, times, switches, active] = springSim(fps, t_stop, stepBound, rFactor)
     %% Initialize particles
     % Create grid of particles at positions
-    pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
+    %pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
+    pP = [-2:2:2]';
 
     % Initialize variables
     cntP = size(pP,1);
@@ -23,10 +24,10 @@ function [positions, lengths, times, switches, active] = springSim(fps, t_stop, 
     cntS = size(sp,1);
 
     % Set up initial measurements for each spring
-    spC = ones(cntS,1) * 0.5;   % Damping
+    spC = ones(cntS,1) * 0.01;   % Damping
     spK = ones(cntS,1) * 0.5;   % Stiffness
     spX0 = pP(sp);
-    spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.8;
+    spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.6;
     spD = ones(cntS,1) * 0.1;   % Viscous drag
     spL = springLength(pP,sp);  % Length at last update
     spSteps = zeros(cntS,1);    % Number of steps since last update
@@ -44,7 +45,8 @@ function [positions, lengths, times, switches, active] = springSim(fps, t_stop, 
 
         % Find active particles pA and active springs spA
         [pA spA spL spSteps_new] = registerActive(pP, sp, spL, pV, time, dt,...
-                                                 rFactor, spSteps, stepBound);
+                                                 rFactor, spSteps, stepBound,...
+                                                 spX0);
         active(end+1) = length(spA);
 
         % Force calculation of active springs/particles
@@ -62,7 +64,7 @@ function [positions, lengths, times, switches, active] = springSim(fps, t_stop, 
             disp(time);
             pSub = unique(sp(mask,:));
             cnt = size(pSub,1);
-            switches(end+1:end+cnt,:) = [repmat(time,[cnt 1]), pSub];
+%            switches(end+1:end+cnt,:) = [repmat(time,[cnt 1]), pSub];
         end
 
         % Registration

### 2013-10-08 src/1D/atomic/displaySim.m

@@ -1,7 +1,7 @@
 function displaySim(data, fps)
     movie = false;
     animation = false;
-    timesteps = false;
+    timesteps = true;
 
     figure();
     cntD = size(data,1);

### 2013-10-08 src/1D/atomic/registerActive.m

@@ -1,5 +1,5 @@
 function [pMask sMask lUpdate, steps] = registerActive(positions, springs,...
-lUpdate, velocities, time, dt, factor, steps, stepBound)
+lUpdate, velocities, time, dt, factor, steps, stepBound, rest)
     if nargin < 9
         stepBound = NaN;
     end
@@ -11,13 +11,13 @@ lUpdate, velocities, time, dt, factor, steps, stepBound)
     l = springLength(positionsFull, springs);
 
     % Ratio between last updated length and next
-    ratio = l ./ lUpdate;
+    ratio = abs(l-rest);
+    ratio2 = abs(l ./ lUpdate);
 
     % Find indices of violating springs and points
+    sMask = zeros(size(l));
     sMask = ratio > factor | ratio <= 1/factor;
-    if time == 0
-        sMask = ones(size(springs,1),1);
-    end
+%    sMask = ratio2 > factor | ratio2 <= 1/factor | sMask;
     if ~isnan(stepBound)
         sMask = sMask | steps >= stepBound;
     end

### 2013-10-08 src/1D/atomic/run.m

@@ -5,13 +5,14 @@ clc;
 
 % Set up simulation parameters
 t_stop = 500;
-fpses = [10,2];
-stepBounds = [10, 8];
-rFactors = [2 2];
+fpses =      [10, 10];
+stepBounds = [100, 100];
+rFactors =   [inf,  1.2];
 results = cell(length(fpses),5);
 
 % Perform simulations
 for i = 1:length(fpses)
+    disp(fpses(i));
     [results{i,1}, results{i,2}, results{i,3}, results{i,4}, results{i,5}] = atomicSim(fpses(i),t_stop,stepBounds(i), rFactors(i));
 end
 

### 2013-10-15 src/1D/inverse/diffSim.m
Inverse heuristic implemented

@@ -1,0 +1,18 @@
+function diffSim(data, fps)
+    cntD = size(data,1);
+
+    if cntD > 1
+        figure();
+        for i = 1:cntD-1
+            subplot(1,cntD-1,i);
+            mult = fps(i) / fps(i+1);
+            diff = data{i,2}(:,1:mult:end) - data{i+1,2};
+            plot(data{i,3}(1:mult:end),diff,'-');
+            title(['Diff between ' num2str(fps(i)) ' and ' num2str(fps(i+1)) ' fps']);
+            xlabel('Time');
+            ylabel('Diff in spring length');
+        end
+    else
+        error('Not able to calculate diff of only 1 simulation');
+    end
+end

### 2013-10-15 src/1D/inverse/inverseSim.m

@@ -1,0 +1,87 @@
+function [positions, lengths, times, switches, active] = inverseSim(fps, t_stop, stepBound, rFactor)
+    %% Initialize particles
+    % Create grid of particles at positions
+   % pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
+    pP = [-2:2:2]';
+
+    % Initialize variables
+    cntP = size(pP,1);
+    positions = [];
+    lengths = [];
+    times = [];
+    switches = [];
+    active = [];
+
+    % Set up initial measurements for each particle
+    pV = zeros(cntP,1);         % Initial velocities
+    pF = zeros(cntP,1);         % Force at time t
+    pM = ones(cntP,1) * 30;     % Mass of each particle
+
+    %% Register springs
+    % Springs between all inner particles (indices)
+    sp = [1:cntP]';
+    sp = [sp(1:end-1),sp(2:end)];
+    cntS = size(sp,1);
+
+    % Set up initial measurements for each spring
+    spC = ones(cntS,1) * 0.1;   % Damping
+    spK = ones(cntS,1) * 0.5;   % Stiffness
+    spX0 = pP(sp);
+    spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.52;
+    spD = ones(cntS,1) * 0.1;   % Viscous drag
+    spL = springLength(pP,sp);  % Length at last update
+    spSteps = zeros(cntS,1);    % Number of steps since last update
+
+    % Initial time
+    dt_base = 1 / fps;
+    dt = dt_base;
+    time = 0;
+
+    dx = zeros(cntS,1);
+    pF = zeros(cntP,2);
+
+    % Loop simulation over time
+    while time <= t_stop + dt/2;
+
+        pP = pP + pV*dt;
+        % Calculate spring timestep
+        [pA spA spL spSteps_new] = registerActive(pP, sp, spL, pV, time, dt,...
+                                                 rFactor, spSteps, stepBound,...
+                                                 spX0);
+        active(end+1) = length(spA);
+
+        % Force calculation of active springs/particles
+        [pF dx] = springForces(pP, sp(spA,:), spL, pV,...
+                                   spK(spA), spX0(spA), spC(spA));
+
+        % Position and velocity update
+        [dp dv] = springStep(pP, pM, pV, sum(pF,2), dt, pA, sp, spSteps);
+        % Detect collisions / switches in position
[…]

### 2013-10-15 src/1D/inverse/run.m

@@ -5,17 +5,17 @@ clc;
 
 % Set up simulation parameters
 t_stop = 500;
-fpses =      [10, 10, 30];
-stepBounds = [100, 100, 1];
-rFactors =   [inf,  0.2, inf];
+fpses =      [10,30];
+stepBounds = [1, 1];
+rFactors =   [inf, inf];
 results = cell(length(fpses),5);
 
 % Perform simulations
 for i = 1:length(fpses)
     disp(fpses(i));
-    [results{i,1}, results{i,2}, results{i,3}, results{i,4}, results{i,5}] = atomicSim(fpses(i),t_stop,stepBounds(i), rFactors(i));
+    [results{i,1}, results{i,2}, results{i,3}, results{i,4}, results{i,5}] = inverseSim(fpses(i),t_stop,stepBounds(i), rFactors(i));
 end
-
+results
 displaySim(results, fpses);
 
 %diffSim(results, fpses);

### 2013-10-15 src/1D/inverse/springTimestep.m

@@ -1,0 +1,13 @@
+function dt = springTimestep(positions, springs, m, K, x0)
+    % Get spring lengths
+    dt = ones(size(positions,1),1) * -Inf;
+    from = springs(:,1);
+    to   = springs(:,2);
+    dx = abs(springLength(positions, springs));
+    divider = abs(-K .* (dx - x0));
+
+    dt(from) = max(dt(from), divider);
+    dt(to) = max(dt(to), divider);
+    dt = (m ./ dt / 1000);
+    dt(dt > 1) = 1;
+end

### 2013-10-15 src/1D/stepbounding/atomicSim.m

@@ -1,0 +1,84 @@
+function [positions, lengths, times, switches, active] = springSim(fps, t_stop, stepBound, rFactor)
+    %% Initialize particles
+    % Create grid of particles at positions
+    %pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
+    pP = [-2:2:2]';
+
+    % Initialize variables
+    cntP = size(pP,1);
+    positions = [];
+    lengths = [];
+    times = [];
+    switches = [];
+    active = [];
+
+    % Set up initial measurements for each particle
+    pV = zeros(cntP,1);         % Initial velocities
+    pF = zeros(cntP,1);         % Force at time t
+    pM = ones(cntP,1) * 30;     % Mass of each particle
+
+    %% Register springs
+    % Springs between all inner particles (indices)
+    sp = [1:cntP]';
+    sp = [sp(1:end-1),sp(2:end)];
+    cntS = size(sp,1);
+
+    % Set up initial measurements for each spring
+    spC = ones(cntS,1) * 0.01;   % Damping
+    spK = ones(cntS,1) * 0.5;   % Stiffness
+    spX0 = pP(sp);
+    spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.6;
+    spD = ones(cntS,1) * 0.1;   % Viscous drag
+    spL = springLength(pP,sp);  % Length at last update
+    spSteps = zeros(cntS,1);    % Number of steps since last update
+
+    % Initial time
+    dt_base = 1 / fps;
+    dt = dt_base;
+    time = 0;
+
+    dx = zeros(cntS,1);
+    pF = zeros(cntP,2);
+
+    % Loop simulation over time
+    while time <= t_stop + dt/2;
+
+        % Find active particles pA and active springs spA
+        [pA spA spL spSteps_new] = registerActive(pP, sp, spL, pV, time, dt,...
+                                                 rFactor, spSteps, stepBound,...
+                                                 spX0);
+        active(end+1) = length(spA);
+
+        % Force calculation of active springs/particles
+        [pF dx] = springForces(pP, sp(spA,:), spL, pV,...
+                                   spK(spA), spX0(spA), spC(spA));
+
+        % Position and velocity update
+        [dp dv] = springStep(pP, pM, pV, sum(pF,2), dt, pA, sp, spSteps);
+        % Detect collisions / switches in position
+        l = springLength(pP + dp, sp);
[…]

### 2013-10-15 src/1D/stepbounding/springForces.m

@@ -1,0 +1,49 @@
+function [F, dx] = springForces(positions, springs, lUpdate, v, K, x0, C)
+    element = false;
+    % Initialize internal variables
+    cntS = size(springs,1);
+    cntP = size(positions,1);
+    Fs  = zeros(cntS,1);
+
+    % Initialize output 
+    F   = zeros(cntP,2);
+    dx = zeros(cntS,1);
+
+    % Calculate force F of each spring
+    if element
+        for s = 1:cntS
+            % Endpoints of spring
+            from = springs(s,1);
+            to = springs(s,2);
+
+            % Lengh of spring
+            dx(s) = springLength(positions, springs(s,:));
+            l = abs(dx(s));
+
+            % Force calculation
+            Fs = K(s) * (l - x0(s));                        % Spring force
+            Fs = Fs + C(s) * (v(from) - v(to)) * dx(s) / l; % Dampening
+            Fs = Fs * (- dx(s) / l);                        % Direction
+
+            % Apply forces to endpoints
+            F(from,1) = F(from,1) + Fs;
+            F(to,2) = F(to,2) - Fs;
+        end
+    else
+        % Matrix alternative to spring (element) wise computation
+        from = springs(:,1);
+        to   = springs(:,2);
+
+        dx = springLength(positions, springs);
+        l  = abs(dx);
+
+        Fs = K .* (l - x0);
+        Fs = Fs + C .* (v(from) - v(to)) .* dx ./ l;
+        Fs = Fs .* (-dx ./ l);
+
+        F(from,1) = F(from,1) + Fs;
+        F(to,2)   = F(to,2) - Fs;
+
+    end
+
+end

### 2013-10-15 src/1D/stepbounding/springLength.m

@@ -1,0 +1,8 @@
+function [l] = springLength(positions, springs)
+    % Get indices
+    left  = positions(springs(:,1));
+    right = positions(springs(:,2));
+
+    % Compute length (left - right)
+    l = left - right;
+end

### 2013-10-15 src/1D/switch/run.m

@@ -1,7 +1,7 @@
 clear all;
 close all;
 t_stop = 500;
-fpses = [10, 2, 1];
+fpses = [10, 2, 0.15];
 results = cell(length(fpses),4);
 for i = 1:length(fpses)
     [results{i,1}, results{i,2}, results{i,3}, results{i,4}] = springSim(fpses(i),t_stop);

### 2013-10-15 src/1D/switch/springSim.m

@@ -29,7 +29,7 @@ function [positions, lengths, times, switches] = springSim(fps_lowest, t_stop)
     spC = ones(cntS,1) * 0.5;   % Damping
     spK = ones(cntS,1) * 0.5;   % Stiffness
     spX0 = pP(sp);
-    spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.8;
+    spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.79;
     spD = ones(cntS,1) * 0.1;   % Viscous drag
 
     distance_show = false;
@@ -70,6 +70,7 @@ function [positions, lengths, times, switches] = springSim(fps_lowest, t_stop)
     dx = zeros(cntS,1);
     pF = zeros(cntP,2);
     while time <= t_stop + 1e-5;
+        pP = pP + pV * dt;
 
         % Check for subdivision
         if dt < dt_base && length(pSub) > 0
@@ -108,7 +109,6 @@ function [positions, lengths, times, switches] = springSim(fps_lowest, t_stop)
         times(end+1) = time;
 
         % Subdivide
-        pP = pP + dp;
         pV = pV + dv;
 
         % Time update

### 2013-10-16 report/article.tex
Moved report into article.tex to get two column layout

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Journal Article
% LaTeX Template
% Version 1.3 (9/9/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[twoside]{article}

\usepackage{lipsum} % Package to generate dummy text throughout this template

\usepackage[sc]{mathpazo} % Use the Palatino font
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\linespread{1.05} % Line spacing - Palatino needs more space between lines
\usepackage{microtype} % Slightly tweak font spacing for aesthetics

\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
\usepackage{multicol} % Used for the two-column layout of the document
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{float} % Required for tables and figures in the multi-column environment - they need to be placed in specific locations with the [H] (e.g. \begin{table}[H])
\usepackage[hidelinks]{hyperref} % For hyperlinks in the PDF

\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text
\usepackage{paralist} % Used for the compactitem environment which makes bullet points with less space between them

\usepackage{amsmath,amssymb,amsfonts}
\providecommand{\abs}[1]{\left \lvert #1 \right \rvert}

\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text

\usepackage{titlesec} % Allows customization of titles
\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
\renewcommand\thesubsection{\Roman{subsection}} % Roman numerals for subsections
\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles

\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
%\fancyhead[C]{Running title $\bullet$ November 2012 $\bullet$ Vol. XXI, No. 1} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Local Adaptive
Optimization of Time step}} % Article title

\author{
\large
\textsc{Malte Stær Nissen}\thanks{Thank to my supervisor Sune Darkner for
helping me realize this project}\\[2mm] % Your name
\normalsize University of Copenhagen \\ % Your institution
\normalsize \href{mailto:nissen@diku.dk}{nissen@diku.dk} % Your email address
\vspace{-5mm}
}
\date{}

%----------------------------------------------------------------------------------------

\begin{document}

\maketitle % Insert title

\thispagestyle{fancy} % All pages have headers and footers

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

\begin{abstract}

\end{abstract}

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\begin{multicols}{2} % Two-column layout throughout the main article text

\section{Introduction}
\lettrine[nindent=0em,lines=3]{S} imulations of large hyper elastic materials
are very time consuming. When working with materials of varying density
of vertices/scale of mesh, the highest densities often make the most
strict bounds on size of the time step of the simulation in order to keep
the materials (and grids) stable and the discretization correct. in a
``standard'' simulation with a global time step size we need to compute
possibly unnecessarily many computations on the coarser parts of the material
caused by this bound. When furthermore having materials with a large amount
of somewhat equally distributed vertices and smaller patches of detailed
areas with higher density of vertices, it would be an advantage to be able
to perform large time steps for the majority of the material and smaller
steps for the local patches where vertices could cause instability. we call
the concept of varying the timestep local adaptive optimization of time step
(LAOT), in which local either refer to spatial locality or time locality. We
will in this report study LAOT in the simple 1 dimensional simulation of a
mass spring system in order to gain knowledge into the field for further use
in simulation of hyper elastic materials. We will first describe the simulation
problem and then go into the study of the application of LOAT herein.

%------------------------------------------------

\section{Previous work}
As according to \cite{Gander:2013} local time stepping was first studied
in the community of ordinary differential equations (ODE) with Rice
\cite{rice:1960} developing the split Runge-Kutta methods (multi rate
Runge-Kutta methods). The concept behind these methods is the splitting of
the ODEs into multiple (two components in \cite{rice:1960}) components of
which each component needs to be integrated in different scales. A strategy
is chosen as proposed in \cite{Kvaernoe:1999} of either computing the coarser
part first (``Fastest first strategy'') followed by the finer part or vice
versa (``Slowest first strategy''). In order to perform these sequential
computations either interpolation or extrapolation of the first computation
is performed to compute the second part in each time step and the different
time step sizes can be adapted to fit the model in each step as well. See
\cite{Kvaernoe:1999} and \cite{Gear:1984} (similar strategy for linear multi
step methods) for more detailed descriptions of the two strategies.

In the partial differential equations (PDE) community, the adaptive time
stepping area was explored and developed with the goal of simulating
very specific known problems such as the (hyperbolic) wave equations
and (parabolic) heat equations instead of more general applications. We
will adhere from describing the methods further in this report, but a
short description and comparison of the specific methods can be found in
\cite{Gander:2013}. Generally the methods developed at first all make use
of the same basic concepts for making the local adaptive time stepping:
Interpolation, extrapolation, prediction and correction, which we are going to
use in the work of this report as well.

Recently newer and faster methods with local time stepping have been developed
on the basis of nonlinear PDEs and the work performed in this field in
the 80s and 90s. First the multi resolution (MR) schemes for creating
space-adaptive discretizations and refinements of these were developed, see
\cite{Berger:1984}. Then multiple MR based schemes were developed. Domingues
et al. \cite{Domingues:2008} is one of the more recent of these MR schemes.
They describe a local scale-dependent time stepping for a space-adaptive multi
resolution scheme using the finite volume method in order to obtain speed-up
using larger time-steps without violating a defined stability constraints,
which is essentially the same motivation as this report. The method is based
on an explicit Runge-Kutta method of second order. As expected the time step
size is imposed by a stability condition of the explicit Runge-Kutta on the
finest scale, which increases with the scale of the mesh and hence we are able
to increase the time step as well without violating the stability condition.

%------------------------------------------------

\section{Methods}

We will in this report consider the 1 dimensional problem of simulating a mass
spring system, which is usually (in 3 dimensions) used for simulating
cloth and hair etc. A mass spring system is a system consisting of a set of
vertices $V$ each having a specified mass $m$, force $F$ and velocity $u$ and
a set of springs $S$ with two endpoints $p_f \in V$ and $p_t \in V \setminus
p_f$. Each spring $s$ has a damping factor $d_s$, a stiffness constant $K_s$ and a rest
length $l_{s0}$ describing the spring's physical properties.

Things to write about (notes):

Damping force $F_d$:
\begin{align}
    \Delta l &= l_{left} - l_{right} \\
    F_d      &= - C_d v(i) =  - C_d \frac{\left(v_{left} - v_{right}\right)
\Delta l}{\abs{\Delta l}}
\end{align}

Spring force $F_s$:
\begin{align}
    F_s = -kx = -k \left( l - l_0 \right)
\end{align}

Explicit Euler method:
\begin{align}
    x(t+h) &= x(t) + h \frac{\partial x(t)}{\partial t}
\end{align}

Midpoint method:
\begin{align}
    x(t+h) = x(t) + h \left( f \left(x + \frac{h}{2} f \left(t \right) \right) \right)
\end{align}

\cite{Keshav:2007}
\subsection{Switch}
Change timestep when detecting collision/switch in next iteration

\subsection{Spatial LAOT}
Atomic update

Fixed update rate and detection of need for re-calculation of individual springs

Heuristic: Length of spring, steps since last update

\subsection{Inverse}
Time-local optimization according to heuristic:

\begin{align}
    \frac{1}{a} &= \frac{1}{\frac{-kx}{m}} = \frac{m}{-kx} \\
    \Delta t &= \abs{\frac{1}{a}} + \frac{1}{0.1}
\end{align}

%------------------------------------------------

\section{Results}

%\begin{table}[H]
%\caption{Example table}
%\centering
%\begin{tabular}{llr}
%\toprule
%\multicolumn{2}{c}{Name} \\
%\cmidrule(r){1-2}
%First name & Last Name & Grade \\
%\midrule
%John & Doe & $7.5$ \\
%Richard & Miles & $2$ \\
%\bottomrule
%\end{tabular}
%\end{table}

%------------------------------------------------

\section{Discussion}

Adaptive timestepping based on highest acceleration/smallest grid

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------

\bibliography{bibliography}
\bibliographystyle{plain}

%----------------------------------------------------------------------------------------

\end{multicols}

\end{document}

### 2013-10-22 report/contract.tex
Litterature study handin and papers
Cleanup of code and report

--- 2013-09-05 report/report.tex
@@
     \item Implement local adaptive optimization of timesteps (LAOT)
-	\item Evaluate LAOT to minimize the risk of unwanted deformations in grids.
+	\item Evaluate LAOT to minimize the risk of unwanted deformations in grids
+        and minimize the computational cost.
     \item Discuss how LAOT is performed in practice using existing 2nd order optimization methods.
-    \item Study of convergence on implemented methods.
 \end{enumerate}
@@
 \subsection{Problem Statement}
-Study the application of locally adaptive timestepping in simulation of hyperelastic materials.
+Study the application of locally adaptive timestepping in 1 dimensional
+simulation of a mass spring system.
 
@@
 smaller areas might need smaller timesteps. These situations are the ones
-we wish to accomodate in this paper.
+we wish to accomodate in this paper by gaining insight into the area of LAOT in
+a simple 1 dimensional simulation (mass spring system).
 
@@
 \subsection{Risk assessment}
-Low to none. The technique has been performed before in various shapes and it
-should therefore be do-able.
+Medium. The technique of LAOT has been performed before in various shapes but
+not the same as our proposed solution.
 

### 2013-10-22 report/report.tex
Cleanup of code and report

@@
 \section{Introduction}
-Simulations of large (order of millions of vertices) hyper elastic materials
-are very time consuming. When working with materials of varying density
+simulations of large (order of millions of vertices) hyper elastic materials
+are very time consuming. when working with materials of varying density
 of vertices, the highest densities often make the most strict bounds on
 size of the time step of the simulation in order to keep the materials (and
-grids) stable and the discretization correct. In a ``standard'' simulation
+grids) stable and the discretization correct. in a ``standard'' simulation
 with a global time step size we need to compute possibly unnecessarily many
 computations on the coarser parts of the material caused by this bound.
-When furthermore having materials with a large amount of somewhat equally
+when furthermore having materials with a large amount of somewhat equally
 distributed vertices and smaller patches of detailed areas with higher density
@@
 for the majority of the material and smaller steps for the local patches when
-vertices cause instability. We call this concept local adaptive optimization
+vertices cause instability. we call this concept local adaptive optimization
 of time step, which we will study further in this report.
@@
 
-\subsection{Initial though}
-Normal timestepping with switch detection and refinement
+\subsection{Switch}
 
-\subsection{Stepping with interpolation and lockout of re-calculations}
+\subsection{Spatial LAOT}
+
+\subsection{Inverse}
 
 \section{Discussion}
+
+Adaptive timestepping based on highest acceleration/smallest grid
 

### 2013-10-22 src/1D/atomic/atomicSim.m

@@ -28,7 +28,6 @@ function [positions, lengths, times, switches, active] = springSim(fps, t_stop, 
     spK = ones(cntS,1) * 0.5;   % Stiffness
     spX0 = pP(sp);
     spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.6;
-    spD = ones(cntS,1) * 0.1;   % Viscous drag
     spL = springLength(pP,sp);  % Length at last update
     spSteps = zeros(cntS,1);    % Number of steps since last update
 
@@ -50,7 +49,7 @@ function [positions, lengths, times, switches, active] = springSim(fps, t_stop, 
         active(end+1) = length(spA);
 
         % Force calculation of active springs/particles
-        [pF dx] = springForces(pP, sp(spA,:), spL, pV,...
+        [pF dx] = springForces(pP, sp(spA,:), pV,...
                                    spK(spA), spX0(spA), spC(spA));
 
         % Position and velocity update
@@ -58,7 +57,7 @@ function [positions, lengths, times, switches, active] = springSim(fps, t_stop, 
         % Detect collisions / switches in position
         l = springLength(pP + dp, sp);
         l2 = springLength(pP, sp);
-        mask = (l .* l2 < 0) & (l+l2 < 0);
+        mask = l .* l2 < 0;
         if any(mask)
             disp('Switch in next iteration!');
             disp(time);

### 2013-10-22 src/1D/displaySim.m

@@ -1,4 +1,4 @@
-function displaySim(data, fps)
+function displaySim(data, fps, settings)
     movie = false;
     animation = false;
     timesteps = true;

### 2013-10-22 src/1D/initSim.m

@@ -1,0 +1,35 @@
+function simConst = initSim(simulation)
+    switch simulation
+        case 'simple'
+            simConst = struct(...
+                        'damping',   0.01,...
+                        'stiffness', 0.5,...
+                        'spX0',      0.6,...
+                        'mass',      30,...
+                        'particles', [-2:2:2]'...
+                       );
+        case 'inverse'
+            simConst = struct(...
+                        'damping',   0.1,...
+                        'stiffness', 0.5,...
+                        'spX0',      0.52,...
+                        'mass',      30,...
+                        'particles', [-2:2:2]'...
+                       );
+        case 'multiscale'
+            simConst = struct(...
+                        'damping',   0.01,...
+                        'stiffness', 0.5,...
+                        'spX0',      0.6,...
+                        'mass',      30,...
+                        'particles', [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]'...
+                       );
+        otherwise
+            simConst = struct(...
+                        'damping',   0.5,...
+                        'stiffness', 0.5,...
+                        'spX0',      0.785,...
+                        'mass',      30,...
+                        'particles', [-2:0.5:1.99, 2:0.15:2.99,3:0.5:6.99]'...
+                       );
+end

### 2013-10-22 src/1D/inverse/springForces.m

@@ -1,4 +1,4 @@
-function [F, dx] = springForces(positions, springs, lUpdate, v, K, x0, C)
+function [F, dx] = springForces(positions, springs, v, K, x0, C)
     element = false;
     % Initialize internal variables
     cntS = size(springs,1);
@@ -41,6 +41,7 @@ function [F, dx] = springForces(positions, springs, lUpdate, v, K, x0, C)
         Fs = Fs + C .* (v(from) - v(to)) .* dx ./ l;
         Fs = Fs .* (-dx ./ l);
 
+        % Apply forces to endpoints
         F(from,1) = F(from,1) + Fs;
         F(to,2)   = F(to,2) - Fs;
 

### 2013-10-22 src/1D/run.m

@@ -1,0 +1,49 @@
+% Clear environment
+clear all;
+close all;
+clc;
+
+% Set up simulation parameters
+t_stop = 500;
+
+
+%% Atomic:
+fpses =      [10, 10, 30];
+stepBounds = [100, 100, 1];
+rFactors =   [inf,  0.2, inf];
+results = cell(length(fpses),5);
+
+% Perform simulations
+for i = 1:length(fpses)
+    disp(fpses(i));
+    [results{i,1}, results{i,2}, results{i,3}, results{i,4}, results{i,5}] = atomicSim(fpses(i),t_stop,stepBounds(i), rFactors(i));
+end
+
+displaySim(results, fpses);
+
+%% Inverse
+fpses =      [10,30];
+stepBounds = [1, 1];
+rFactors =   [inf, inf];
+results = cell(length(fpses),5);
+
+% Perform simulations
+for i = 1:length(fpses)
+    disp(fpses(i));
+    [results{i,1}, results{i,2}, results{i,3}, results{i,4}, results{i,5}] = inverseSim(fpses(i),t_stop,stepBounds(i), rFactors(i));
+end
+results
+displaySim(results, fpses);
+
+
+displaySim(results, fpses);
+
+%% Switch
+fpses = [10, 2, 0.15, 30];
+results = cell(length(fpses),4);
+for i = 1:length(fpses)
+    disp(['Simulation: ' num2str(fpses(i))]);
+    [results{i,1}, results{i,2}, results{i,3}, results{i,4}] = springSim(fpses(i),t_stop);
+end
+
+displaySim(results, fpses);

### 2013-10-22 src/1D/springForces.m

@@ -1,4 +1,4 @@
-function [F, dx] = springForces(positions, springs, lUpdate, v, K, x0, C)
+function [F, dx] = springForces(positions, springs, v, K, x0, C)
     element = false;
     % Initialize internal variables
     cntS = size(springs,1);
@@ -16,7 +16,7 @@ function [F, dx] = springForces(positions, springs, lUpdate, v, K, x0, C)
             from = springs(s,1);
             to = springs(s,2);
 
-            % Lengh of spring
+            % Lenght of spring
             dx(s) = springLength(positions, springs(s,:));
             l = abs(dx(s));
 
@@ -34,13 +34,15 @@ function [F, dx] = springForces(positions, springs, lUpdate, v, K, x0, C)
         from = springs(:,1);
         to   = springs(:,2);
 
+        % Length of spring
         dx = springLength(positions, springs);
         l  = abs(dx);
 
-        Fs = K .* (l - x0);
-        Fs = Fs + C .* (v(from) - v(to)) .* dx ./ l;
-        Fs = Fs .* (-dx ./ l);
+        Fs = K .* (l - x0);                             % Spring force
+        Fs = Fs + C .* (v(from) - v(to)) .* dx ./ l;    % Dampening
+        Fs = Fs .* (-dx ./ l);                          % Direction
 
+        % Apply forces to endpoints
         F(from,1) = F(from,1) + Fs;
         F(to,2)   = F(to,2) - Fs;
 

### 2013-10-22 src/1D/switch/displaySim.m

@@ -7,7 +7,7 @@ function displaySim(data, fps)
     for i = 1:cntD
         i0 = i-1;
         plot1 = i;
-        plot2 = plot1+3;
+        plot2 = plot1+size(data,1);
         subplot(2, cntD, plot1);
         plot(data{i,3},data{i,2});
         title(['Spring lengths for ' num2str(fps(i)) ' fps']);

### 2013-10-22 src/1D/switch/run.m

@@ -1,9 +1,14 @@
+% Clear environment
 clear all;
 close all;
+clc;
+
+% Set up simulation parameters
 t_stop = 500;
-fpses = [10, 2, 0.15];
+fpses = [10, 2, 0.15, 30];
 results = cell(length(fpses),4);
 for i = 1:length(fpses)
+    disp(['Simulation: ' num2str(fpses(i))]);
     [results{i,1}, results{i,2}, results{i,3}, results{i,4}] = springSim(fpses(i),t_stop);
 end
 

### 2013-10-22 src/1D/switch/springForces.m

@@ -1,4 +1,5 @@
 function [F, dx] = springForces(positions, springs, v, K, x0, C)
+    element = false;
     % Initialize internal variables
     cntS = size(springs,1);
     cntP = size(positions,1);
@@ -9,23 +10,41 @@ function [F, dx] = springForces(positions, springs, v, K, x0, C)
     dx = zeros(cntS,1);
 
     % Calculate force F of each spring
-    for s = 1:cntS
-        % Endpoints of spring
-        from = springs(s,1);
-        to = springs(s,2);
+    if element
+        for s = 1:cntS
+            % Endpoints of spring
+            from = springs(s,1);
+            to = springs(s,2);
 
-        % Lengh of spring
-        dx(s) = springLength(positions, springs(s,:));
-        l = abs(dx(s));
+            % Lengh of spring
+            dx(s) = springLength(positions, springs(s,:));
+            l = abs(dx(s));
 
-        % Force calculation
-        Fs = K(s) * (l - x0(s));                        % Spring force
-        Fs = Fs + C(s) * (v(from) - v(to)) * dx(s) / l; % Dampening
-        Fs = Fs * (- dx(s) / l);                        % Direction
+            % Force calculation
+            Fs = K(s) * (l - x0(s));                        % Spring force
+            Fs = Fs + C(s) * (v(from) - v(to)) * dx(s) / l; % Dampening
+            Fs = Fs * (- dx(s) / l);                        % Direction
+
+            % Apply forces to endpoints
+            F(from,1) = F(from,1) + Fs;
+            F(to,2) = F(to,2) - Fs;
+        end
+    else
+        % Matrix alternative to spring (element) wise computation
+        from = springs(:,1);
+        to   = springs(:,2);
+
+        dx = springLength(positions, springs);
+        l  = abs(dx);
+
+        Fs = K .* (l - x0);
+        Fs = Fs + C .* (v(from) - v(to)) .* dx ./ l;
+        Fs = Fs .* (-dx ./ l);
 
         % Apply forces to endpoints
         F(from,1) = F(from,1) + Fs;
-        F(to,2) = F(to,2) - Fs;
+        F(to,2)   = F(to,2) - Fs;
+
     end
 
 end

### 2013-10-22 report/article.tex
Introduction and Previous Work shipped for review. Mass spring system initiated

@@
 
+\usepackage{todonotes}
 \usepackage{amsmath,amssymb,amsfonts}
@@
 \title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Local Adaptive
-Optimization of Time step}} % Article title
+Optimization of Time Step}} % Article title
 
@@
 \lettrine[nindent=0em,lines=3]{S} imulations of large hyper elastic materials
-are very time consuming. When working with materials of varying density
+can be very time consuming. When working with materials of varying density
 of vertices/scale of mesh, the highest densities often make the most
@@
 areas with higher density of vertices, it would be an advantage to be able
-to perform large time steps for the majority of the material and smaller
-steps for the local patches where vertices could cause instability. we call
-the concept of varying the timestep local adaptive optimization of time step
-(LAOT), in which local either refer to spatial locality or time locality. We
-will in this report study LAOT in the simple 1 dimensional simulation of a
-mass spring system in order to gain knowledge into the field for further use
-in simulation of hyper elastic materials. We will first describe the simulation
-problem and then go into the study of the application of LOAT herein.
+to perform large time steps for the majority of the material (the coarse
+parts) and smaller steps for the local patches where vertices could cause
+instability. we call the concept of varying the timestep local adaptive
+optimization of time step (LAOTS), in which ``local'' either refers to spatial
+locality (as described above) or time locality where the time step is global
+for all vertices but varies during the simulation. We will in this report
+study possible approaches for application of LAOTS. In order to keep the focus
+on possible LAOTS methods and remove unneccessary complexities we will study
+the simple 1 dimensional simulation of a mass spring system and try to apply
+LAOTS for it. Afterwards we will shortly discuss the application of our gained
+knowledge to the domain of hyper elastic material simulation and simulations
+in general.
 
@@
 
-We will in this report consider the 1 dimensional problem of simulating a mass
-spring system, which is usually (in 3 dimensions) used for simulating
-cloth and hair etc. A mass spring system is a system consisting of a set of
-vertices $V$ each having a specified mass $m$, force $F$ and velocity $u$ and
-a set of springs $S$ with two endpoints $p_f \in V$ and $p_t \in V \setminus
-p_f$. Each spring $s$ has a damping factor $d_s$, a stiffness constant $K_s$ and a rest
-length $l_{s0}$ describing the spring's physical properties.
+With no prerequisites within the field of mass simulations we will first
+consider the 1D problem of simulating a mass spring system using the explicit
+(forward) Euler method.
+
+Second we will propose and test out three different methods for applying LAOTS
+to the simulation.
+
+Third and last we will discussion the results and how we
+could apply our knowledge of LAOTS for hyper elastic material.
+
+\subsection{1D mass spring system}
+Mass spring systems are systems simulating a set of vertices interconnected by
+springs in various ways to obtain desirable physical properties. The motion of
+the springs and vertices are then simulated as a function of the time elapsed.
+Examples of application of mass spring systems are for simulation of cloth or
+hair.
+
+Our mass spring system consists of a set $V$ of vertices and a set $S$ of
+springs. Each vertex $v \in V$ has a specified position $p_v$, mass $m_v$, force
+$F_v$
+and velocity $u_v$ and each spring $s$ has two endpoints vertices $p_{sf} \in V$
+and $p_{st} \in V \setminus p_{sf}$, a damping factor $d_s$, a stiffness
+constant $K_s$ and a rest length $l_{s0}$. Combined all these variables and
+constants describe the state of the mass spring system at each timestep during a
+simulation.
+
+As each vertex $v$ is connected to 1 or 2 springs, its motion can be calculated by
+the total force $F_v$ applied by the springs it is connected to. The spring
+force is calculated using Hooke's Law and adding damping [TODO: Add reference (wiki):
+    \url{http://en.wikipedia.org/wiki/Harmonic_oscillator}]:
+\begin{align}
+    \label{eq:hooke}
+    F_{h,d} &= -k_d x_d = -k_d (l_d - l_{0,d}) \\
+    \label{eq:damping}
+    F_{C,d} &= -C_d \frac{dx_d}{dt} = - C_d \frac{\left(v_{left} - v_{right}\right)
+\Delta l}{\abs{\Delta l}}
+\end{align}
+in which $F_h$ is the spring force of Hooke's Law, $F_c$ is the damping force,
+$k$ is the spring constant, $x$ is the difference between the current length of
+the spring and the spring rest length, $c$ is the damping constant and
+$\frac{dx}{dt}$ is the velocity of the vertex $v$.
 
@@
 
-\subsection{Spatial LAOT}
+\subsection{Spatial LAOTS}
 Atomic update

### 2013-10-22 review/review.tex
Initiated review

\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage{lmodern}

\usepackage{amsmath,amssymb,amsfonts}

\title{Review of Human Pose Tracking}
\author{Malte Stær Nissen}

\begin{document}
\maketitle

\noindent Corrections/Questions:
\begin{itemize}
    \item 1.L2: Comma before ``and video games'' (Oxford comma)
    \item 1.L13 and 2.4.1.L2: Gradient ascent/descent - which is it? Looks as if you refer to
        the same paper and the same part of the algorithm.
    \item Figure 1: Is the figure your own product or from somewhere else?
    \item 2.1.L3, 2.1.L11: Shouldn't you always start with a capital letter after colon?
    \item 2.3.2.L2: ``requires a shape model, described above'' (where exactly?)
    \item 2.3.3: Any reference? When you have chopped the section into
        subsections it might be a good idea to have references for all
        methods
    \item 2.6.L3: ``the then state-of-the-art method''. The ``then''-part Doesn't sound correct
        in my head/mind. I have however no alternative solution at hand.
\end{itemize}

\noindent General:
\begin{itemize}
    \item Introduction: Good and sufficient describtion to get a grasp of the
        problem and the work being performed (exclusing results).
    \item When reading the litterature study (all the subsections following the
        introduction) it first seems quite messy with a
    lot of terminology but no red line. After a few times of reading the
    subsections and introduction (2.1)
    it gets more obvious what your intention/red line of the litterature study
    is. Maybe emphasize in 2.1.L7 that the overview is the other subsections of
    section 2. Would've helped me a bit when first reading the report.

    \item Quite long litterature study explaining a lot of the therminology
        within the field. Maybe some of it is more suitable in the method
        section?
    \item Consistency in writing style (sections): Some places (2.4) you describe (name) the
        following subsections in the overall section. Other places (2.3.2, 2.3,
        2.5) you only write that there are X (number) of methods/subsections,
        not what they are.
    \item Maybe some sort of conclusion of the litterature study could be added
        at the end? Some comments on what you're going to implement/use of all
        the stuff you present.
\end{itemize}

\end{document}

### 2013-10-24 review/review.tex
Review done

@@
     \item Figure 1: Is the figure your own product or from somewhere else?
-    \item 2.1.L3, 2.1.L11: Shouldn't you always start with a capital letter after colon?
+    \item Figure 1: Write which of the figures is which (good = left, bad =
+        right) - it might be obvious but it's always good to be on the safe
+        side.
+    \item 1.L1, 2.1.L3, 2.1.L11: Shouldn't you always start with a capital letter after colon?
+    \item 2.1: Why make a subsection called Introduction. Why not just write the
+        entire part directly after section 2 starts?
     \item 2.3.2.L2: ``requires a shape model, described above'' (where exactly?)
@@
         subsections it might be a good idea to have references for all
-        methods
+        methods/parts
     \item 2.6.L3: ``the then state-of-the-art method''. The ``then''-part Doesn't sound correct
@@
     section 2. Would've helped me a bit when first reading the report.
-
-    \item Quite long litterature study explaining a lot of the therminology
-        within the field. Maybe some of it is more suitable in the method
-        section?
-    \item Consistency in writing style (sections): Some places (2.4) you describe (name) the
-        following subsections in the overall section. Other places (2.3.2, 2.3,
-        2.5) you only write that there are X (number) of methods/subsections,
-        not what they are.
+    \item Litterature study level of detail good and sufficient.
+    \item Consistency in writing style (sections/subsections): Some places
+    (2.4) you describe (name) the following subsections in the overall
+    section. Other places (2.3.2, 2.3, 2.5) you only write that there are X
+    (number) of methods/subsections, not what they are.
     \item Maybe some sort of conclusion of the litterature study could be added
-        at the end? Some comments on what you're going to implement/use of all
-        the stuff you present.
+        at the end? Some comments on what exactly you're going to implement/use of all
+        the stuff you present. I do know that you've described what you're going
+        to implement earlier but a nice round-up of exact references might be an
+        idea?
 \end{itemize}

### 2013-10-24 report/article.tex
1D mass-spring simulation described and minor grammatical corrections

@@
 \large
-\textsc{Malte Stær Nissen}\thanks{Thank to my supervisor Sune Darkner for
+\textsc{Malte Stær Nissen}\thanks{Thanks to my supervisor Sune Darkner for
 helping me realize this project}\\[2mm] % Your name
@@
 
-\begin{abstract}
+%\begin{abstract}
 
-\end{abstract}
+%\end{abstract}
 
@@
 \section{Introduction}
-\lettrine[nindent=0em,lines=3]{S} imulations of large hyper elastic materials
+\lettrine[nindent=0em,lines=3]{S} imulations of large hyperelastic materials
 can be very time consuming. When working with materials of varying density
-of vertices/scale of mesh, the highest densities often make the most
+of vertices/scale of mesh, the highest densities often give the most
 strict bounds on size of the time step of the simulation in order to keep
-the materials (and grids) stable and the discretization correct. in a
+the materials (and grids) stable and the discretization correct. In a
 ``standard'' simulation with a global time step size we need to compute
@@
 parts) and smaller steps for the local patches where vertices could cause
-instability. we call the concept of varying the timestep local adaptive
+instability. We call the concept of varying the timestep local adaptive
 optimization of time step (LAOTS), in which ``local'' either refers to spatial
@@
 study possible approaches for application of LAOTS. In order to keep the focus
-on possible LAOTS methods and remove unneccessary complexities we will study
-the simple 1 dimensional simulation of a mass spring system and try to apply
+on possible LAOTS methods and remove unneccessary complexities, we will study
+the simple 1 dimensional simulation of a mass-spring system and try to apply
 LAOTS for it. Afterwards we will shortly discuss the application of our gained
-knowledge to the domain of hyper elastic material simulation and simulations
+knowledge to the domain of hyperelastic material simulation and simulations
 in general.
@@
 \section{Previous work}
-As according to \cite{Gander:2013} local time stepping was first studied
+According to \cite{Gander:2013} local time stepping was first studied
 in the community of ordinary differential equations (ODE) with Rice
@@
 computations either interpolation or extrapolation of the first computation
-is performed to compute the second part in each time step and the different
+is performed to compute the second part in each time step. The different
 time step sizes can be adapted to fit the model in each step as well. See
@@
 and (parabolic) heat equations instead of more general applications. We
-will adhere from describing the methods further in this report, but a
+will refrain from describing the methods further in this report, but a
 short description and comparison of the specific methods can be found in
@@
 With no prerequisites within the field of mass simulations we will first
-consider the 1D problem of simulating a mass spring system using the explicit
+consider the 1D problem of simulating a mass-spring system using the explicit
 (forward) Euler method.
@@
 Third and last we will discussion the results and how we
-could apply our knowledge of LAOTS for hyper elastic material.
+could apply our knowledge of LAOTS for hyperelastic material.
 
-\subsection{1D mass spring system}
-Mass spring systems are systems simulating a set of vertices interconnected by
+\subsection{1D mass-spring system}
+Mass-spring systems are systems simulating a set of vertices interconnected by
 springs in various ways to obtain desirable physical properties. The motion of
 the springs and vertices are then simulated as a function of the time elapsed.
-Examples of application of mass spring systems are for simulation of cloth or
-hair.
+Examples of application of mass-spring systems are simulation of cloth or hair.
 
-Our mass spring system consists of a set $V$ of vertices and a set $S$ of
+When simulating the mass-spring system we perform the following steps:
+\begin{compactenum}
+    \item Initialize system
+    \item Calculate forces
+    \item Calculate derrivatives
+    \item Update positions and velocities
+    \item Update time
+    \item Go to step 2 if the simulation has not reached it's maximum simulation
+        time.
+\end{compactenum}
+We now describe the details of each of the steps of the simulation.
+
+Our 1D mass-spring system consists of a set $V$ of vertices and a set $S$ of
 springs. Each vertex $v \in V$ has a specified position $p_v$, mass $m_v$, force
 $F_v$
-and velocity $u_v$ and each spring $s$ has two endpoints vertices $p_{sf} \in V$
-and $p_{st} \in V \setminus p_{sf}$, a damping factor $d_s$, a stiffness
-constant $K_s$ and a rest length $l_{s0}$. Combined all these variables and
-constants describe the state of the mass spring system at each timestep during a
-simulation.
+and velocity $u_v$, and each spring $s \in S$ has two endpoints vertices $p_{f,s} \in V$
+and $p_{t,s} \in V \setminus p_{f,s}$, a damping factor $d_s$, a stiffness
+constant $K_s$ and a rest length $l_{0,s}$. All these variables and
+constants combined describe the state of the mass-spring system at each time $t$ during a
+simulation. All these variables are initialized in step 1.
 
-As each vertex $v$ is connected to 1 or 2 springs, its motion can be calculated by
-the total force $F_v$ applied by the springs it is connected to. The spring
-force is calculated using Hooke's Law and adding damping [TODO: Add reference (wiki):
+The total force of a spring is the sum of the
+spring force $F_{h,d}$ calculated by using Hooke's Law and the damping force
+$F_{C,d}$ using a simple damping method [TODO: Add reference (wiki):
 \url{http://en.wikipedia.org/wiki/Harmonic_oscillator}]:
 \begin{align}
+    \nonumber l_d &= p_{l,d} - p_{r,d} \\
     \label{eq:hooke}
@@
     \label{eq:damping}
-    F_{C,d} &= -C_d \frac{dx_d}{dt} = - C_d \frac{\left(v_{left} - v_{right}\right)
-\Delta l}{\abs{\Delta l}}
+    F_{C,d} &= -C_d \frac{\partial x_d}{ \partial t} = - C_d
+        \frac{\left(v_{l,d} - v_{r,d}\right) l_d}{\abs{l_d}} \\
+    F_d &= F_{h,d} + F_{C,d}
 \end{align}
-in which $F_h$ is the spring force of Hooke's Law, $F_c$ is the damping force,
-$k$ is the spring constant, $x$ is the difference between the current length of
-the spring and the spring rest length, $c$ is the damping constant and
-$\frac{dx}{dt}$ is the velocity of the vertex $v$.
+[TODO: Maybe add direction correction: $F_d \cdot (-l_d / \abs{l_d})$]
+in which $p_{l,d}$ and $p_{r,d}$ are the positions of the left and right
+endpoints of the spring $d$ respectively, and $v_{l,d}$ and $v_{r,d}$ are
+the velocities of the left and right endpoint vertices of the spring $d$
+respectively, and $F_d$ is the total force of spring $d$.
 
-Things to write about (notes):
+When calculating the forces applied by each spring, we add the total spring
+force $F_d$ to the left endpoint of the spring and subtract $F_d$ from the right
+endpoint since both endpoints are affected equally in opposite direction by the
+spring $d$ connecting them. These spring forces are calculated in step 2 of the
+simulation.
 
-Damping force $F_d$:
+We now know how to calculate the force of the spring system and apply
+it to the vertices. In order to calculate the derrivatives of the
+system (step 3) we use Newton's second law of motion [TODO: Add reference,
+\url{http://en.wikipedia.org/wiki/Newton's_laws_of_motion}] to get the
+difference in speed $\Delta u_v$ and position $\Delta p_v$ of each vertex $v$ using
+a given time step size $\Delta t$:
 \begin{align}
-    \Delta l &= l_{left} - l_{right} \\
-    F_d      &= - C_d v(i) =  - C_d \frac{\left(v_{left} - v_{right}\right)
-\Delta l}{\abs{\Delta l}}
+    \label{eq:delta_u}
+    \Delta u_v &= \frac{\partial u_v}{\partial t} \Delta t = \frac{F_v}{m_v}
+    \Delta t \\
+    \label{eq:delta_p}
+    \Delta p_v &= u \Delta t
 \end{align}
 
-Spring force $F_s$:
+As mentioned in the beginning of this section we have chosen to use the
+explicit Euler method defined below in equation \ref{eq:explicit_euler}
+to calculate the time ``integration'' [TODO: Add reference,
+\url{http://en.wikipedia.org/wiki/Euler_method}] when stepping through the
+simulation. Given a variable $x$ as a function of the time $t$, the variable $x$
+at the time $t + \Delta t$ is defined as:
 \begin{align}
-    F_s = -kx = -k \left( l - l_0 \right)
+    \label{eq:explicit_euler}
+    x(t+\Delta t) &= x(t) + \Delta t \frac{\partial x(t)}{\partial t}
 \end{align}
+Applying this method for the timestepping of our simulation we get the two
+following position and velocity update equations using equations
+\ref{eq:delta_u} and \ref{eq:delta_p}:
+\begin{align}
+    p_v(t+\Delta t) &= p_v(t) + \Delta p_v \\
+    u_v(t+\Delta t) &= u_v(t) + \Delta u_v
+\end{align}
+We use these equations to perform step 4 of the simulation.
 
-Explicit Euler method:
-\begin{align}
-    x(t+h) &= x(t) + h \frac{\partial x(t)}{\partial t}
-\end{align}
+Step 5 is a simple update of the time: $t = t + \Delta t$ and step 6 describes
+itself.
 
-Midpoint method:
-\begin{align}
-    x(t+h) = x(t) + h \left( f \left(x + \frac{h}{2} f \left(t \right) \right) \right)
-\end{align}
+We have now fully described the simple case of simulating a mass-spring system
+using the explicit Euler method and the next subsections describe the three
+different methods/modifications for applying LAOTS to the simulation just
+described.
 
-\cite{Keshav:2007}
 \subsection{Switch}
@@
 
-\section{Results}
+\section{Experiments and results}
+
+Examples of
 
@@
 Adaptive timestepping based on highest acceleration/smallest grid
+Relate to hyperelastic materials (some linear heuristic = win)
 
@@
 %----------------------------------------------------------------------------------------
+
+\section{Conclusion}
 

### 2013-10-27 report/article.tex
Methods described

@@
 LAOTS for it. Afterwards we will shortly discuss the application of our gained
-knowledge to the domain of hyperelastic material simulation and simulations
-in general.
+knowledge to the domain of hyperelastic material simulation and simulations in
+general.
 
@@
 
-\section{Previous work}
-According to \cite{Gander:2013} local time stepping was first studied
-in the community of ordinary differential equations (ODE) with Rice
-\cite{rice:1960} developing the split Runge-Kutta methods (multi rate
-Runge-Kutta methods). The concept behind these methods is the splitting of
-the ODEs into multiple (two components in \cite{rice:1960}) components of
+\section{Previous work} According to \cite{Gander:2013} local time stepping
+was first studied in the community of ordinary differential equations (ODE)
+with Rice \cite{rice:1960} developing the split Runge-Kutta methods (multi
+rate Runge-Kutta methods). The concept behind these methods is the splitting
+of the ODEs into multiple (two components in \cite{rice:1960}) components of
 which each component needs to be integrated in different scales. A strategy
@@
 
-Third and last we will discussion the results and how we
-could apply our knowledge of LAOTS for hyperelastic material.
+Third and last we will discussion the results and how we could apply our
+knowledge of LAOTS for hyperelastic material.
 
-\subsection{1D mass-spring system}
-Mass-spring systems are systems simulating a set of vertices interconnected by
-springs in various ways to obtain desirable physical properties. The motion of
-the springs and vertices are then simulated as a function of the time elapsed.
-Examples of application of mass-spring systems are simulation of cloth or hair.
+\subsection{1D mass-spring system} Mass-spring systems are systems simulating
+a set of vertices interconnected by springs in various ways to obtain
+desirable physical properties. The motion of the springs and vertices are
+then simulated as a function of the time elapsed. Examples of application of
+mass-spring systems are simulation of cloth or hair. When performing simulations
+we want to avoid a switch of position of spring-connected vertices caused by the
+timestep size of the simulation, since this invalidates our discretization of
+the domain. Furthermore we want to perform as few timesteps as possible in order
+to get to the final simulation state by using as big timesteps as possible
+without affecting the outcome of the simulation.
 
 When simulating the mass-spring system we perform the following steps:
-\begin{compactenum}
-    \item Initialize system
-    \item Calculate forces
-    \item Calculate derrivatives
-    \item Update positions and velocities
-    \item Update time
+\begin{compactenum} \item Initialize system \item Calculate forces \item
+Calculate derrivatives \item Update positions and velocities \item Update time
 \item Go to step 2 if the simulation has not reached it's maximum simulation
-        time.
-\end{compactenum}
-We now describe the details of each of the steps of the simulation.
+time. \end{compactenum} We now describe the details of each of the steps of
+the simulation.
 
-Our 1D mass-spring system consists of a set $V$ of vertices and a set $S$ of
-springs. Each vertex $v \in V$ has a specified position $p_v$, mass $m_v$, force
-$F_v$
-and velocity $u_v$, and each spring $s \in S$ has two endpoints vertices $p_{f,s} \in V$
-and $p_{t,s} \in V \setminus p_{f,s}$, a damping factor $d_s$, a stiffness
-constant $K_s$ and a rest length $l_{0,s}$. All these variables and
-constants combined describe the state of the mass-spring system at each time $t$ during a
-simulation. All these variables are initialized in step 1.
+Our 1D mass-spring system consists of a set $V$ of vertices and a set
+$S$ of springs. Each vertex $v \in V$ has a specified position $p_v$,
+mass $m_v$, force $F_v$ and velocity $u_v$, and each spring $s \in S$
+has two endpoints vertices $p_{f,s} \in V$ and $p_{t,s} \in V \setminus
+p_{f,s}$, a damping factor $d_s$, a stiffness constant $K_s$ and a rest length
+$l_{0,s}$. All these variables and constants combined describe the state
+of the mass-spring system at each time $t$ during a simulation. All these
+variables are initialized in step 1.
 
-The total force of a spring is the sum of the
-spring force $F_{h,d}$ calculated by using Hooke's Law and the damping force
-$F_{C,d}$ using a simple damping method [TODO: Add reference (wiki):
-\url{http://en.wikipedia.org/wiki/Harmonic_oscillator}]:
-\begin{align}
-    \nonumber l_d &= p_{l,d} - p_{r,d} \\
-    \label{eq:hooke}
-    F_{h,d} &= -k_d x_d = -k_d (l_d - l_{0,d}) \\
-    \label{eq:damping}
-    F_{C,d} &= -C_d \frac{\partial x_d}{ \partial t} = - C_d
-        \frac{\left(v_{l,d} - v_{r,d}\right) l_d}{\abs{l_d}} \\
-    F_d &= F_{h,d} + F_{C,d}
-\end{align}
-[TODO: Maybe add direction correction: $F_d \cdot (-l_d / \abs{l_d})$]
-in which $p_{l,d}$ and $p_{r,d}$ are the positions of the left and right
-endpoints of the spring $d$ respectively, and $v_{l,d}$ and $v_{r,d}$ are
-the velocities of the left and right endpoint vertices of the spring $d$
-respectively, and $F_d$ is the total force of spring $d$.
+The total force of a spring is the sum of the spring force $F_{h,d}$
+calculated by using Hooke's Law and the damping force $F_{C,d}$
+using a simple damping method [TODO: Add reference (wiki):
+\url{http://en.wikipedia.org/wiki/Harmonic_oscillator}]: \begin{align}
+\nonumber l_d &= p_{l,d} - p_{r,d} \\ \label{eq:hooke} F_{h,d} &= -k_d x_d
+= -k_d (l_d - l_{0,d}) \\ \label{eq:damping} F_{C,d} &= -C_d \frac{\partial
+x_d}{ \partial t} = - C_d \frac{\left(v_{l,d} - v_{r,d}\right) l_d}{\abs{l_d}}
+\\ F_d &= F_{h,d} + F_{C,d} \end{align} [TODO: Maybe add direction correction:
+$F_d \cdot (-l_d / \abs{l_d})$] in which $p_{l,d}$ and $p_{r,d}$ are the
+positions of the left and right endpoints of the spring $d$ respectively, and
+$v_{l,d}$ and $v_{r,d}$ are the velocities of the left and right endpoint
+vertices of the spring $d$ respectively, and $F_d$ is the total force of
+spring $d$.
 
 When calculating the forces applied by each spring, we add the total spring
-force $F_d$ to the left endpoint of the spring and subtract $F_d$ from the right
-endpoint since both endpoints are affected equally in opposite direction by the
-spring $d$ connecting them. These spring forces are calculated in step 2 of the
-simulation.
+force $F_d$ to the left endpoint of the spring and subtract $F_d$ from the
+right endpoint since both endpoints are affected equally in opposite direction
+by the spring $d$ connecting them. These spring forces are calculated in step
+2 of the simulation.
 
-We now know how to calculate the force of the spring system and apply
-it to the vertices. In order to calculate the derrivatives of the
-system (step 3) we use Newton's second law of motion [TODO: Add reference,
+We now know how to calculate the force of the spring system and apply it
+to the vertices. In order to calculate the derrivatives of the system
+(step 3) we use Newton's second law of motion [TODO: Add reference,
 \url{http://en.wikipedia.org/wiki/Newton's_laws_of_motion}] to get the
-difference in speed $\Delta u_v$ and position $\Delta p_v$ of each vertex $v$ using
-a given time step size $\Delta t$:
-\begin{align}
-    \label{eq:delta_u}
+difference in speed $\Delta u_v$ and position $\Delta p_v$ of each vertex $v$
+using a given time step size $\Delta t$: \begin{align} \label{eq:delta_u}
 \Delta u_v &= \frac{\partial u_v}{\partial t} \Delta t = \frac{F_v}{m_v}
-    \Delta t \\
-    \label{eq:delta_p}
-    \Delta p_v &= u \Delta t
-\end{align}
+\Delta t \\ \label{eq:delta_p} \Delta p_v &= u \Delta t \end{align}
 
@@
 to calculate the time ``integration'' [TODO: Add reference,
-\url{http://en.wikipedia.org/wiki/Euler_method}] when stepping through the
-simulation. Given a variable $x$ as a function of the time $t$, the variable $x$
-at the time $t + \Delta t$ is defined as:
-\begin{align}
-    \label{eq:explicit_euler}
-    x(t+\Delta t) &= x(t) + \Delta t \frac{\partial x(t)}{\partial t}
-\end{align}
-Applying this method for the timestepping of our simulation we get the two
-following position and velocity update equations using equations
-\ref{eq:delta_u} and \ref{eq:delta_p}:
-\begin{align}
-    p_v(t+\Delta t) &= p_v(t) + \Delta p_v \\
-    u_v(t+\Delta t) &= u_v(t) + \Delta u_v
-\end{align}
-We use these equations to perform step 4 of the simulation.
+\url{http://en.wikipedia.org/wiki/Euler_method}] when stepping through
+the simulation. Given a variable $x$ as a function of the time $t$, the
+variable $x$ at the time $t + \Delta t$ is defined as: \begin{align}
+\label{eq:explicit_euler} x(t+\Delta t) &= x(t) + \Delta t \frac{\partial
+x(t)}{\partial t} \end{align} Applying this method for the timestepping
+of our simulation we get the two following position and velocity update
+equations using equations \ref{eq:delta_u} and \ref{eq:delta_p}: \begin{align}
+p_v(t+\Delta t) &= p_v(t) + \Delta p_v \\ u_v(t+\Delta t) &= u_v(t) + \Delta
+u_v \end{align} We use these equations to perform step 4 of the simulation.
 
@@
 
-\subsection{Switch}
-Change timestep when detecting collision/switch in next iteration
+\subsection{Adapt-on-switch}
+The first LAOT method we propose is time-adaptive. We perform a simulation
+with an initial timestep size. In each step we try to predict whether a timestep
+of the same size using the current velocities will result in a switch of the
+left and right vertices of any of the springs. If this is the case, we should
+try to avoid the switch by decreasing the timestep of the current iteration,
+re-calculate the velocity and position changes and resume stepping of the
+simulation using the new timestep size. When switches aren't detected anymore
+we gradually increase the timestep of the simulation again. We call this
+method the adapt-on-switch method.
 
-\subsection{Spatial LAOTS}
-Atomic update
+The switch of two endspoints connected by a spring $d$ can easily be
+calculated by checking the length ($p_{l,d} - p_{r,d}$) of the spring before
+and after a position update. If the sign of the length differs, the two
+points must have switched sides and hence adaption of the timestep should be
+performed. The adaption of the timestep can either be done by division of the
+current timestep by a constant or by having a finite set of timestep sizes to
+choose from.
 
-Fixed update rate and detection of need for re-calculation of individual springs
+\subsection{Atomic updating using heuristics}
+The second LAOT method we propose is space-adaptive. We perform simulations
+using a fixed timestep size. In each iteration we perform position update of
+all vertices, but only re-calculate acceleration and velocity of the springs
+and points that fulfill one of the defined heuristics. The heuristics can be
+chosen to fit the domain of the simulation, but we have chosen to use the
+number of steps since last update and the ratio between the current spring
+length and the spring length at the last update. This should force a bound on
+the maximal timestep before update as well as making sure updates are done
+more frequently in case the springlength is changing more rapidly. Further
+details are explained in section \ref{sec:experiments}.
 
-Heuristic: Length of spring, steps since last update
+\subsection{Global estimated acceleration}
+The third and last LAOT method we propose is time-adaptive. The simulations are
+done using an initial timestep size. We know that the force and hence
+acceleration of a spring is somewhat linear close to the spring's rest length
+and it ``bends'' off as the spring is streched or contracted further away from
+the rest length. Based on this knowledge we will try to make a heuristic
+telling us about the needed timestep size at any given point during the
+simulation. The smallest needed timestep defines the global timestep for that
+given step.
 
-\subsection{Inverse}
-Time-local optimization according to heuristic:
-
+For this simple 1D example we use the inverse of the acceleration
+for adapting the timestep and hence the heuristic is defined as:
 \begin{align}
     \frac{1}{a} &= \frac{1}{\frac{-kx}{m}} = \frac{m}{-kx} \\
-    \Delta t &= \abs{\frac{1}{a}} + \frac{1}{0.1}
+    \Delta t &= \text{min}\left( \frac{1}{\abs{a} + 0.1} \right )
 \end{align}
+We have added a small value (0.1) in order to avoid the timestep to increase
+exponentially towards infinity as $\abs{a}$ approach 0. Furthermore a scaling of the timestep
+could be needed depending on the scale of the simulation mesh. Further details
+are explained in section \ref{sec:experiments}.
+
+Our though is that for some more
+complex simulations we are able to make estimates of this type less costly than
+performing the actual simulations of a small fixed size timestep.
 
@@
 \section{Experiments and results}
+\label{sec:experiments}
+
+Show results for:
+
+\begin{itemize}
+    \item Normal simulation
+    \item Switch
+    \item Atomic
+    \item Inverse
+\end{itemize}
 

### 2013-10-29 report/article.tex
Report section on experiments for standard simulation

@@
 
-\documentclass[twoside]{article}
-
-\usepackage{lipsum} % Package to generate dummy text throughout this template
+\documentclass[11pt]{article}
 
@@
 
-\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
-\usepackage{multicol} % Used for the two-column layout of the document
+%\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
+%\usepackage{multicol} % Used for the two-column layout of the document
 \usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
@@
 \usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text
-\usepackage{paralist} % Used for the compactitem environment which makes bullet points with less space between them
+%\usepackage{paralist} % Used for the compactitem environment which makes bullet points with less space between them
 
@@
 \usepackage{amsmath,amssymb,amsfonts}
+\usepackage{subcaption}
 \providecommand{\abs}[1]{\left \lvert #1 \right \rvert}
@@
 
-\begin{multicols}{2} % Two-column layout throughout the main article text
+%\begin{multicols}{2} % Two-column layout throughout the main article text
 
@@
 strict bounds on size of the time step of the simulation in order to keep
-the materials (and grids) stable and the discretization correct. In a
-``standard'' simulation with a global time step size we need to compute
+the materials (and grids) stable and the discretization correct.
+
+In a ``standard'' simulation with a global time step size we need to compute
 possibly unnecessarily many computations on the coarser parts of the material
@@
 locality (as described above) or time locality where the time step is global
-for all vertices but varies during the simulation. We will in this report
+for all vertices but varies during the simulation.
+
+We will in this report
 study possible approaches for application of LAOTS. In order to keep the focus
@@
 \section{Methods}
+\label{sec:methods}
 
@@
 When simulating the mass-spring system we perform the following steps:
-\begin{compactenum} \item Initialize system \item Calculate forces \item
+\begin{enumerate} \item Initialize system \item Calculate forces \item
 Calculate derrivatives \item Update positions and velocities \item Update time
 \item Go to step 2 if the simulation has not reached it's maximum simulation
-time. \end{compactenum} We now describe the details of each of the steps of
+time. \end{enumerate} We now describe the details of each of the steps of
 the simulation.
@@
 calculated by using Hooke's Law and the damping force $F_{C,d}$
-using a simple damping method [TODO: Add reference (wiki):
-\url{http://en.wikipedia.org/wiki/Harmonic_oscillator}]: \begin{align}
+using a simple damping method \todo{Add reference (wiki):
+    \url{http://en.wikipedia.org/wiki/Harmonic_oscillator}}: \begin{align}
 \nonumber l_d &= p_{l,d} - p_{r,d} \\ \label{eq:hooke} F_{h,d} &= -k_d x_d
@@
 x_d}{ \partial t} = - C_d \frac{\left(v_{l,d} - v_{r,d}\right) l_d}{\abs{l_d}}
-\\ F_d &= F_{h,d} + F_{C,d} \end{align} [TODO: Maybe add direction correction:
-$F_d \cdot (-l_d / \abs{l_d})$] in which $p_{l,d}$ and $p_{r,d}$ are the
+\\ F_d &= F_{h,d} + F_{C,d} \end{align} \todo{Maybe add direction correction:
+    $F_d \cdot (-l_d / \abs{l_d})$} in which $p_{l,d}$ and $p_{r,d}$ are the
 positions of the left and right endpoints of the spring $d$ respectively, and
@@
 to the vertices. In order to calculate the derrivatives of the system
-(step 3) we use Newton's second law of motion [TODO: Add reference,
-\url{http://en.wikipedia.org/wiki/Newton's_laws_of_motion}] to get the
+(step 3) we use Newton's second law of motion \todo{Add reference,
+    \url{http://en.wikipedia.org/wiki/Newton's_laws_of_motion}} to get the
 difference in speed $\Delta u_v$ and position $\Delta p_v$ of each vertex $v$
@@
 explicit Euler method defined below in equation \ref{eq:explicit_euler}
-to calculate the time ``integration'' [TODO: Add reference,
-\url{http://en.wikipedia.org/wiki/Euler_method}] when stepping through
+to calculate the time ``integration'' \todo{Add reference,
+    \url{http://en.wikipedia.org/wiki/Euler_method}} when stepping through
 the simulation. Given a variable $x$ as a function of the time $t$, the
@@
 all vertices, but only re-calculate acceleration and velocity of the springs
-and points that fulfill one of the defined heuristics. The heuristics can be
-chosen to fit the domain of the simulation, but we have chosen to use the
-number of steps since last update and the ratio between the current spring
-length and the spring length at the last update. This should force a bound on
-the maximal timestep before update as well as making sure updates are done
-more frequently in case the springlength is changing more rapidly. Further
-details are explained in section \ref{sec:experiments}.
+and points that fulfill some defined heuristics.
+
+The heuristics can be chosen to fit the domain of the simulation, but we have
+chosen to use the number of steps since last update and the ratio between the
+current spring length and the spring length at the last update. This should
+force a bound on the maximal timestep before update as well as making sure
+updates are done more frequently in case the springlength is changing more
+rapidly. Further details are explained in section \ref{sec:experiments}.
 
@@
 We have added a small value (0.1) in order to avoid the timestep to increase
-exponentially towards infinity as $\abs{a}$ approach 0. Furthermore a scaling of the timestep
-could be needed depending on the scale of the simulation mesh. Further details
-are explained in section \ref{sec:experiments}.
+exponentially towards infinity as $\abs{a}$ approach 0. Furthermore a scaling
+of the timestep could be needed depending on the scale of the simulation mesh.
+Further details are explained in section \ref{sec:experiments}.
 
@@
 
-Show results for:
+We first give an example of the actual application of LAOTS using the standard
+simulation of mass spring particles. Figure \ref{fig:std} shows two simulations
+of the same set of particles distributed with two different scales simulated for
+500 seconds each. All the springs $s \in S$ are initiated in a stretched state/position
+with a length 1.25 times longer than the rest length, damping and stiffness
+constants $C_s = 0.5, K_s = 0.5$, and all vertices $v \in V$
+have a mass of $m_v = 30$.
+\ref{fig:std_multi_30fps} shows a simulation using 30 fps and Figure
+\ref{fig:std_multi_1fps} shows a simulation using 1 fps. We clearly see that
+using a timestep of 30 fps the simulation is stable and able to handle the
+compression of all the small springs around $t = 110$ whereas a switch occurs
+when simulating at 1 fps. This switch causes ripples of sideeffects for the rest
+of the simulation and hence the final result is incorrect. No switches of the longer springs are however occuring and hence these pose no problem at 1 fps. We will use this simulation for
+testing all three LAOTS methods proposed in section \ref{sec:methods}.
 
-\begin{itemize}
-    \item Normal simulation
-    \item Switch
-    \item Atomic
-    \item Inverse
-\end{itemize}
+\begin{figure}[H]
+    \begin{subfigure}{0.5\textwidth}
+        \includegraphics[width=\textwidth]{../images/standard_multiscale_30fps.png}
+        \caption{30 fps}
+        \label{fig:std_multi_30fps}
+    \end{subfigure}
+    \begin{subfigure}{0.5\textwidth}
+        \includegraphics[width=\textwidth]{../images/standard_multiscale_1fps.png}
+        \caption{1 fps}
+        \label{fig:std_multi_1fps}
+    \end{subfigure}
+    \caption{Standard simulations of particles distributed with two scales using
+    different timestep sizes}
+    \label{fig:std}
+\end{figure}
 
-Examples of
+\todo{Write something about the explicit euler based on uniform}
+\begin{figure}[H]
+    \begin{subfigure}{0.5\textwidth}
+        \includegraphics[width=\textwidth]{../images/standard_uniform_30fps.png}
+        \caption{30 fps}
+        \label{fig:std_uni_30fps}
+    \end{subfigure}
+    \begin{subfigure}{0.5\textwidth}
+        \includegraphics[width=\textwidth]{../images/standard_uniform_1fps.png}
+        \caption{1 fps}
+        \label{fig:std_uni_1fps}
+    \end{subfigure}
+    \caption{Standard simulations of particles distributed uniformly using
+    different timestep sizes}
+\end{figure}
 
+\subsection{Adapt-on-switch}
+Special things to note: How subdivision is done: new fixed (30) or relative (dt / 10)
+new fixed: We don't catch it early enough. Maybe something with the explicit
+euler, maybe something with the linear model
+
+Relative: We get too fine-grained - maybe related to the above
+
+\subsection{Atomic updating using heuristics}
+
+Heuristic: steps since last update and ratio of length since last update
+
+\subsection{Global estimated acceleration}
+
+%Examples of
+%
 %\begin{table}[H]
@@
 
-\end{multicols}
-
 \end{document}

### 2013-10-29 src/1D/run.m

@@ -6,6 +6,18 @@ clc;
 % Set up simulation parameters
 t_stop = 500;
 
+pP = [-2:1:2];
+[positions, ~, times, ~] = springSim(30, t_stop, pP);
+fi1 = figure();
+
+plot(times, positions);
+hold on;
+axis([0 max(times) min(positions(:)) max(positions(:))]);
+xlabel('Time');
+ylabel('Position');
+
+
+return;
 
 %% Atomic:
 fpses =      [10, 10, 30];
@@ -43,7 +55,7 @@ fpses = [10, 2, 0.15, 30];
 results = cell(length(fpses),4);
 for i = 1:length(fpses)
     disp(['Simulation: ' num2str(fpses(i))]);
-    [results{i,1}, results{i,2}, results{i,3}, results{i,4}] = springSim(fpses(i),t_stop);
+    [results{i,1}, results{i,2}, results{i,3}, results{i,4}] = switchSim(fpses(i),t_stop);
 end
 
 displaySim(results, fpses);

### 2013-10-29 src/1D/runStandard.m

@@ -1,0 +1,50 @@
+% Set up simulation parameters
+t_stop = 500;
+
+%% Uniform plots
+pP = [-2:1:2];
+[positions, ~, times, ~] = springSim(30, t_stop, pP);
+fi1 = figure();
+impath = '../../images/';
+
+plot(times, positions);
+hold on;
+axis([0 max(times) min(positions(:)) max(positions(:))]);
+xlabel('Time');
+ylabel('Position');
+title('Standard simulation (30 fps)');
+saveas(fi1, [impath 'standard_uniform_30fps.png'])
+
+[positions, ~, times, ~] = springSim(1, t_stop, pP);
+fi2 = figure();
+impath = '../../images/';
+
+plot(times, positions);
+hold on;
+axis([0 max(times) min(positions(:)) max(positions(:))]);
+xlabel('Time');
+ylabel('Position');
+title('Standard simulation (1 fps)');
+saveas(fi2, [impath 'standard_uniform_1fps.png'])
+
+%% Multi-scale plots
+pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
+[positions, ~, times, ~] = springSim(30, t_stop, pP);
+fi3 = figure();
+plot(times, positions);
+hold on
+axis([0 max(times) min(positions(:)) max(positions(:))]);
+xlabel('Time');
+ylabel('Position');
+title('Standard simulation (30 fps)');
+saveas(fi3, [impath 'standard_multiscale_30fps.png'])
+
+[positions, ~, times, ~] = springSim(1, t_stop, pP);
+fi4 = figure();
+plot(times, positions);
+hold on
+axis([0 max(times) min(positions(:)) max(positions(:))]);
+xlabel('Time');
+ylabel('Position');
+title('Standard simulation (1 fps)');
+saveas(fi4, [impath 'standard_multiscale_1fps.png'])

### 2013-10-29 src/1D/switchSim.m

@@ -1,9 +1,7 @@
-function [positions, lengths, times, switches] = springSim(fps_lowest, t_stop, pP)
+function [positions, lengths, times, switches] = switchSim(fps_lowest, t_stop)
     % Create grid of particles at positions
-    if nargin < 3
-        pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
-        %pP = [-2:1:2];
-    end
+    pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
+    %pP = [-2:1:2];
     if size(pP,2) > size(pP,1)
         pP = pP';
     end
@@ -31,7 +29,7 @@ function [positions, lengths, times, switches] = springSim(fps_lowest, t_stop, p
     spC = ones(cntS,1) * 0.5;   % Damping
     spK = ones(cntS,1) * 0.5;   % Stiffness
     spX0 = pP(sp);
-    spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.8;
+    spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.785;
 
     distance_show = false;
     momentum_show = false;
@@ -71,19 +69,46 @@ function [positions, lengths, times, switches] = springSim(fps_lowest, t_stop, p
     dx = zeros(cntS,1);
     pF = zeros(cntP,2);
     while time <= t_stop + dt/2;
+        p_last = pP;
+        pP = pP + pV * dt;
+
+        % Check for subdivision
+        if dt < dt_base && size(pSub,1) > 0
+
+            % Find springs to re-calculate force for
+            springs = sp(ismember(sp,pSub));
+        end
+
         % Force calculation
         [pF dx] = springForces(pP, sp, pV, spK, spX0, spC);
 
-        % Position and velocity update values
+        % Position and velocity update
         [dp dv] = springStep(pP, pM, pV, sum(pF,2), dt);
+
+        % Detect collisions / switches in position
+        l = springLength(p_last, sp);
+        l2 = springLength(pP + dp, sp);
+        mask = l .* l2 < 0;
+        if any(mask)
+%            dt = dt/10;
+            dt = 1/30;
+            disp('Switch in next iteration!');
+            pSub = sp(mask,:);
+            cnt = size(pSub,1);
+            switches(end+1:end+cnt,:) = [repmat(time,[cnt 1]), pSub];
+%            dt = dt/10;
+%            pV(pSub) = 0;
[…]

### 2013-10-29 report/article.tex
Added references for various simple concepts

@@
 The total force of a spring is the sum of the spring force $F_{h,d}$
-calculated by using Hooke's Law and the damping force $F_{C,d}$
-using a simple damping method \todo{Add reference (wiki):
-    \url{http://en.wikipedia.org/wiki/Harmonic_oscillator}}: \begin{align}
-\nonumber l_d &= p_{l,d} - p_{r,d} \\ \label{eq:hooke} F_{h,d} &= -k_d x_d
-= -k_d (l_d - l_{0,d}) \\ \label{eq:damping} F_{C,d} &= -C_d \frac{\partial
-x_d}{ \partial t} = - C_d \frac{\left(v_{l,d} - v_{r,d}\right) l_d}{\abs{l_d}}
-\\ F_d &= F_{h,d} + F_{C,d} \end{align} \todo{Maybe add direction correction:
-    $F_d \cdot (-l_d / \abs{l_d})$} in which $p_{l,d}$ and $p_{r,d}$ are the
-positions of the left and right endpoints of the spring $d$ respectively, and
-$v_{l,d}$ and $v_{r,d}$ are the velocities of the left and right endpoint
-vertices of the spring $d$ respectively, and $F_d$ is the total force of
-spring $d$.
+calculated by using Hooke's Law (see \cite[p.~439]{Young:2010})
+and the damping force $F_{C,d}$ using a simple damping method (see
+\cite[p.~457]{Young:2010}): \begin{align} \nonumber l_d &= p_{l,d} -
+p_{r,d} \\ \label{eq:hooke} F_{h,d} &= -k_d x_d = -k_d (l_d - l_{0,d}) \\
+\label{eq:damping} F_{C,d} &= -C_d \frac{\partial x_d}{ \partial t} = - C_d
+\frac{\left(v_{l,d} - v_{r,d}\right) l_d}{\abs{l_d}} \\ F_d &= F_{h,d} +
+F_{C,d} \end{align} \todo{Maybe add direction correction: $F_d \cdot (-l_d
+/ \abs{l_d})$} in which $p_{l,d}$ and $p_{r,d}$ are the positions of the
+left and right endpoints of the spring $d$ respectively, and $v_{l,d}$ and
+$v_{r,d}$ are the velocities of the left and right endpoint vertices of the
+spring $d$ respectively, and $F_d$ is the total force of spring $d$.
 
@@
 
-We now know how to calculate the force of the spring system and apply it
-to the vertices. In order to calculate the derrivatives of the system
-(step 3) we use Newton's second law of motion \todo{Add reference,
-    \url{http://en.wikipedia.org/wiki/Newton's_laws_of_motion}} to get the
-difference in speed $\Delta u_v$ and position $\Delta p_v$ of each vertex $v$
-using a given time step size $\Delta t$: \begin{align} \label{eq:delta_u}
+We now know how to calculate the force of the spring system and apply it to
+the vertices. In order to calculate the derrivatives of the system (step 3) we
+use Newton's second law of motion (see \cite[p.~112-115]{Young:2010}) to get
+the difference in speed $\Delta u_v$ and position $\Delta p_v$ of each vertex
+$v$ using a given time step size $\Delta t$: \begin{align} \label{eq:delta_u}
 \Delta u_v &= \frac{\partial u_v}{\partial t} \Delta t = \frac{F_v}{m_v}
@@
 explicit Euler method defined below in equation \ref{eq:explicit_euler}
-to calculate the time ``integration'' \todo{Add reference,
-    \url{http://en.wikipedia.org/wiki/Euler_method}} when stepping through
+to calculate the time ``integration'' \cite{Flaherty} when stepping through
 the simulation. Given a variable $x$ as a function of the time $t$, the

### 2013-10-29 src/1D/switchSim.m
Changed switch to use proper explicit euler

@@ -70,7 +70,6 @@ function [positions, lengths, times, switches] = switchSim(fps_lowest, t_stop)
     pF = zeros(cntP,2);
     while time <= t_stop + dt/2;
         p_last = pP;
-        pP = pP + pV * dt;
 
         % Check for subdivision
         if dt < dt_base && size(pSub,1) > 0
@@ -109,6 +108,7 @@ function [positions, lengths, times, switches] = switchSim(fps_lowest, t_stop)
         times(end+1) = time;
 
         % Subdivide
+        pP = pP + dp;
         pV = pV + dv;
 
         % Time update

### 2013-10-29 report/article.tex
Added examples for switch

@@
 new fixed: We don't catch it early enough. Maybe something with the explicit
-euler, maybe something with the linear model
+euler, maybe something with the linear model. We should catch the error before
+it escalates into a state in which we cannot rescue it anymore.
 
 Relative: We get too fine-grained - maybe related to the above
+\begin{figure}[H]
+    \begin{subfigure}{0.5\textwidth}
+        \includegraphics[width=\textwidth]{../images/switch_multiscale_1fps_relative.png}
+        \caption{Relative timestepping ($\Delta t = \frac{\Delta t}{100}$)}
+        \label{fig:switch_multi_1fps_relative}
+    \end{subfigure}
+    \begin{subfigure}{0.5\textwidth}
+        \includegraphics[width=\textwidth]{../images/switch_multiscale_1fps_absolute.png}
+        \caption{Absolute timestepping ($\Delta t = \frac{1}{30}$)}
+        \label{fig:switch_multi_1fps_absolute}
+    \end{subfigure}
+    \caption{Adapt-on-switch simulation of particles distributed with two
+    scales using 1 fps}
+\end{figure}
 
@@
 
+interpolation of positions makes the integration more stable since the stepping
+is based on a point closer in time
+
 \subsection{Global estimated acceleration}
+
+The heuristic dampens the system
 

### 2013-10-29 src/1D/atomicSim.m
Changed switch scheme to use relative or absolute subdivision

@@ -1,8 +1,10 @@
-function [positions, lengths, times, switches, active] = springSim(fps, t_stop, stepBound, rFactor)
+function [positions, lengths, times, switches, active] = springSim(fps, t_stop, stepBound, rFactor, pP)
     %% Initialize particles
-    % Create grid of particles at positions
-    %pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
-    pP = [-2:2:2]';
+    if nargin < 5
+        % Create grid of particles at positions
+        %pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
+        pP = [-2:2:2]';
+    end
 
     % Initialize variables
     cntP = size(pP,1);

### 2013-10-29 src/1D/run.m

@@ -55,7 +55,7 @@ fpses = [10, 2, 0.15, 30];
 results = cell(length(fpses),4);
 for i = 1:length(fpses)
     disp(['Simulation: ' num2str(fpses(i))]);
-    [results{i,1}, results{i,2}, results{i,3}, results{i,4}] = switchSim(fpses(i),t_stop);
+    [results{i,1}, results{i,2}, results{i,3}, results{i,4}] = switchSim(fpses(i),t_stop, false);
 end
 
 displaySim(results, fpses);

### 2013-10-29 src/1D/runSwitch.m

@@ -1,0 +1,25 @@
+% Set up simulation parameters
+t_stop = 500;
+impath = '../../images/';
+
+%% Multi-scale plots
+pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
+[positions, ~, times, ~] = switchSim(1, t_stop, true, pP);
+fi1 = figure();
+plot(times, positions);
+hold on
+axis([0 max(times) min(positions(:)) max(positions(:))]);
+xlabel('Time');
+ylabel('Position');
+title('Standard simulation (1 fps, relative)');
+saveas(fi1, [impath 'switch_multiscale_1fps_relative.png'])
+
+[positions, ~, times, ~] = switchSim(1, t_stop, false, pP);
+fi2 = figure();
+plot(times, positions);
+hold on
+axis([0 max(times) min(positions(:)) max(positions(:))]);
+xlabel('Time');
+ylabel('Position');
+title('Standard simulation (1 fps, absolute)');
+saveas(fi2, [impath 'switch_multiscale_1fps_absolute.png'])

### 2013-10-29 src/1D/switchSim.m

@@ -1,7 +1,10 @@
-function [positions, lengths, times, switches] = switchSim(fps_lowest, t_stop)
-    % Create grid of particles at positions
-    pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
-    %pP = [-2:1:2];
+function [positions, lengths, times, switches] = switchSim(fps_lowest, t_stop, relative, pP)
+    if nargin < 4
+        % Create grid of particles at positions
+        pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
+        %pP = [-2:1:2];
+    end
+
     if size(pP,2) > size(pP,1)
         pP = pP';
     end
@@ -29,7 +32,7 @@ function [positions, lengths, times, switches] = switchSim(fps_lowest, t_stop)
     spC = ones(cntS,1) * 0.5;   % Damping
     spK = ones(cntS,1) * 0.5;   % Stiffness
     spX0 = pP(sp);
-    spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.785;
+    spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.8;
 
     distance_show = false;
     momentum_show = false;
@@ -69,14 +72,6 @@ function [positions, lengths, times, switches] = switchSim(fps_lowest, t_stop)
     dx = zeros(cntS,1);
     pF = zeros(cntP,2);
     while time <= t_stop + dt/2;
-        p_last = pP;
-
-        % Check for subdivision
-        if dt < dt_base && size(pSub,1) > 0
-
-            % Find springs to re-calculate force for
-            springs = sp(ismember(sp,pSub));
-        end
 
         % Force calculation
         [pF dx] = springForces(pP, sp, pV, spK, spX0, spC);
@@ -85,21 +80,29 @@ function [positions, lengths, times, switches] = switchSim(fps_lowest, t_stop)
         [dp dv] = springStep(pP, pM, pV, sum(pF,2), dt);
 
         % Detect collisions / switches in position
-        l = springLength(p_last, sp);
+        l = springLength(pP, sp);
         l2 = springLength(pP + dp, sp);
         mask = l .* l2 < 0;
         if any(mask)
-%            dt = dt/10;
-            dt = 1/30;
+            if relative
+                dt = dt/100
+            else
+                dt = 1/30;
+            end
             disp('Switch in next iteration!');
             pSub = sp(mask,:);
             cnt = size(pSub,1);
[…]

### 2013-10-31 report/article.tex
Report figures generated by runReport and report section experiments for switch and atomic (half done)

@@
 different methods/modifications for applying LAOTS to the simulation just
-described.
+described. The three methods are using different ideas to implement LOATS. They
+could however possibly be combined in case they turn out to be working as
+intended to achieve the best possible LAOTS scheme.
 
 \subsection{Adapt-on-switch}
+\label{sec:method_switch}
 The first LAOT method we propose is time-adaptive. We perform a simulation
@@
 
-\subsection{Atomic updating using heuristics}
+\subsection{Atomic Updating using Heuristics}
 The second LAOT method we propose is space-adaptive. We perform simulations
@@
 all vertices, but only re-calculate acceleration and velocity of the springs
-and points that fulfill some defined heuristics.
+and points that fulfill some defined heuristics and hence the method is called
+``Atomic Updating using Heuristics'' (AUH).
 
-The heuristics can be chosen to fit the domain of the simulation, but we have
-chosen to use the number of steps since last update and the ratio between the
-current spring length and the spring length at the last update. This should
-force a bound on the maximal timestep before update as well as making sure
-updates are done more frequently in case the springlength is changing more
-rapidly. Further details are explained in section \ref{sec:experiments}.
+The heuristics can be chosen to fit the domain of the simulation, but we
+have chosen to use the number of steps since last update, $n_{last}$, and
+the ratio between the current spring length and the spring length at the
+last update $r_{last}$. This should force a bound on the maximal timestep
+before update as well as making sure updates are done more frequently in case
+the springlength is changing more rapidly. Further details are explained in
+section \ref{sec:experiments}.
 
@@
 compression of all the small springs around $t = 110$ whereas a switch occurs
-when simulating at 1 fps. This switch causes ripples of sideeffects for the rest
-of the simulation and hence the final result is incorrect. No switches of the longer springs are however occuring and hence these pose no problem at 1 fps. We will use this simulation for
-testing all three LAOTS methods proposed in section \ref{sec:methods}.
+when simulating at 1 fps. This switch causes ripples of sideeffects for the
+rest of the simulation and hence the final result is incorrect. No switches of
+the longer springs are however occuring and hence these pose no problem at 1
+fps. We will use this simulation for testing all three LAOTS methods proposed
+in section \ref{sec:methods}.
 
@@
 
-\todo{Write something about the explicit euler based on uniform}
-\begin{figure}[H]
-    \begin{subfigure}{0.5\textwidth}
-        \includegraphics[width=\textwidth]{../images/standard_uniform_30fps.png}
-        \caption{30 fps}
-        \label{fig:std_uni_30fps}
-    \end{subfigure}
-    \begin{subfigure}{0.5\textwidth}
-        \includegraphics[width=\textwidth]{../images/standard_uniform_1fps.png}
-        \caption{1 fps}
-        \label{fig:std_uni_1fps}
-    \end{subfigure}
-    \caption{Standard simulations of particles distributed uniformly using
-    different timestep sizes}
-\end{figure}
+\todo{Write something about the explicit euler based on uniform example}
+%\begin{figure}[H]
+%    \begin{subfigure}{0.5\textwidth}
+%        \includegraphics[width=\textwidth]{../images/standard_uniform_30fps.png}
+%        \caption{30 fps}
+%        \label{fig:std_uni_30fps}
+%    \end{subfigure}
+%    \begin{subfigure}{0.5\textwidth}
+%        \includegraphics[width=\textwidth]{../images/standard_uniform_1fps.png}
+%        \caption{1 fps}
+%        \label{fig:std_uni_1fps}
+%    \end{subfigure}
+%    \caption{Standard simulations of particles distributed uniformly using
+%    different timestep sizes}
+%\end{figure}
 
 \subsection{Adapt-on-switch}
-Special things to note: How subdivision is done: new fixed (30) or relative (dt / 10)
-new fixed: We don't catch it early enough. Maybe something with the explicit
-euler, maybe something with the linear model. We should catch the error before
-it escalates into a state in which we cannot rescue it anymore.
+As mentioned in section \ref{sec:method_switch} the adapt-on-switch method
+can be implemented using either a relative or a absolute adaption of the
+timestep. In our experiments we have tested out both methods defining the
+relative adaption to scale the timestep by a factor of $\frac{1}{100}$ and
+the absolute solution to set the timestep to $\frac{1}{30}$ when switches are
+detected since this generated nice simulations for the standard simulation .
+Furthermore the up-scaling for the relative scheme is set to a factor of 10
+since we wish to avoid running into new switches immediately.
 
-Relative: We get too fine-grained - maybe related to the above
+Figure \ref{fig:switch} shows the result of running simulations
+described earlier at 1 fps using the adapt-on-switch scheme, where
+the red crosses mark timesteps where a switch is predicted in the
+following step and hence the schemes are ``activated''. Figure
+\ref{fig:switch_multi_1fps_relative} is simulated using the relative scheme
+and figure \ref{fig:switch_multi_1fps_absolute} is simulated using the
+absolute scheme. As we clearly see in both figures the schemes fail to correct
+the simulations in spite of the switch prediction. Once the first switch
+occurs the simulation will change behavior and hence the first red cross marks
+the point at which the simulation can be rejected from being correct.
+
+After having made these observations we furthermore tested out a third way of
+creating the scheme by predicting how long a timestep we could take before the
+switch would occur and then divide this steplength by 2 to get a new timestep
+length. This generated infeasibly small timesteps forcing us to stop the
+simulation too early.
+
+Taking these observations into account we can safely say that the
+adapt-on-switch method/scheme is not working for our simulation. This does
+however not mean that the method couldn't work for other domains or other
+refinements of the method. The problem with the scheme is that the prediction
+of switches is done too late in the simulation and hence once we detect
+the switches it's too late to do anything about them given our model of
+the mass-spring system. In order for the scheme to work properly it should
+therefore be modified to be able to perform the detection of switches earlier
+in the simulation.
+
+In the real world the spring force would ``bend'' towards $\pm \infty$ once
+the spring is stretched or contracted too much and hence the linear model that
+we use doesn't suffice when using the adapt-on-switch scheme. Having used a
+more realistic choice of model could possibly have had a positive impact of
+the use of this scheme since this would cause the spring force of the spring
+in question to grow in order to avoid the switch. \todo{Maybe the section
+above should be moved to a discussion section.}
+
 \begin{figure}[H]
@@
     \caption{Adapt-on-switch simulation of particles distributed with two
-    scales using 1 fps}
+scales using 1 fps. The red crosses mark times at which the scheme predicts a
+switch in the following iteration.}
+    \label{fig:switch}
 \end{figure}
 
-\subsection{Atomic updating using heuristics}
+\subsection{Atomic Updating using Heuristics}
+The AUH method concists of two heuristics for determining whether an element
+should be updated or not: steps since last update $l_{last}$, and the ratio
+between the spring length in last update and the current length $r_{last}$.
 
-Heuristic: steps since last update and ratio of length since last update
+We first test the method using only the first heuristic which will give
+us synchronized updates of all vertices every $r_{last}$ step. Using the
+same setup as the standard simulation we perform simulations using the AUH
+method with the length ratio heuristic ignored by setting the limit of
+the ratio to infinity. We set the timestep to $\Delta t = \frac{1}{10}$
+(10 fps) and vary $l_{last}$. Figure \ref{fig:atomic_multi_sbound} shows
+the resulting simulations where figures \ref{fig:atomic_multi_10sbound},
+\ref{fig:atomic_multi_68sbound}, and \ref{fig:atomic_multi_69sbound} uses
+an $l_{last}$ of 10, 68, and 69 respectively.
 
-interpolation of positions makes the integration more stable since the stepping
-is based on a point closer in time
+Simulating using 10 fps and $l_{last} = 10$ gives us the same amount
+of velocity updates as the standard simulation using 1 fps (which had
+switches), but gives us a nice and smooth simulation with no problems.
+As can be seen in the figures \ref{fig:atomic_multi_68sbound} and
+\ref{fig:atomic_multi_69sbound} we are able to perform the simulation using
+$l_{last} = 68$ without having switches occuring whereas using $l_{last} =
+69$ results in switches occuring. In other words we are able to perform the
+simulation at 10 fps but with 6.8 seconds between acceleration and velocity
+calculations. This is a result of combining the explicit Euler method with
+the AUH scheme since the scheme simply extrapolates the position at the last
+acceleration and velocity update to a position close in time to the next
+update position and then use this position when updating the acceleration and
+velocity again. Had we chosen an alternative time integration method this
+advantage would most likely have been eliminated and let us perform the larger
+timesteps with the standard simulation.
+
+We now turn our attention towards the second heuristic.
+\todo{Decide on the exact heuristic to display after having talked to Sune.}
+
+\begin{figure}[H]
+    \begin{subfigure}{0.3\textwidth}
+        \includegraphics[width=\textwidth]{../images/atomic_multiscale_10fps_10sbound.png}
+        \caption{$l_{last} = 10$}
+        \label{fig:atomic_multi_10sbound}
+    \end{subfigure}
+    \begin{subfigure}{0.3\textwidth}
+        \includegraphics[width=\textwidth]{../images/atomic_multiscale_10fps_68sbound.png}
+        \caption{$l_{last} = 68$}
+        \label{fig:atomic_multi_68sbound}
+    \end{subfigure}
+    \begin{subfigure}{0.3\textwidth}
+        \includegraphics[width=\textwidth]{../images/atomic_multiscale_10fps_69sbound.png}
+        \caption{$l_{last} = 69$}
+        \label{fig:atomic_multi_69sbound}
+    \end{subfigure}
+    \caption{Simulations using the AUH method at 10 fps with various number of
+    steps since last update. Length heuristic is ignored (set to $\infty$).}
+    \label{fig:atomic_multi_sbound}
+\end{figure}
+
+\begin{figure}[H]
+    \begin{subfigure}{0.5\textwidth}
+        \includegraphics[width=\textwidth]{../images/atomic_uniform_10fps_0.png}
+        \caption{}
+        \label{fig:atomic_uniform_10fps_0}
+    \end{subfigure}
+    \begin{subfigure}{0.5\textwidth}
+        \includegraphics[width=\textwidth]{../images/atomic_uniform_10fps_1.png}
+        \caption{}
+        \label{fig:atomic_uniform_10fps_1}
+    \end{subfigure}
+\end{figure}
+
+if ratio then too much dampening caused by lack of updates: Example: Simple with
+3 springs
 
@@
 The heuristic dampens the system
+\begin{figure}[H]
+    \begin{subfigure}{0.5\textwidth}
+        \includegraphics[width=\textwidth]{../images/inverse_uniform_10fps_0.png}
+        \caption{}
+        \label{fig:inverse_uniform_10fps_0}
+    \end{subfigure}
+    \begin{subfigure}{0.5\textwidth}
+        \includegraphics[width=\textwidth]{../images/inverse_uniform_10fps_1.png}
+        \caption{}
+        \label{fig:inverse_uniform_10fps_1}
+    \end{subfigure}
+\end{figure}
 

### 2013-10-31 src/1D/atomicSim.m

@@ -1,9 +1,18 @@
-function [positions, lengths, times, switches, active] = springSim(fps, t_stop, stepBound, rFactor, pP)
+function [positions, lengths, times, switches, active] = atomicSim(fps, t_stop, stepBound, rFactor, pP, x0factor, dFactor, sFactor)
     %% Initialize particles
     if nargin < 5
         % Create grid of particles at positions
         %pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
         pP = [-2:2:2]';
+    end
+    if nargin < 6
+        x0factor = 0.6;
+    end
+    if nargin < 7
+        dFactor = 0.01;
+    end
+    if nargin < 8
+        sFactor = 0.5;
     end
 
     % Initialize variables
@@ -26,10 +35,10 @@ function [positions, lengths, times, switches, active] = springSim(fps, t_stop, 
     cntS = size(sp,1);
 
     % Set up initial measurements for each spring
-    spC = ones(cntS,1) * 0.01;   % Damping
-    spK = ones(cntS,1) * 0.5;   % Stiffness
+    spC = ones(cntS,1) * dFactor;   % Damping
+    spK = ones(cntS,1) * sFactor;   % Stiffness
     spX0 = pP(sp);
-    spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.6;
+    spX0 = abs(spX0(:,1) - spX0(:,2)) * x0factor;
     spL = springLength(pP,sp);  % Length at last update
     spSteps = zeros(cntS,1);    % Number of steps since last update
 
@@ -56,6 +65,7 @@ function [positions, lengths, times, switches, active] = springSim(fps, t_stop, 
 
         % Position and velocity update
         [dp dv] = springStep(pP, pM, pV, sum(pF,2), dt, pA, sp, spSteps);
+
         % Detect collisions / switches in position
         l = springLength(pP + dp, sp);
         l2 = springLength(pP, sp);
@@ -63,9 +73,9 @@ function [positions, lengths, times, switches, active] = springSim(fps, t_stop, 
         if any(mask)
             disp('Switch in next iteration!');
             disp(time);
-            pSub = unique(sp(mask,:));
+            pSub = sp(mask,:);
             cnt = size(pSub,1);
-%            switches(end+1:end+cnt,:) = [repmat(time,[cnt 1]), pSub];
+            switches(end+1:end+cnt,:) = [repmat(time,[cnt 1]), pSub];
         end
 
         % Registration

### 2013-10-31 src/1D/run.m

@@ -7,31 +7,31 @@ clc;
 t_stop = 500;
 
 pP = [-2:1:2];
-[positions, ~, times, ~] = springSim(30, t_stop, pP);
-fi1 = figure();
+%[positions, ~, times, ~] = springSim(30, t_stop, pP);
+%fi1 = figure();
+%
+%plot(times, positions);
+%hold on;
+%axis([0 max(times) min(positions(:)) max(positions(:))]);
+%xlabel('Time');
+%ylabel('Position');
+%
+%
+%
+%%% Atomic:
+%fpses =      [10, 10, 30];
+%stepBounds = [100, 100, 1];
+%rFactors =   [inf,  0.2, inf];
+%results = cell(length(fpses),5);
+%
+%% Perform simulations
+%for i = 1:length(fpses)
+%    disp(fpses(i));
+%    [results{i,1}, results{i,2}, results{i,3}, results{i,4}, results{i,5}] = atomicSim(fpses(i),t_stop,stepBounds(i), rFactors(i));
+%end
+%
+%displaySim(results, fpses);
 
-plot(times, positions);
-hold on;
-axis([0 max(times) min(positions(:)) max(positions(:))]);
-xlabel('Time');
-ylabel('Position');
-
-
-return;
-
-%% Atomic:
-fpses =      [10, 10, 30];
-stepBounds = [100, 100, 1];
-rFactors =   [inf,  0.2, inf];
-results = cell(length(fpses),5);
-
-% Perform simulations
-for i = 1:length(fpses)
-    disp(fpses(i));
-    [results{i,1}, results{i,2}, results{i,3}, results{i,4}, results{i,5}] = atomicSim(fpses(i),t_stop,stepBounds(i), rFactors(i));
-end
-
-displaySim(results, fpses);
 
 %% Inverse
 fpses =      [10,30];
@@ -47,6 +47,7 @@ end
 results
 displaySim(results, fpses);
 
+return
 
 displaySim(results, fpses);
 

### 2013-10-31 src/1D/runAtomic.m

@@ -1,0 +1,74 @@
+clear all;
+close all;
+clc;
+impath = '../../images/';
+% Set up simulation parameters
+t_stop = 500;
+fpses =      [10, 10, 30];
+stepBounds = [60, 60, 1];
+rFactors =   [inf,  0.2, inf];
+results = cell(length(fpses),5);
+
+%% Uniform plots
+pP = [-2:2:2];
+[positions, ~, times, ~] = atomicSim(10, t_stop, 100, 0.1);
+fi1 = figure();
+
+plot(times, positions);
+hold on;
+axis([0 max(times) min(positions(:)) max(positions(:))]);
+xlabel('Time');
+ylabel('Position');
+title('AUH simulation (10 fps)');
+saveas(fi1, [impath 'atomic_uniform_10fps_0.png'])
+
+[positions, ~, times, ~] = atomicSim(10, t_stop, 1, inf);
+fi2 = figure();
+
+plot(times, positions);
+hold on;
+axis([0 max(times) min(positions(:)) max(positions(:))]);
+xlabel('Time');
+ylabel('Position');
+title('AUH simulation (10 fps)');
+saveas(fi2, [impath 'atomic_uniform_10fps_1.png'])
+
+%% Multi-scale plots
+t_stop = 250;
+pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
+[positions, ~, times, ~,~] = atomicSim(10, t_stop, 10, inf, pP, 0.8, 0.5,0.5);
+fi3 = figure();
+plot(times, positions);
+hold on
+axis([0 max(times) min(positions(:)) max(positions(:))]);
+xlabel('Time');
+ylabel('Position');
+title('AUH simulation (10 fps, sbound = 10)');
+saveas(fi3, [impath 'atomic_multiscale_10fps_10sbound.png'])
+
+[positions, ~, times, ~,~] = atomicSim(10, t_stop, 68, inf, pP, 0.8, 0.5, 0.5);
+fi4 = figure();
+plot(times, positions);
+hold on
+axis([0 max(times) min(positions(:)) max(positions(:))]);
+xlabel('Time');
+ylabel('Position');
+title('AUH simulation (10 fps, sbound = 68)');
+saveas(fi4, [impath 'atomic_multiscale_10fps_68sbound.png'])
+
+[positions, ~, times, switches,~] = atomicSim(10, t_stop, 69, inf, pP, 0.8, 0.5, 0.5);
+fi5 = figure();
[…]

### 2013-10-31 src/1D/runInverse.m

@@ -1,0 +1,59 @@
+clear all;
+close all;
+clc;
+impath = '../../images/';
+% Set up simulation parameters
+t_stop = 500;
+fpses =      [10, 10, 30];
+stepBounds = [60, 60, 1];
+rFactors =   [inf,  0.2, inf];
+results = cell(length(fpses),5);
+
+%% Uniform plots
+pP = [-2:2:2];
+[positions, ~, times, ~] = inverseSim(10, t_stop, 1, inf);
+fi1 = figure();
+
+plot(times, positions);
+hold on;
+axis([0 max(times) min(positions(:)) max(positions(:))]);
+xlabel('Time');
+ylabel('Position');
+title('AUH simulation (10 fps)');
+saveas(fi1, [impath 'inverse_uniform_10fps_0.png'])
+
+[positions, ~, times, ~] = inverseSim(30, t_stop, 1, inf);
+fi2 = figure();
+
+plot(times, positions);
+hold on;
+axis([0 max(times) min(positions(:)) max(positions(:))]);
+xlabel('Time');
+ylabel('Position');
+title('AUH simulation (10 fps)');
+saveas(fi2, [impath 'inverse_uniform_10fps_1.png'])
+
+return
+
+%% Multi-scale plots
+t_stop = 250;
+pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
+[positions, ~, times, ~,~] = inverseSim(10, t_stop, 10, inf, pP, 0.8, 0.5,0.5);
+fi3 = figure();
+plot(times, positions);
+hold on
+axis([0 max(times) min(positions(:)) max(positions(:))]);
+xlabel('Time');
+ylabel('Position');
+title('AUH simulation (10 fps, sbound = 10)');
+saveas(fi3, [impath 'inverse_multiscale_10fps_0.png'])
+
+[positions, ~, times, ~,~] = inverseSim(10, t_stop, 68, inf, pP, 0.8, 0.5, 0.5);
+fi4 = figure();
+plot(times, positions);
+hold on
+axis([0 max(times) min(positions(:)) max(positions(:))]);
+xlabel('Time');
+ylabel('Position');
+title('AUH simulation (10 fps, sbound = 68)');
+saveas(fi4, [impath 'inverse_multiscale_10fps_1.png'])

### 2013-10-31 src/1D/runReport.m

@@ -1,0 +1,11 @@
+clear all;
+close all;
+clc;
+
+runStandard;
+
+runSwitch;
+
+runAtomic;
+
+runInverse;

### 2013-10-31 src/1D/runStandard.m

@@ -1,5 +1,9 @@
+clear all;
+close all;
+clc;
+
 % Set up simulation parameters
-t_stop = 500;
+t_stop = 250;
 
 %% Uniform plots
 pP = [-2:1:2];

### 2013-10-31 src/1D/runSwitch.m

@@ -1,25 +1,43 @@
+clear all;
+close all;
+clc;
+
 % Set up simulation parameters
-t_stop = 500;
+t_stop = 250;
 impath = '../../images/';
 
 %% Multi-scale plots
 pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
-[positions, ~, times, ~] = switchSim(1, t_stop, true, pP);
+[positions, ~, times, switches] = switchSim(1, t_stop, true, pP);
 fi1 = figure();
 plot(times, positions);
 hold on
+for j = 1:size(switches,1)
+    time = switches(j,1);
+    ind1 = switches(j,2);
+    ind2 = switches(j,3);
+    plot(time,positions(ind1, times == time),'xr','MarkerSize', 10);
+    plot(time,positions(ind2, times == time),'xr','MarkerSize', 10);
+end
 axis([0 max(times) min(positions(:)) max(positions(:))]);
 xlabel('Time');
 ylabel('Position');
-title('Standard simulation (1 fps, relative)');
+title('Adapt-on-switch simulation (1 fps, relative)');
 saveas(fi1, [impath 'switch_multiscale_1fps_relative.png'])
 
-[positions, ~, times, ~] = switchSim(1, t_stop, false, pP);
+[positions, ~, times, switches] = switchSim(1, t_stop, false, pP);
 fi2 = figure();
 plot(times, positions);
 hold on
+for j = 1:size(switches,1)
+    time = switches(j,1);
+    ind1 = switches(j,2);
+    ind2 = switches(j,3);
+    plot(time,positions(ind1, times == time),'xr','MarkerSize', 10);
+    plot(time,positions(ind2, times == time),'xr','MarkerSize', 10);
+end
 axis([0 max(times) min(positions(:)) max(positions(:))]);
 xlabel('Time');
 ylabel('Position');
-title('Standard simulation (1 fps, absolute)');
+title('Adapt-on-switch simulation (1 fps, absolute)');
 saveas(fi2, [impath 'switch_multiscale_1fps_absolute.png'])

### 2013-11-03 report/article.tex
Added GEA section and code/images for loads of the report simulations

@@
 
-%\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
+%\usepackage[hmarginratio=1:1,top=32mm]{geometry} % Document margins
 %\usepackage{multicol} % Used for the two-column layout of the document
@@
 \usepackage{todonotes}
+\usepackage{graphicx}
 \usepackage{amsmath,amssymb,amsfonts}
@@
 \lettrine[nindent=0em,lines=3]{S} imulations of large hyperelastic materials
+such as brain tissue
 can be very time consuming. When working with materials of varying density
 of vertices/scale of mesh, the highest densities often give the most
-strict bounds on size of the time step of the simulation in order to keep
+strict bounds on size of the timestep of the simulation in order to keep
 the materials (and grids) stable and the discretization correct.
 
-In a ``standard'' simulation with a global time step size we need to compute
+In a ``standard'' simulation with a global timestep size we need to compute
 possibly unnecessarily many computations on the coarser parts of the material
@@
 areas with higher density of vertices, it would be an advantage to be able
-to perform large time steps for the majority of the material (the coarse
+to perform large timesteps for the majority of the material (the coarse
 parts) and smaller steps for the local patches where vertices could cause
 instability. We call the concept of varying the timestep local adaptive
-optimization of time step (LAOTS), in which ``local'' either refers to spatial
-locality (as described above) or time locality where the time step is global
+optimization of timestep (LAOT), in which ``local'' either refers to spatial
+locality (as described above) or time locality where the timestep is global
 for all vertices but varies during the simulation.
@@
 We will in this report
-study possible approaches for application of LAOTS. In order to keep the focus
-on possible LAOTS methods and remove unneccessary complexities, we will study
+study possible approaches for application of LAOT. In order to keep the focus
+on possible LAOT methods and remove unneccessary complexities, we will study
 the simple 1 dimensional simulation of a mass-spring system and try to apply
-LAOTS for it. Afterwards we will shortly discuss the application of our gained
+LAOT for it. Afterwards we will shortly discuss the application of our gained
 knowledge to the domain of hyperelastic material simulation and simulations in
@@
 
-\section{Previous work} According to \cite{Gander:2013} local time stepping
+\section{Previous work} According to \cite{Gander:2013} local timestepping
 was first studied in the community of ordinary differential equations (ODE)
@@
 computations either interpolation or extrapolation of the first computation
-is performed to compute the second part in each time step. The different
-time step sizes can be adapted to fit the model in each step as well. See
+is performed to compute the second part in each timestep. The different
+timestep sizes can be adapted to fit the model in each step as well. See
 \cite{Kvaernoe:1999} and \cite{Gear:1984} (similar strategy for linear multi
@@
 \cite{Gander:2013}. Generally the methods developed at first all make use
-of the same basic concepts for making the local adaptive time stepping:
+of the same basic concepts for making the local adaptive timestepping:
 Interpolation, extrapolation, prediction and correction, which we are going to
@@
 
-Recently newer and faster methods with local time stepping have been developed
+Recently newer and faster methods with local timestepping have been developed
 on the basis of nonlinear PDEs and the work performed in this field in
@@
 et al. \cite{Domingues:2008} is one of the more recent of these MR schemes.
-They describe a local scale-dependent time stepping for a space-adaptive multi
+They describe a local scale-dependent timestepping for a space-adaptive multi
 resolution scheme using the finite volume method in order to obtain speed-up
@@
 which is essentially the same motivation as this report. The method is based
-on an explicit Runge-Kutta method of second order. As expected the time step
+on an explicit Runge-Kutta method of second order. As expected the timestep
 size is imposed by a stability condition of the explicit Runge-Kutta on the
 finest scale, which increases with the scale of the mesh and hence we are able
-to increase the time step as well without violating the stability condition.
+to increase the timestep as well without violating the stability condition.
 
@@
 
-Second we will propose and test out three different methods for applying LAOTS
+Second we will propose and test out three different methods for applying LAOT
 to the simulation.
@@
 Third and last we will discussion the results and how we could apply our
-knowledge of LAOTS for hyperelastic material.
+knowledge of LAOT for hyperelastic material.
 
@@
 the difference in speed $\Delta u_v$ and position $\Delta p_v$ of each vertex
-$v$ using a given time step size $\Delta t$: \begin{align} \label{eq:delta_u}
+$v$ using a given timestep size $\Delta t$: \begin{align} \label{eq:delta_u}
 \Delta u_v &= \frac{\partial u_v}{\partial t} \Delta t = \frac{F_v}{m_v}
@@
 using the explicit Euler method and the next subsections describe the three
-different methods/modifications for applying LAOTS to the simulation just
-described. The three methods are using different ideas to implement LOATS. They
-could however possibly be combined in case they turn out to be working as
-intended to achieve the best possible LAOTS scheme.
+different methods/modifications for applying LAOT to the simulation just
+described. The three methods are using different ideas to implement LAOT. They
+all include using some sort of heuristics in order to perform the adaption of
+the methods. We developed and tested these heuristics as we developed the
+methods (chronologically) and hence most of the heuristics could potentially be
+used on all three methods and not only the methods on which they were used. In
+case the methods are working succesfully, they could furthermore be combined to
+generate ie. fully atomic timestepping (both time and spatial adaptive).
 
@@
 
-The heuristics can be chosen to fit the domain of the simulation, but we
-have chosen to use the number of steps since last update, $n_{last}$, and
-the ratio between the current spring length and the spring length at the
-last update $r_{last}$. This should force a bound on the maximal timestep
-before update as well as making sure updates are done more frequently in case
-the springlength is changing more rapidly. Further details are explained in
-section \ref{sec:experiments}.
+The heuristics can be chosen to fit the domain of the simulation, but we have
+chosen to use the number of steps since last update, $n_{last}$, and the ratio
+between the current spring length and the spring length at the last update
+$r_{last}$. This should force a bound on the maximal number of timesteps
+before update as well as force frequent updates when the spring is affected by a
+high acceleration in opposite direction of the current velocity/direction of
+change. \todo{Explain length ratio thought}. Further details are explained in
+section \ref{sec:experiments_atomic}.
 
-\subsection{Global estimated acceleration}
+\subsection{Global Estimated Acceleration}
 The third and last LAOT method we propose is time-adaptive. The simulations are
-done using an initial timestep size. We know that the force and hence
+done using an initial timestep size and define the timestep of each iteration
+based on a heuristic using the acceleration of the springs, and hence we call
+the method Global Estimated Acceleration (GEA). We know that the force and hence
 acceleration of a spring is somewhat linear close to the spring's rest length
@@
     \frac{1}{a} &= \frac{1}{\frac{-kx}{m}} = \frac{m}{-kx} \\
+    \label{eq:inverse}
     \Delta t &= \text{min}\left( \frac{1}{\abs{a} + 0.1} \right )
@@
 
-Our though is that for some more
+Our though is that for more
 complex simulations we are able to make estimates of this type less costly than
@@
 
-\section{Experiments and results}
+\section{Experiments, results and discussion}
 \label{sec:experiments}
 
-We first give an example of the actual application of LAOTS using the standard
-simulation of mass spring particles. Figure \ref{fig:std} shows two simulations
-of the same set of particles distributed with two different scales simulated for
-500 seconds each. All the springs $s \in S$ are initiated in a stretched state/position
-with a length 1.25 times longer than the rest length, damping and stiffness
-constants $C_s = 0.5, K_s = 0.5$, and all vertices $v \in V$
-have a mass of $m_v = 30$.
-\ref{fig:std_multi_30fps} shows a simulation using 30 fps and Figure
-\ref{fig:std_multi_1fps} shows a simulation using 1 fps. We clearly see that
-using a timestep of 30 fps the simulation is stable and able to handle the
-compression of all the small springs around $t = 110$ whereas a switch occurs
-when simulating at 1 fps. This switch causes ripples of sideeffects for the
-rest of the simulation and hence the final result is incorrect. No switches of
-the longer springs are however occuring and hence these pose no problem at 1
-fps. We will use this simulation for testing all three LAOTS methods proposed
-in section \ref{sec:methods}.
+We first give an example of the actual application of LAOT using the
+standard simulation of mass spring particles. Figure \ref{fig:std} shows two
+simulations of the same set of particles distributed with two different scales
+simulated for 250 seconds each. All the springs $s \in S$ are initiated in a
+stretched state/position with a length 1.25 times longer than the rest length,
+damping and stiffness constants $C_s = 0.5, K_s = 0.5$, and all vertices $v
+\in V$ have a mass of $m_v = 30$. Given this initial state all springs should
+start constracting since they are in a stretched position and eventually
+somewhere further in time stabilize at the rest position since we have added
+the damping force to the system.
+
+\ref{fig:std_multi_30fps} shows a simulation of the system described above
+using 30 fps and Figure \ref{fig:std_multi_1fps} shows the simulation using 1
+fps. We clearly see that using a timestep of 30 fps the simulation is stable
+and able to handle the compression of all the small springs around $t = 110$
+whereas a switch occurs when simulating at 1 fps. This switch causes ripples
+of sideeffects for the rest of the simulation and hence the final result is
+incorrect. No switches of the longer springs are however occuring and hence
+these pose no problem at 1 fps. We will use this simulation for testing all
+three LAOT methods proposed in section \ref{sec:methods}.
 
 \begin{figure}[H]
-    \begin{subfigure}{0.5\textwidth}
-        \includegraphics[width=\textwidth]{../images/standard_multiscale_30fps.png}
+    \begin{subfigure}[t]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{../images/standard_multiscale_30fps.pdf}
         \caption{30 fps}
@@
     \end{subfigure}
-    \begin{subfigure}{0.5\textwidth}
-        \includegraphics[width=\textwidth]{../images/standard_multiscale_1fps.png}
+    \begin{subfigure}[t]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{../images/standard_multiscale_1fps.pdf}
         \caption{1 fps}
@@
 
-\todo{Write something about the explicit euler based on uniform example}
 %\begin{figure}[H]
-%    \begin{subfigure}{0.5\textwidth}
-%        \includegraphics[width=\textwidth]{../images/standard_uniform_30fps.png}
+%    \begin{subfigure}[t]{0.5\textwidth}
+%        \includegraphics[width=\textwidth]{../images/standard_uniform_30fps.pdf}
 %        \caption{30 fps}
@@
 %    \end{subfigure}
-%    \begin{subfigure}{0.5\textwidth}
-%        \includegraphics[width=\textwidth]{../images/standard_uniform_1fps.png}
+%    \begin{subfigure}[t]{0.5\textwidth}
+%        \includegraphics[width=\textwidth]{../images/standard_uniform_1fps.pdf}
 %        \caption{1 fps}
@@
 \begin{figure}[H]
-    \begin{subfigure}{0.5\textwidth}
-        \includegraphics[width=\textwidth]{../images/switch_multiscale_1fps_relative.png}
+    \begin{subfigure}[t]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{../images/switch_multiscale_1fps_relative.pdf}
         \caption{Relative timestepping ($\Delta t = \frac{\Delta t}{100}$)}
@@
     \end{subfigure}
-    \begin{subfigure}{0.5\textwidth}
-        \includegraphics[width=\textwidth]{../images/switch_multiscale_1fps_absolute.png}
+    \begin{subfigure}[t]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{../images/switch_multiscale_1fps_absolute.pdf}
         \caption{Absolute timestepping ($\Delta t = \frac{1}{30}$)}
@@
 \subsection{Atomic Updating using Heuristics}
+\label{sec:experiments_atomic}
 The AUH method concists of two heuristics for determining whether an element
@@
 acceleration and velocity update to a position close in time to the next
-update position and then use this position when updating the acceleration and
+update position and then uses this position when updating the acceleration and
 velocity again. Had we chosen an alternative time integration method this
@@
 We now turn our attention towards the second heuristic.
-\todo{Decide on the exact heuristic to display after having talked to Sune.}
 
 \begin{figure}[H]
-    \begin{subfigure}{0.3\textwidth}
-        \includegraphics[width=\textwidth]{../images/atomic_multiscale_10fps_10sbound.png}
-        \caption{$l_{last} = 10$}
+    \begin{subfigure}[t]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{../images/atomic_multiscale_10fps_10sbound.pdf}
+        \caption{$l_{last} = 10$, $r_{last}$ ignored}
         \label{fig:atomic_multi_10sbound}
     \end{subfigure}
-    \begin{subfigure}{0.3\textwidth}
-        \includegraphics[width=\textwidth]{../images/atomic_multiscale_10fps_68sbound.png}
-        \caption{$l_{last} = 68$}
+    \begin{subfigure}[t]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{../images/atomic_multiscale_10fps_68sbound.pdf}
+        \caption{$l_{last} = 68$, $r_{last}$ ignored}
         \label{fig:atomic_multi_68sbound}
     \end{subfigure}
-    \begin{subfigure}{0.3\textwidth}
-        \includegraphics[width=\textwidth]{../images/atomic_multiscale_10fps_69sbound.png}
-        \caption{$l_{last} = 69$}
+    \begin{subfigure}[t]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{../images/atomic_multiscale_10fps_69sbound.pdf}
+        \caption{$l_{last} = 69$, $r_{last}$ ignored}
         \label{fig:atomic_multi_69sbound}
     \end{subfigure}
+    \begin{subfigure}[t]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{../images/atomic_multiscale_10fps_both.pdf}
+        \caption{$l_{last} = 10, r_{last} = 1.01$}
+        \label{fig:atomic_multi_both}
+    \end{subfigure}
     \caption{Simulations using the AUH method at 10 fps with various number of
-    steps since last update. Length heuristic is ignored (set to $\infty$).}
+    steps since last update.}
     \label{fig:atomic_multi_sbound}
@@
 
-\begin{figure}[H]
-    \begin{subfigure}{0.5\textwidth}
-        \includegraphics[width=\textwidth]{../images/atomic_uniform_10fps_0.png}
+\begin{figure}
+    \begin{subfigure}[t]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{../images/atomic_uniform_10fps_0.pdf}
         \caption{}
@@
     \end{subfigure}
-    \begin{subfigure}{0.5\textwidth}
-        \includegraphics[width=\textwidth]{../images/atomic_uniform_10fps_1.png}
+    \begin{subfigure}[t]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{../images/atomic_uniform_10fps_1.pdf}
         \caption{}
@@
     \end{subfigure}
+    \caption{}
+    \label{fig:atomic_uniform}
 \end{figure}
@@
 
-\subsection{Global estimated acceleration}
+\subsection{Global Estimated Acceleration}
+In the GEA method we perform adaption of the timestep according to equation
+\ref{eq:inverse}, and therefore we first wish to explore what this adaption
+means for a simple simulation before exploring the effect on the particle
+system used in the previous method experiments. We use the following
+constants: $d_s = 0.5$, $K_s = 0.3$, and $l_0,s$ is 1.25 times initial
+positions (2, 0, and -2). Furthermore given our simulation we define our
+timestep to $\Delta t_{GEA} = \Delta t^{0.8}\cdot \frac{1}{250}$ using the definition of
+$\Delta t$ in equation \ref{eq:inverse} in order to get steps of proper size.
+Figure \ref{fig:inverse_uniform_30fps_0} shows the resulting simulation
+displaying the position as a function of the time. The red lines show a standard
+simulation at 5 fps, the black lines show a standard simulation at 2.5 fps, and
+the blue lines show the GEA simulation. We observe that the three simulations a
+coinciding. Looking at figure \ref{fig:inverse_uniform_30fps_steps} we see the
+timestep sizes during the simulations displayed with the same colors as the
+simulation figure. We see that the GEA simulation varies it's timestep in
+between the two standard simulation timesteps and hence in this simulation the
+varying timestep isn't really beneficial as a standard simulation with lower fps
+is sufficient and hence less costly.
 
-The heuristic dampens the system
-\begin{figure}[H]
-    \begin{subfigure}{0.5\textwidth}
-        \includegraphics[width=\textwidth]{../images/inverse_uniform_10fps_0.png}
-        \caption{}
-        \label{fig:inverse_uniform_10fps_0}
+We now compare the GEA and standard simulation methods using the same simulation
+setup as we did when evaluating the previous methods. Here we have defined the varying
+timestep as $\Delta t_{GEA} = \Delta t^{0.8} \cdot \frac{1}{120}-0.038$ again
+using the definition of $\Delta t$ in equation \ref{eq:inverse}.
+Figure \ref{fig:inverse_multiscale} shows resulting plots of the timestep size
+as a function of the time of the simulation (figure
+\ref{fig:inverse_multiscale_steps}), and the accumulated number of timesteps as a
+function of the time of the simulation (figure
+\ref{fig:inverse_multiscale_cumstep}). The blue lines are the results for the GEA
+method and the red lines are the results for the standard method simulated at
+1.8 fps, which is the lowest fps at which no switches occur.
+Looking solely at the simulation-part of the previous methods (the first 250
+seconds) the GEA simulation used 455 steps while the standard simulation used 451 steps.
+If we however let the
+simulation run for a longer period of time, the GEA method surpasses the
+standard simulation. As we see in figure \ref{fig:inverse_multiscale_cumstep} the
+difference in accumulated number of timesteps rises as time passes and 20,000
+seconds into the simulation the standard simulation has used 36001 steps and GEA
+has used 30145 steps.
+This is caused by the damping of
+the system. As the acceleration of the springs decreases, the timesteps rise as can
+be seen in figure \ref{fig:inverse_uniform_30fps_steps} and
+\ref{fig:inverse_multiscale_steps}. By carefully scaling the
+timestep varying heuristic, the timestep adaption gains benefit over the
+standard simulation. The downside is however, that we first have to put an
+effort into scaling the heuristic to fit our simulation, but for the standard
+simulation we likewise have to experiment with the timestep size in order to
+find the threshold at which switches start occuring.
+
+Looking at the execution time the GEA method takes around 4.6 seconds and the
+standard simulation takes around 3.7 seconds to simulate 20,000 seconds at my
+personal laptop (2.53 GHz dualcore, 4GB memory) using my current Matlab 2013b
+implementation. This shows that the overhead of computing the heuristic and
+modifying the timestep for the GEA method in the current setup and simulation
+is more expensive than simply performing the standard simulation and hence we
+have gained no advantage. The execution time for the GEA method could however
+possibly be optimized and decreased using another programming language.
+Furthermore the advantage of the adapting is dependent of the simulation in
+question and hence further investigation of the possible use of the GEA method
+should be performed before a final conclusion can be made.
+
+Another heuristic which potentially could be used for the adaptive timestepping
+would be the acceleration (as opposed to the inverse of the acceleration in
+equation \ref{eq:inverse}). This could be done by having a basis timestep which
+is sufficient when the springs have a low acceleration. When the acceleration
+rises the timestep is decreased in stages instead of a direct one-to-one mapping
+depending on the order of magnitude of the acceleration.
+
+\begin{figure}
+    \begin{subfigure}[t]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{../images/inverse_uniform_30fps.pdf}
+        \caption{Simulation using GEA (blue) and standard (red) simulation methods with
+        position as a function of the time of the simulation.}
+        \label{fig:inverse_uniform_30fps_0}
     \end{subfigure}
-    \begin{subfigure}{0.5\textwidth}
-        \includegraphics[width=\textwidth]{../images/inverse_uniform_10fps_1.png}
-        \caption{}
-        \label{fig:inverse_uniform_10fps_1}
+    \begin{subfigure}[t]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{../images/inverse_uniform_30fps_steps.pdf}
+        \caption{Timestep size for GEA (blue) and standard (red) of the simulation in
+            \ref{fig:inverse_uniform_30fps_0}}
+        \label{fig:inverse_uniform_30fps_steps}
     \end{subfigure}
+    \caption{GEA compared with standard simulation on simple 3-vertex
+    mass-spring system}
+    \label{fig:inverse}
+\end{figure}
+\begin{figure}
+    \begin{subfigure}[t]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{../images/inverse_multiscale_steps.pdf}
+        \caption{Timestep size for GEA (blue) and standard (red) as a function
+        of the time passed}
+        \label{fig:inverse_multiscale_steps}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{../images/inverse_multiscale_cumstep.pdf}
+        \caption{Accumulated number of steps taken as a function of the time passed for GEA
+        (blue) and standard (red)}
+        \label{fig:inverse_multiscale_cumstep}
+    \end{subfigure}
+    \caption{GEA compared with standard simulation on particles distributed with
+        two scales simulated for 20,000 seconds}
+    \label{fig:inverse_multiscale}
 \end{figure}
@@
 
-\section{Discussion}
+\section{Future work}
 
-Adaptive timestepping based on highest acceleration/smallest grid
+Combine atomic approach with inverse acceleration heuristic
+
+Test methods with other integration schemes
+
+Test methods with hyperelastic materials
+
 Relate to hyperelastic materials (some linear heuristic = win)
+
+%------------------------------------------------
+
+\section{Conclusion}
 
@@
 %----------------------------------------------------------------------------------------
-
-\section{Conclusion}
 

### 2013-11-03 src/1D/atomicSim.m

@@ -57,7 +57,9 @@ function [positions, lengths, times, switches, active] = atomicSim(fps, t_stop, 
         [pA spA spL spSteps_new] = registerActive(pP, sp, spL, pV, time, dt,...
                                                  rFactor, spSteps, stepBound,...
                                                  spX0);
-        active(end+1) = length(spA);
+        pMask = zeros(1,cntP);
+        pMask(pA) = 1;
+        active(end+1,:) = pMask;
 
         % Force calculation of active springs/particles
         [pF dx] = springForces(pP, sp(spA,:), pV,...

### 2013-11-03 src/1D/inverseSim.m

@@ -1,8 +1,16 @@
-function [positions, lengths, times, switches, active] = inverseSim(fps, t_stop, stepBound, rFactor)
+function [positions, lengths, times, switches, active, dForce] = inverseSim(fps,t_stop, stepBound, rFactor,tScale, pP,x0factor)
     %% Initialize particles
-    % Create grid of particles at positions
-   % pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
-    pP = [-2:2:2]';
+    if nargin < 5
+        tScale = @(x) (x.^(0.8)).*1/250;
+    end
+    if nargin < 6
+        % Create grid of particles at positions
+        % pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
+        pP = [-2:2:2]';
+    end
+    if nargin < 7
+        x0factor = 0.3;
+    end
 
     % Initialize variables
     cntP = size(pP,1);
@@ -11,6 +19,7 @@ function [positions, lengths, times, switches, active] = inverseSim(fps, t_stop,
     times = [];
     switches = [];
     active = [];
+    dForce = [];
 
     % Set up initial measurements for each particle
     pV = zeros(cntP,1);         % Initial velocities
@@ -24,11 +33,10 @@ function [positions, lengths, times, switches, active] = inverseSim(fps, t_stop,
     cntS = size(sp,1);
 
     % Set up initial measurements for each spring
-    spC = ones(cntS,1) * 0.1;   % Damping
-    spK = ones(cntS,1) * 0.5;   % Stiffness
+    spC = ones(cntS,1) * 0.5;   % Damping
+    spK = ones(cntS,1) * x0factor;   % Stiffness
     spX0 = pP(sp);
-    spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.52;
-    spD = ones(cntS,1) * 0.1;   % Viscous drag
+    spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.8;
     spL = springLength(pP,sp);  % Length at last update
     spSteps = zeros(cntS,1);    % Number of steps since last update
 
@@ -43,45 +51,48 @@ function [positions, lengths, times, switches, active] = inverseSim(fps, t_stop,
     % Loop simulation over time
     while time <= t_stop + dt/2;
 
-        pP = pP + pV*dt;
         % Calculate spring timestep
-        [pA spA spL spSteps_new] = registerActive(pP, sp, spL, pV, time, dt,...
[…]

### 2013-11-03 src/1D/registerActive.m

@@ -11,14 +11,18 @@ lUpdate, velocities, time, dt, factor, steps, stepBound, rest)
     l = abs(springLength(positionsFull, springs));
 
     % Ratio between last updated length and next
-    ratio = abs(l-rest);
-%    ratio2 = abs(l ./ lUpdate);
+%    ratio = abs(l-rest);
+    ratio2 = abs(l ./ lUpdate);
 
     % Find indices of violating springs and points
     sMask = zeros(size(l));
-    sMask = ratio > factor;
-%    sMask = ratio > factor | ratio <= 1/factor;
-%    sMask = ratio2 > factor | ratio2 <= 1/factor | sMask;
+%    sMask = ratio > factor;
+%    sMask = ratio2 > factor | ratio2 <= 1/factor;
+    if ~isnan(factor)
+        sMask = ratio2 < factor & ratio2 >= 1/factor | sMask;
+    end
+%    sMask = ratio2 < factor | sMask;
+%    sMask = ratio2 >= 1/factor & sMask;
     if ~isnan(stepBound)
         sMask = sMask | steps >= stepBound;
     end

### 2013-11-03 src/1D/runAtomic.m

@@ -3,26 +3,30 @@ close all;
 clc;
 impath = '../../images/';
 % Set up simulation parameters
-t_stop = 500;
+t_stop = 250;
 fpses =      [10, 10, 30];
 stepBounds = [60, 60, 1];
-rFactors =   [inf,  0.2, inf];
+rFactors =   [inf,  1, inf];
 results = cell(length(fpses),5);
 
 %% Uniform plots
-pP = [-2:2:2];
-[positions, ~, times, ~] = atomicSim(10, t_stop, 100, 0.1);
+pP = [-2:2:2]';
+disp('With length');
+[positions, ~, times, ~, active] = atomicSim(10, t_stop, 20, 1.002, pP, 0.8);
 fi1 = figure();
-
 plot(times, positions);
 hold on;
+for i = 1:size(active,2)
+    updPoints = find(active(:,i));
+    plot(times(updPoints),positions(:,updPoints),'xg');
+end
 axis([0 max(times) min(positions(:)) max(positions(:))]);
 xlabel('Time');
 ylabel('Position');
 title('AUH simulation (10 fps)');
-saveas(fi1, [impath 'atomic_uniform_10fps_0.png'])
 
-[positions, ~, times, ~] = atomicSim(10, t_stop, 1, inf);
+disp('Without length');
+[positions, ~, times, ~] = atomicSim(10, t_stop, 1, NaN, pP, 0.8);
 fi2 = figure();
 
 plot(times, positions);
@@ -31,22 +35,20 @@ axis([0 max(times) min(positions(:)) max(positions(:))]);
 xlabel('Time');
 ylabel('Position');
 title('AUH simulation (10 fps)');
-saveas(fi2, [impath 'atomic_uniform_10fps_1.png'])
 
 %% Multi-scale plots
 t_stop = 250;
 pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
-[positions, ~, times, ~,~] = atomicSim(10, t_stop, 10, inf, pP, 0.8, 0.5,0.5);
+[positions, ~, times, ~, active] = atomicSim(10, t_stop, 10, NaN, pP, 0.8, 0.5,0.5);
 fi3 = figure();
 plot(times, positions);
-hold on
+hold on;
 axis([0 max(times) min(positions(:)) max(positions(:))]);
 xlabel('Time');
 ylabel('Position');
[…]

### 2013-11-03 src/1D/runInverse.m

@@ -3,7 +3,7 @@ close all;
 clc;
 impath = '../../images/';
 % Set up simulation parameters
-t_stop = 500;
+t_stop = 250;
 fpses =      [10, 10, 30];
 stepBounds = [60, 60, 1];
 rFactors =   [inf,  0.2, inf];
@@ -11,49 +11,118 @@ results = cell(length(fpses),5);
 
 %% Uniform plots
 pP = [-2:2:2];
-[positions, ~, times, ~] = inverseSim(10, t_stop, 1, inf);
+[positions, ~, times, ~, active, dForces] = inverseSim(2, t_stop, 1, inf);
+[positions2, ~, times2, ~, dForces2] = springSim(5, t_stop, pP, 0.3);
+[positions3, ~, times3, ~, dForces3] = springSim(2.5, t_stop, pP, 0.3);
 fi1 = figure();
 
-plot(times, positions);
+plot(times, positions(1,:), '-b','linewidth',2);
 hold on;
+plot(times2, positions2(1,:),'-r');
+plot(times3, positions3(1,:), '-k');
+plot(times, positions(2:end,:), '-b','linewidth',2);
+plot(times2, positions2(2:end,:), '-r');
+plot(times3, positions3(2:end,:), '-k');
+%updPoints = find(active);
+%plot(times(updPoints),positions(:,updPoints),'xg');
 axis([0 max(times) min(positions(:)) max(positions(:))]);
 xlabel('Time');
 ylabel('Position');
-title('AUH simulation (10 fps)');
-saveas(fi1, [impath 'inverse_uniform_10fps_0.png'])
+title('Simple simulation');
+legend('GEA','Standard (5 fps)','Standard (2.5 fps)','Location','best')
 
-[positions, ~, times, ~] = inverseSim(30, t_stop, 1, inf);
-fi2 = figure();
+%fi2 = figure();
+%plot(times2, dForces2,'-r');
+%hold on
+%plot(times, dForces, '-b');
+%title('Damping forces during simulation');
+%legend('Standard','GEA');
+%xlabel('Time');
+%ylabel('Sum of damping force');
+%scaleaxis([times,times2],[dForces,dForces2]);
 
-plot(times, positions);
+fi3 = figure();
+steps = times(2:end)-times(1:end-1);
+steps2 = times2(2:end)-times2(1:end-1);
[…]

### 2013-11-03 src/1D/runStandard.m

@@ -17,7 +17,6 @@ axis([0 max(times) min(positions(:)) max(positions(:))]);
 xlabel('Time');
 ylabel('Position');
 title('Standard simulation (30 fps)');
-saveas(fi1, [impath 'standard_uniform_30fps.png'])
 
 [positions, ~, times, ~] = springSim(1, t_stop, pP);
 fi2 = figure();
@@ -29,7 +28,6 @@ axis([0 max(times) min(positions(:)) max(positions(:))]);
 xlabel('Time');
 ylabel('Position');
 title('Standard simulation (1 fps)');
-saveas(fi2, [impath 'standard_uniform_1fps.png'])
 
 %% Multi-scale plots
 pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
@@ -41,7 +39,6 @@ axis([0 max(times) min(positions(:)) max(positions(:))]);
 xlabel('Time');
 ylabel('Position');
 title('Standard simulation (30 fps)');
-saveas(fi3, [impath 'standard_multiscale_30fps.png'])
 
 [positions, ~, times, ~] = springSim(1, t_stop, pP);
 fi4 = figure();
@@ -51,4 +48,12 @@ axis([0 max(times) min(positions(:)) max(positions(:))]);
 xlabel('Time');
 ylabel('Position');
 title('Standard simulation (1 fps)');
-saveas(fi4, [impath 'standard_multiscale_1fps.png'])
+
+fi = [fi1 fi2 fi3 fi4];
+set(fi, 'PaperPosition', [-0.2 -0.25 6 5.5]);
+set(fi, 'PaperSize', [5.4 5.2]);
+
+saveas(fi1, [impath 'standard_uniform_30fps'], 'pdf')
+saveas(fi2, [impath 'standard_uniform_1fps'], 'pdf')
+saveas(fi3, [impath 'standard_multiscale_30fps'], 'pdf')
+saveas(fi4, [impath 'standard_multiscale_1fps'], 'pdf')

### 2013-11-03 src/1D/runSwitch.m

@@ -23,7 +23,6 @@ axis([0 max(times) min(positions(:)) max(positions(:))]);
 xlabel('Time');
 ylabel('Position');
 title('Adapt-on-switch simulation (1 fps, relative)');
-saveas(fi1, [impath 'switch_multiscale_1fps_relative.png'])
 
 [positions, ~, times, switches] = switchSim(1, t_stop, false, pP);
 fi2 = figure();
@@ -40,4 +39,9 @@ axis([0 max(times) min(positions(:)) max(positions(:))]);
 xlabel('Time');
 ylabel('Position');
 title('Adapt-on-switch simulation (1 fps, absolute)');
-saveas(fi2, [impath 'switch_multiscale_1fps_absolute.png'])
+
+fi = [fi1 fi2];
+set(fi, 'PaperPosition', [-0.2 -0.25 6 5.5]);
+set(fi, 'PaperSize', [5.4 5.2]);
+saveas(fi1, [impath 'switch_multiscale_1fps_relative'],'pdf');
+saveas(fi2, [impath 'switch_multiscale_1fps_absolute'],'pdf');

### 2013-11-03 src/1D/scaleaxis.m

@@ -1,0 +1,8 @@
+function scaleaxis (xpoints, ypoints)
+    x_min = min(xpoints(:));
+    x_max = max(xpoints(:));
+    y_min = min(ypoints(:));
+    y_max = max(ypoints(:));
+
+    axis([x_min x_max y_min y_max]);
+end

### 2013-11-03 src/1D/springForces.m

@@ -1,4 +1,4 @@
-function [F, dx] = springForces(positions, springs, v, K, x0, C)
+function [F, dx, Fc] = springForces(positions, springs, v, K, x0, C)
     element = false;
     % Initialize internal variables
     cntS = size(springs,1);
@@ -38,13 +38,17 @@ function [F, dx] = springForces(positions, springs, v, K, x0, C)
         dx = springLength(positions, springs);
         l  = abs(dx);
 
-        Fs = K .* (l - x0);                             % Spring force
-        Fs = Fs + C .* (v(from) - v(to)) .* dx ./ l;    % Dampening
-        Fs = Fs .* (-dx ./ l);                          % Direction
+        Fs = -K .* (l - x0);                             % Spring force
+        Fc = -C .* (v(from) - v(to)) .* dx ./ l;
+%        Fs = Fs + C .* (v(from) - v(to)) .* dx ./ l;    % Dampening
+        Ft = Fs + Fc;
+        Ft = Ft .* (-dx ./ l);                          % Direction
 
         % Apply forces to endpoints
-        F(from,1) = F(from,1) + Fs;
-        F(to,2)   = F(to,2) - Fs;
+        F(from,1) = F(from,1) - Ft;
+        F(to,2)   = F(to,2) + Ft;
+        Fc = Fs;
+
 
     end
 

### 2013-11-03 src/1D/springSim.m

@@ -1,8 +1,11 @@
-function [positions, lengths, times, switches] = springSim(fps_lowest, t_stop, pP)
+function [positions, lengths, times, switches, dForces] = springSim(fps_lowest, t_stop, pP, sFactor)
     % Create grid of particles at positions
     if nargin < 3
         pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
         %pP = [-2:1:2];
+    end
+    if nargin < 4
+        sFactor = 0.5;
     end
     if size(pP,2) > size(pP,1)
         pP = pP';
@@ -29,7 +32,7 @@ function [positions, lengths, times, switches] = springSim(fps_lowest, t_stop, p
 
     % Set up initial measurements for each spring
     spC = ones(cntS,1) * 0.5;   % Damping
-    spK = ones(cntS,1) * 0.5;   % Stiffness
+    spK = ones(cntS,1) * sFactor;   % Stiffness
     spX0 = pP(sp);
     spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.8;
 
@@ -62,6 +65,7 @@ function [positions, lengths, times, switches] = springSim(fps_lowest, t_stop, p
     lengths = [];
     times = [];
     switches = [];
+    dForces = [];
 
     % subdivide particles
     pSub = [];
@@ -72,7 +76,7 @@ function [positions, lengths, times, switches] = springSim(fps_lowest, t_stop, p
     pF = zeros(cntP,2);
     while time <= t_stop + dt/2;
         % Force calculation
-        [pF dx] = springForces(pP, sp, pV, spK, spX0, spC);
+        [pF dx Fc] = springForces(pP, sp, pV, spK, spX0, spC);
 
         % Position and velocity update values
         [dp dv] = springStep(pP, pM, pV, sum(pF,2), dt);
@@ -81,6 +85,7 @@ function [positions, lengths, times, switches] = springSim(fps_lowest, t_stop, p
         positions(:,end+1) = pP;
         lengths(:,end+1) = -dx;
         times(end+1) = time;
+        dForces(:,end+1) = Fc;
 
         % Update position and velocity of simulation
         pP = pP + dp;

### 2013-11-03 src/1D/springTimestep.m

@@ -1,13 +1,13 @@
-function dt = springTimestep(positions, springs, m, K, x0)
+function dt = springTimestep(positions, springs, m, K, x0,Fk)
     % Get spring lengths
     dt = ones(size(positions,1),1) * -Inf;
     from = springs(:,1);
     to   = springs(:,2);
-    dx = abs(springLength(positions, springs));
-    divider = abs(-K .* (dx - x0))+0.1;
+%    dx = abs(springLength(positions, springs));
+%    divider = abs(-K .* (dx - x0))+0.1;
+    divider = abs(Fk)+0.1;
 
     dt(from) = max(dt(from), divider);
     dt(to) = max(dt(to), divider);
-    dt = (m ./ dt).^(0.1);
-%    dt(dt > 1) = 1;
+    dt = (m ./ dt);
 end

### 2013-11-04 report/article.tex
Small notes added and cleanup of code

@@
 positions (2, 0, and -2). Furthermore given our simulation we define our
-timestep to $\Delta t_{GEA} = \Delta t^{0.8}\cdot \frac{1}{250}$ using the definition of
-$\Delta t$ in equation \ref{eq:inverse} in order to get steps of proper size.
-Figure \ref{fig:inverse_uniform_30fps_0} shows the resulting simulation
-displaying the position as a function of the time. The red lines show a standard
-simulation at 5 fps, the black lines show a standard simulation at 2.5 fps, and
-the blue lines show the GEA simulation. We observe that the three simulations a
-coinciding. Looking at figure \ref{fig:inverse_uniform_30fps_steps} we see the
-timestep sizes during the simulations displayed with the same colors as the
-simulation figure. We see that the GEA simulation varies it's timestep in
-between the two standard simulation timesteps and hence in this simulation the
-varying timestep isn't really beneficial as a standard simulation with lower fps
-is sufficient and hence less costly.
+timestep to $\Delta t_{GEA} = \Delta t^{0.8}\cdot \frac{1}{250}$ using
+the definition of $\Delta t$ in equation \ref{eq:inverse} in order to get
+steps of proper size. Figure \ref{fig:inverse_uniform_30fps_0} shows the
+resulting simulation displaying the position as a function of the time.
+The red lines show a standard simulation at 5 fps, the black lines show a
+standard simulation at 2.5 fps, and the blue lines show the GEA simulation.
+We observe that the three simulations a coinciding. Looking at figure
+\ref{fig:inverse_uniform_30fps_steps} we see the timestep sizes during the
+simulations displayed with the same colors as the simulation figure. We see
+that the GEA simulation varies it's timestep in between the two standard
+simulation timesteps and hence in this simulation the varying timestep isn't
+really beneficial as a standard simulation with lower fps is sufficient and
+hence less costly.
 
-We now compare the GEA and standard simulation methods using the same simulation
-setup as we did when evaluating the previous methods. Here we have defined the varying
-timestep as $\Delta t_{GEA} = \Delta t^{0.8} \cdot \frac{1}{120}-0.038$ again
-using the definition of $\Delta t$ in equation \ref{eq:inverse}.
-Figure \ref{fig:inverse_multiscale} shows resulting plots of the timestep size
-as a function of the time of the simulation (figure
-\ref{fig:inverse_multiscale_steps}), and the accumulated number of timesteps as a
-function of the time of the simulation (figure
-\ref{fig:inverse_multiscale_cumstep}). The blue lines are the results for the GEA
-method and the red lines are the results for the standard method simulated at
-1.8 fps, which is the lowest fps at which no switches occur.
+We now compare the GEA and standard simulation methods using the same
+simulation setup as we did when evaluating the previous methods. Here we
+have defined the varying timestep as $\Delta t_{GEA} = \Delta t^{0.8}
+\cdot \frac{1}{120}-0.038$ again using the definition of $\Delta t$ in
+equation \ref{eq:inverse}. Figure \ref{fig:inverse_multiscale} shows
+resulting plots of the timestep size as a function of the time of the
+simulation (figure \ref{fig:inverse_multiscale_steps}), and the accumulated
+number of timesteps as a function of the time of the simulation (figure
+\ref{fig:inverse_multiscale_cumstep}). The blue lines are the results for
+the GEA method and the red lines are the results for the standard method
+simulated at 1.8 fps, which is the lowest fps at which no switches occur.
 Looking solely at the simulation-part of the previous methods (the first 250
-seconds) the GEA simulation used 455 steps while the standard simulation used 451 steps.
-If we however let the
-simulation run for a longer period of time, the GEA method surpasses the
-standard simulation. As we see in figure \ref{fig:inverse_multiscale_cumstep} the
-difference in accumulated number of timesteps rises as time passes and 20,000
-seconds into the simulation the standard simulation has used 36001 steps and GEA
-has used 30145 steps.
-This is caused by the damping of
-the system. As the acceleration of the springs decreases, the timesteps rise as can
-be seen in figure \ref{fig:inverse_uniform_30fps_steps} and
-\ref{fig:inverse_multiscale_steps}. By carefully scaling the
-timestep varying heuristic, the timestep adaption gains benefit over the
-standard simulation. The downside is however, that we first have to put an
-effort into scaling the heuristic to fit our simulation, but for the standard
-simulation we likewise have to experiment with the timestep size in order to
-find the threshold at which switches start occuring.
+seconds) the GEA simulation used 455 steps while the standard simulation
+used 451 steps. If we however let the simulation run for a longer period
+of time, the GEA method surpasses the standard simulation. As we see in
+figure \ref{fig:inverse_multiscale_cumstep} the difference in accumulated
+number of timesteps rises as time passes and 20,000 seconds into the
+simulation the standard simulation has used 36001 steps and GEA has used
+30145 steps. This is caused by the damping of the system. As the acceleration
+of the springs decreases, the timesteps rise as can be seen in figure
+\ref{fig:inverse_uniform_30fps_steps} and \ref{fig:inverse_multiscale_steps}.
+By carefully scaling the timestep varying heuristic, the timestep adaption
+gains benefit over the standard simulation. The downside is however, that we
+first have to put an effort into scaling the heuristic to fit our simulation,
+but for the standard simulation we likewise have to experiment with the
+timestep size in order to find the threshold at which switches start occuring.
 
@@
 
-Another heuristic which potentially could be used for the adaptive timestepping
-would be the acceleration (as opposed to the inverse of the acceleration in
-equation \ref{eq:inverse}). This could be done by having a basis timestep which
-is sufficient when the springs have a low acceleration. When the acceleration
-rises the timestep is decreased in stages instead of a direct one-to-one mapping
-depending on the order of magnitude of the acceleration.
+Another heuristic which potentially could be used for the adaptive
+timestepping would be the acceleration (as opposed to the inverse of the
+acceleration in equation \ref{eq:inverse}). This could be done by having a
+basis timestep which is sufficient when the springs have a low acceleration.
+When the acceleration rises the timestep is decreased in stages instead
+of a direct one-to-one mapping depending on the order of magnitude of the
+acceleration.
 
@@
 
-Relate to hyperelastic materials (some linear heuristic = win)
+Alternative acceleration heuristic (steps in order of magnitude)
 
@@
 \section{Conclusion}
+
+The two first of our methods didn't work. The last of our methods worked
+(GEA). Further work should be put into optimizing it and testing with
+hyperelastic materials.
 

### 2013-11-04 src/1D/atomic/atomicSim.m

@@ -1,83 +1,0 @@
-function [positions, lengths, times, switches, active] = springSim(fps, t_stop, stepBound, rFactor)
-    %% Initialize particles
-    % Create grid of particles at positions
-    %pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
-    pP = [-2:2:2]';
-
-    % Initialize variables
-    cntP = size(pP,1);
-    positions = [];
-    lengths = [];
-    times = [];
-    switches = [];
-    active = [];
-
-    % Set up initial measurements for each particle
-    pV = zeros(cntP,1);         % Initial velocities
-    pF = zeros(cntP,1);         % Force at time t
-    pM = ones(cntP,1) * 30;     % Mass of each particle
-
-    %% Register springs
-    % Springs between all inner particles (indices)
-    sp = [1:cntP]';
-    sp = [sp(1:end-1),sp(2:end)];
-    cntS = size(sp,1);
-
-    % Set up initial measurements for each spring
-    spC = ones(cntS,1) * 0.01;   % Damping
-    spK = ones(cntS,1) * 0.5;   % Stiffness
-    spX0 = pP(sp);
-    spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.6;
-    spL = springLength(pP,sp);  % Length at last update
-    spSteps = zeros(cntS,1);    % Number of steps since last update
-
-    % Initial time
-    dt_base = 1 / fps;
-    dt = dt_base;
-    time = 0;
-
-    dx = zeros(cntS,1);
-    pF = zeros(cntP,2);
-
-    % Loop simulation over time
-    while time <= t_stop + dt/2;
-
-        % Find active particles pA and active springs spA
-        [pA spA spL spSteps_new] = registerActive(pP, sp, spL, pV, time, dt,...
-                                                 rFactor, spSteps, stepBound,...
-                                                 spX0);
-        active(end+1) = length(spA);
-
-        % Force calculation of active springs/particles
-        [pF dx] = springForces(pP, sp(spA,:), pV,...
-                                   spK(spA), spX0(spA), spC(spA));
-
-        % Position and velocity update
-        [dp dv] = springStep(pP, pM, pV, sum(pF,2), dt, pA, sp, spSteps);
-        % Detect collisions / switches in position
-        l = springLength(pP + dp, sp);
-        l2 = springLength(pP, sp);
-        mask = l .* l2 < 0;
-        if any(mask)
[…]

### 2013-11-04 src/1D/atomic/registerActive.m

@@ -1,33 +1,0 @@
-function [pMask sMask lUpdate, steps] = registerActive(positions, springs,...
-lUpdate, velocities, time, dt, factor, steps, stepBound, rest)
-    if nargin < 9
-        stepBound = NaN;
-    end
-
-    % New positions using existing velocity
-    positionsFull = positions + velocities * dt;
-
-    % New length
-    l = abs(springLength(positionsFull, springs));
-
-    % Ratio between last updated length and next
-    ratio = abs(l-rest);
-%    ratio2 = abs(l ./ lUpdate);
-
-    % Find indices of violating springs and points
-    sMask = zeros(size(l));
-    sMask = ratio > factor;
-%    sMask = ratio > factor | ratio <= 1/factor;
-%    sMask = ratio2 > factor | ratio2 <= 1/factor | sMask;
-    if ~isnan(stepBound)
-        sMask = sMask | steps >= stepBound;
-    end
-    % Temporary override to debug the rest of the simulator
-    sMask = find(sMask > 0);
-    pMask = unique(springs(sMask,:));
-
-    % Update last updated lengths and number of steps since update
-    lUpdate(sMask) = l(sMask);
-    steps = steps+1;
-    steps(sMask) = 1;
-end

### 2013-11-04 src/1D/atomic/run.m

@@ -1,21 +1,0 @@
-% Clear environment
-clear all;
-close all;
-clc;
-
-% Set up simulation parameters
-t_stop = 500;
-fpses =      [10, 10, 30];
-stepBounds = [100, 100, 1];
-rFactors =   [inf,  0.2, inf];
-results = cell(length(fpses),5);
-
-% Perform simulations
-for i = 1:length(fpses)
-    disp(fpses(i));
-    [results{i,1}, results{i,2}, results{i,3}, results{i,4}, results{i,5}] = atomicSim(fpses(i),t_stop,stepBounds(i), rFactors(i));
-end
-
-displaySim(results, fpses);
-
-%diffSim(results, fpses);

### 2013-11-04 src/1D/atomic/springForces.m

@@ -1,50 +1,0 @@
-function [F, dx] = springForces(positions, springs, v, K, x0, C)
-    element = false;
-    % Initialize internal variables
-    cntS = size(springs,1);
-    cntP = size(positions,1);
-    Fs  = zeros(cntS,1);
-
-    % Initialize output 
-    F   = zeros(cntP,2);
-    dx = zeros(cntS,1);
-
-    % Calculate force F of each spring
-    if element
-        for s = 1:cntS
-            % Endpoints of spring
-            from = springs(s,1);
-            to = springs(s,2);
-
-            % Lengh of spring
-            dx(s) = springLength(positions, springs(s,:));
-            l = abs(dx(s));
-
-            % Force calculation
-            Fs = K(s) * (l - x0(s));                        % Spring force
-            Fs = Fs + C(s) * (v(from) - v(to)) * dx(s) / l; % Dampening
-            Fs = Fs * (- dx(s) / l);                        % Direction
-
-            % Apply forces to endpoints
-            F(from,1) = F(from,1) + Fs;
-            F(to,2) = F(to,2) - Fs;
-        end
-    else
-        % Matrix alternative to spring (element) wise computation
-        from = springs(:,1);
-        to   = springs(:,2);
-
-        dx = springLength(positions, springs);
-        l  = abs(dx);
-
-        Fs = K .* (l - x0);
-        Fs = Fs + C .* (v(from) - v(to)) .* dx ./ l;
-        Fs = Fs .* (-dx ./ l);
-
-        % Apply forces to endpoints
-        F(from,1) = F(from,1) + Fs;
-        F(to,2)   = F(to,2) - Fs;
-
-    end
-
-end

### 2013-11-04 src/1D/atomic/springLength.m

@@ -1,8 +1,0 @@
-function [l] = springLength(positions, springs)
-    % Get indices
-    left  = positions(springs(:,1));
-    right = positions(springs(:,2));
-
-    % Compute length (left - right)
-    l = left - right;
-end

### 2013-11-04 src/1D/atomic/springStep.m

@@ -1,14 +1,0 @@
-function [dp, dv] = springStep(p, m, v, F, dt, activeP, springs, steps)
-    cntA = size(activeP,1);
-    pSprings = zeros(cntA,1);
-    for i = 1:cntA
-        pSprings(i) = find(springs(:,1) == activeP(i) |...
-                           springs(:,2) == activeP(i),1);
-    end
-
-    % Position change
-    dp = v * dt;
-
-    % Velocity change
-    dv = F(activeP) ./ m(activeP) .* dt .* steps(pSprings);
-end

### 2013-11-04 src/1D/inverse/diffSim.m

@@ -1,18 +1,0 @@
-function diffSim(data, fps)
-    cntD = size(data,1);
-
-    if cntD > 1
-        figure();
-        for i = 1:cntD-1
-            subplot(1,cntD-1,i);
-            mult = fps(i) / fps(i+1);
-            diff = data{i,2}(:,1:mult:end) - data{i+1,2};
-            plot(data{i,3}(1:mult:end),diff,'-');
-            title(['Diff between ' num2str(fps(i)) ' and ' num2str(fps(i+1)) ' fps']);
-            xlabel('Time');
-            ylabel('Diff in spring length');
-        end
-    else
-        error('Not able to calculate diff of only 1 simulation');
-    end
-end

### 2013-11-04 src/1D/inverse/displaySim.m

@@ -1,59 +1,0 @@
-function displaySim(data, fps)
-    movie = false;
-    animation = false;
-    timesteps = true;
-
-    figure();
-    cntD = size(data,1);
-    for i = 1:cntD
-        i0 = i-1;
-        plot1 = i;
-        plot2 = plot1+cntD;
-        subplot(2, cntD, plot1);
-        plot(data{i,3},data{i,2});
-        title(['Spring lengths for ' num2str(fps(i)) ' fps']);
-        xlabel('Time (s)');
-        ylabel('Spring length');
-        axis([0 max(data{i,3}) -0.3 0.55]);
-        subplot(2, cntD, plot2);
-        plot(data{i,3},data{i,1});
-        title(['Simulation trace ' num2str(fps(i)) ' fps']);
-        xlabel('Time (s)');
-        ylabel('Position');
-        hold on;
-        for j = 1:size(data{i,4},1)
-            time = data{i,4}(j,1);
-            ind1 = data{i,4}(j,2);
-            ind2 = data{i,4}(j,3);
-            plot(time,data{i,1}(ind1, data{i,3} == time),'xr');
-            plot(time,data{i,1}(ind2, data{i,3} == time),'xr');
-        end
-        if timesteps
-            updPoints = find(data{i,5});
-            plot(data{i,3}(updPoints),data{i,1}(:,updPoints),'xb');
-            disp(length(updPoints) / max(data{i,3}));
-        end
-        axis([0 max(data{i,3}) -2 7]);
-    end
-
-%    if animation
-%        figure();
-%        y_values = zeros(size(positions,1),1);
-%        plot(positions(:,1),y_values,'x');
-%        set(gca,'NextPlot','replaceChildren');
-%        for i=1:length(times)
-%            plot(positions(:,i),y_values,'xr');
-%            hold on;
-%            title(['Time frame: ' num2str(times(i))]);
-%            %    drawnow;
-%            hold off;
-%            %    pause(0.1);
-%            F(i) = getframe;
-%        end
-%    end
-%
-%    if movie
-%        figure();
-%        movie(F,1,fps);
-%    end
-end

### 2013-11-04 src/1D/inverse/inverseSim.m

@@ -1,87 +1,0 @@
-function [positions, lengths, times, switches, active] = inverseSim(fps, t_stop, stepBound, rFactor)
-    %% Initialize particles
-    % Create grid of particles at positions
-   % pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
-    pP = [-2:2:2]';
-
-    % Initialize variables
-    cntP = size(pP,1);
-    positions = [];
-    lengths = [];
-    times = [];
-    switches = [];
-    active = [];
-
-    % Set up initial measurements for each particle
-    pV = zeros(cntP,1);         % Initial velocities
-    pF = zeros(cntP,1);         % Force at time t
-    pM = ones(cntP,1) * 30;     % Mass of each particle
-
-    %% Register springs
-    % Springs between all inner particles (indices)
-    sp = [1:cntP]';
-    sp = [sp(1:end-1),sp(2:end)];
-    cntS = size(sp,1);
-
-    % Set up initial measurements for each spring
-    spC = ones(cntS,1) * 0.1;   % Damping
-    spK = ones(cntS,1) * 0.5;   % Stiffness
-    spX0 = pP(sp);
-    spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.52;
-    spD = ones(cntS,1) * 0.1;   % Viscous drag
-    spL = springLength(pP,sp);  % Length at last update
-    spSteps = zeros(cntS,1);    % Number of steps since last update
-
-    % Initial time
-    dt_base = 1 / fps;
-    dt = dt_base;
-    time = 0;
-
-    dx = zeros(cntS,1);
-    pF = zeros(cntP,2);
-
-    % Loop simulation over time
-    while time <= t_stop + dt/2;
-
-        pP = pP + pV*dt;
-        % Calculate spring timestep
-        [pA spA spL spSteps_new] = registerActive(pP, sp, spL, pV, time, dt,...
-                                                 rFactor, spSteps, stepBound,...
-                                                 spX0);
-        active(end+1) = length(spA);
-
-        % Force calculation of active springs/particles
-        [pF dx] = springForces(pP, sp(spA,:), pV,...
-                                   spK(spA), spX0(spA), spC(spA));
-
-        % Position and velocity update
-        [dp dv] = springStep(pP, pM, pV, sum(pF,2), dt, pA, sp, spSteps);
-        % Detect collisions / switches in position
-        l = springLength(pP + dp, sp);
[…]

### 2013-11-04 src/1D/inverse/run.m

@@ -1,21 +1,0 @@
-% Clear environment
-clear all;
-close all;
-clc;
-
-% Set up simulation parameters
-t_stop = 500;
-fpses =      [10,30];
-stepBounds = [1, 1];
-rFactors =   [inf, inf];
-results = cell(length(fpses),5);
-
-% Perform simulations
-for i = 1:length(fpses)
-    disp(fpses(i));
-    [results{i,1}, results{i,2}, results{i,3}, results{i,4}, results{i,5}] = inverseSim(fpses(i),t_stop,stepBounds(i), rFactors(i));
-end
-results
-displaySim(results, fpses);
-
-%diffSim(results, fpses);

### 2013-11-04 src/1D/inverse/springTimestep.m

@@ -1,13 +1,0 @@
-function dt = springTimestep(positions, springs, m, K, x0)
-    % Get spring lengths
-    dt = ones(size(positions,1),1) * -Inf;
-    from = springs(:,1);
-    to   = springs(:,2);
-    dx = abs(springLength(positions, springs));
-    divider = abs(-K .* (dx - x0))+0.1;
-
-    dt(from) = max(dt(from), divider);
-    dt(to) = max(dt(to), divider);
-    dt = (m ./ dt).^(0.1);
-%    dt(dt > 1) = 1;
-end

### 2013-11-04 src/1D/switch/displaySim.m

@@ -1,55 +1,0 @@
-function displaySim(data, fps)
-    movie = false;
-    animation = false;
-
-    fi = figure();
-    cntD = size(data,1);
-    for i = 1:cntD
-        i0 = i-1;
-        plot1 = i;
-        plot2 = plot1+size(data,1);
-        subplot(2, cntD, plot1);
-        plot(data{i,3},data{i,2});
-        title(['Spring lengths for ' num2str(fps(i)) ' fps']);
-        xlabel('Time');
-        ylabel('Spring length');
-        axis([0 max(data{i,3}) -0.3 0.55]);
-        subplot(2, cntD, plot2);
-        plot(data{i,3},data{i,1});
-        title(['Simulation trace ' num2str(fps(i)) ' fps']);
-        xlabel('Time');
-        ylabel('Position');
-        hold on;
-        for j = 1:size(data{i,4},1)
-            time = data{i,4}(j,1);
-            ind1 = data{i,4}(j,2);
-            ind2 = data{i,4}(j,3);
-            plot(time,data{i,1}(ind1, data{i,3} == time),'xr');
-            plot(time,data{i,1}(ind2, data{i,3} == time),'xr');
-        end
-        axis([0 max(data{i,3}) -2 7]);
-    end
-
-    saveas(fi, '../../../images/spring_motivation.png');
-
-%    if animation
-%        figure();
-%        y_values = zeros(size(positions,1),1);
-%        plot(positions(:,1),y_values,'x');
-%        set(gca,'NextPlot','replaceChildren');
-%        for i=1:length(times)
-%            plot(positions(:,i),y_values,'xr');
-%            hold on;
-%            title(['Time frame: ' num2str(times(i))]);
-%            %    drawnow;
-%            hold off;
-%            %    pause(0.1);
-%            F(i) = getframe;
-%        end
-%    end
-%
-%    if movie
-%        figure();
-%        movie(F,1,fps);
-%    end
-end

### 2013-11-04 src/1D/switch/run.m

@@ -1,17 +1,0 @@
-% Clear environment
-clear all;
-close all;
-clc;
-
-% Set up simulation parameters
-t_stop = 500;
-fpses = [10, 2, 0.15, 30];
-results = cell(length(fpses),4);
-for i = 1:length(fpses)
-    disp(['Simulation: ' num2str(fpses(i))]);
-    [results{i,1}, results{i,2}, results{i,3}, results{i,4}] = springSim(fpses(i),t_stop);
-end
-
-displaySim(results, fpses);
-
-%diffSim(results, fpses);

### 2013-11-04 src/1D/switch/springLength.m

@@ -1,8 +1,0 @@
-function [l] = springLength(positions, spring)
-    % Get indices
-    left  = positions(spring(:,1));
-    right = positions(spring(:,2));
-
-    % Compute length (left - right)
-    l = left - right;
-end

### 2013-11-04 src/1D/switch/springSim.m

@@ -1,118 +1,0 @@
-function [positions, lengths, times, switches] = springSim(fps_lowest, t_stop)
-    % Create grid of particles at positions
-    pP = [-2:0.5:1.99, 2:0.15:2.99, 3:0.5:6.99]';
-    %pP = [-2:1:2];
-    if size(pP,2) > size(pP,1)
-        pP = pP';
-    end
-
-    % Initialize variables
-    cntP = size(pP,1);
-
-    % Set up initial measurements for each particle
-    pV = zeros(cntP,1);         % Initial velocities
-    pF = zeros(cntP,1);         % Force at time t
-    pM = ones(cntP,1) * 30;    % Mass of each particle
-
-    % Fix the left-most point
-    % Find left side
-    val = min(pP);
-    pFixed = find(pP == val);
-
-    % Register all springs
-    % Springs between all inner particles (indices)
-    sp = [1:cntP]';
-    sp = [sp(1:end-1),sp(2:end)];
-    cntS = size(sp,1);
-
-    % Set up initial measurements for each spring
-    spC = ones(cntS,1) * 0.5;   % Damping
-    spK = ones(cntS,1) * 0.5;   % Stiffness
-    spX0 = pP(sp);
-    spX0 = abs(spX0(:,1) - spX0(:,2)) * 0.785;
-
-    distance_show = false;
-    momentum_show = false;
-
-    t_start = 0;
-    if t_stop == 0
-        t_stop = 500;
-    end
-    if fps_lowest == 0
-        fps_lowest = 30;
-    end
-    disp(fps_lowest);
-    dt_base = 1 / fps_lowest;
-    dt = dt_base;
-
-    valMax = max(pP);
-    rind = find(pP == valMax);
-
-    % External force:
-    f_ext = 0;
-
-    % Initial time
-    time = 0;
-    explicit = true;
-    h = 1;
-
-    positions = [];
-    lengths = [];
-    times = [];
-    switches = [];
-
-    % subdivide particles
-    pSub = [];
-    pSt = [];
-    pSdt = [];
-
-    dx = zeros(cntS,1);
-    pF = zeros(cntP,2);
-    while time <= t_stop + dt/2;
-        p_last = pP;
-        pP = pP + pV * dt;
-
-        % Check for subdivision
-        if dt < dt_base && size(pSub,1) > 0
-
-            % Find springs to re-calculate force for
-            springs = sp(ismember(sp,pSub));
-        end
-
-        % Force calculation
[…]

### 2013-11-04 src/1D/switch/springStep.m

@@ -1,23 +1,0 @@
-function [dp, dv] = springStep(p, m, v, F, dt, activeP, springs, steps)
-    if nargin <= 6
-        % Calculate derivative of velocity
-        dv = F ./ m;
-
-        % Velocity change
-        dv = dv * dt;
-    else
-        cntA = size(activeP,1);
-        pSprings = zeros(cntA,1);
-        for i = 1:cntA
-            pSprings(i) = find(springs(:,1) == activeP(i) |...
-                               springs(:,2) == activeP(i),1);
-        end
-        % Velocity change
-        dv = F(activeP) ./ m(activeP) .* dt .* steps(pSprings);
-        
-    end
-
-    % Position change
-    dp = v * dt;
-
-end

### 2013-11-04 report/article.tex
Small changes from Lisbeths correction added

@@
 has two endpoints vertices $p_{f,s} \in V$ and $p_{t,s} \in V \setminus
-p_{f,s}$, a damping factor $d_s$, a stiffness constant $K_s$ and a rest length
+p_{f,s}$, a damping factor $C_s$, a stiffness constant $K_s$ and a rest length
 $l_{0,s}$. All these variables and constants combined describe the state
@@
 high acceleration in opposite direction of the current velocity/direction of
-change. \todo{Explain length ratio thought}. Further details are explained in
+change. \todo{Explain length ratio thought} Further details are explained in
 section \ref{sec:experiments_atomic}.

### 2013-11-05 report/contract.tex
Draft for final contract

@@
 
+\section{Problem Statement}
+Study the application of locally adaptive timestepping in 1 dimensional
+simulation of a mass spring system.
+
 \section{Learning Goals}
@@
 \begin{enumerate}
-    \item Implement local adaptive optimization of timesteps (LAOT)
+    \item Implement local adaptive optimization of timesteps in a 1
+dimensional mass-spring system
+    \item Develop and implement possible heuristics for use in LAOT schemes
 	\item Evaluate LAOT to minimize the risk of unwanted deformations in grids
         and minimize the computational cost.
-    \item Discuss how LAOT is performed in practice using existing 2nd order optimization methods.
+    \item Discuss the use and combination of LAOT schemes and heuristics based
+        on the evaluation of these
 \end{enumerate}
@@
 \section{Problem Description}
-
-\subsection{Problem Statement}
-Study the application of locally adaptive timestepping in 1 dimensional
-simulation of a mass spring system.
 
@@
 \subsection{Proposed solution}
-We wish to develop a method for locally adapting the timestep size in
-order to avoid curl ups/explosions while still maintaining the original
-timestep size of the rest of the simulation. In other words for each of
-the original timesteps we perform further simulations of problematic local
-areas of the grid with smaller timesteps.
+We wish to develop a method for locally adapting the timestep size (either
+spatial or in time) in
+order to avoid curl ups/explosions for minimizing the number of computations
+needed for a simulation. We do this by exploring possible methods and heuristics
+for LAOT application without affecting the outcome of a standard simulation.
 

### 2013-11-05 report/article.tex
Final report done

--- 2013-11-03 report/article.tex
@@
 \usepackage{titlesec} % Allows customization of titles
-\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
-\renewcommand\thesubsection{\Roman{subsection}} % Roman numerals for subsections
+%\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
+%\renewcommand\thesubsection{\Roman{subsection}} % Roman numerals for subsections
 \titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
@@
 \textsc{Malte Stær Nissen}\thanks{Thanks to my supervisor Sune Darkner for
-helping me realize this project}\\[2mm] % Your name
+helping and guiding me through this project}\\[2mm] % Your name
 \normalsize University of Copenhagen \\ % Your institution
@@
 We will in this report
-study possible approaches for application of LAOT. In order to keep the focus
+study three possible approaches/methods for application of LAOT and the
+heuristics included in these methods. In order to keep the focus
 on possible LAOT methods and remove unneccessary complexities, we will study
 the simple 1 dimensional simulation of a mass-spring system and try to apply
-LAOT for it. Afterwards we will shortly discuss the application of our gained
-knowledge to the domain of hyperelastic material simulation and simulations in
-general.
+LAOT for it. Afterwards we will test our methods and discuss on the results and
+the applied heuristics.
 
@@
 
-\section{Previous work} According to \cite{Gander:2013} local timestepping
+\section{Previous work}
+According to \cite{Gander:2013} local timestepping
 was first studied in the community of ordinary differential equations (ODE)
@@
 space-adaptive discretizations and refinements of these were developed, see
-\cite{Berger:1984}. Then multiple MR based schemes were developed. Domingues
+\cite{Berger:1984}. Then multiple schemes based on these MR schemes were
+developed. Domingues
 et al. \cite{Domingues:2008} is one of the more recent of these MR schemes.
@@
 resolution scheme using the finite volume method in order to obtain speed-up
-using larger time-steps without violating a defined stability constraints,
+using larger time-steps without violating defined stability constraints,
 which is essentially the same motivation as this report. The method is based
@@
 \label{sec:methods}
-
 With no prerequisites within the field of mass simulations we will first
@@
 
-Second we will propose and test out three different methods for applying LAOT
-to the simulation.
+Second we will propose three different methods and heuristics for applying
+LAOT to the simulation. These methods will be tested and the results will be
+discussed in section \ref{sec:experiments} and \ref{sec:future_work}.
 
-Third and last we will discussion the results and how we could apply our
-knowledge of LAOT for hyperelastic material.
+Both the standard 1D mass-spring system simulator and simulators for each
+of the three LAOT methods have been implemented as described in this section.
+All figures are results of the implemented simulators.
 
-\subsection{1D mass-spring system} Mass-spring systems are systems simulating
-a set of vertices interconnected by springs in various ways to obtain
-desirable physical properties. The motion of the springs and vertices are
-then simulated as a function of the time elapsed. Examples of application of
-mass-spring systems are simulation of cloth or hair. When performing simulations
-we want to avoid a switch of position of spring-connected vertices caused by the
-timestep size of the simulation, since this invalidates our discretization of
-the domain. Furthermore we want to perform as few timesteps as possible in order
-to get to the final simulation state by using as big timesteps as possible
-without affecting the outcome of the simulation.
+\subsection{1D mass-spring system}
+Mass-spring systems are systems simulating a set of vertices interconnected by
+springs in various ways to obtain desirable physical properties. The motion
+of the springs and vertices are simulated as a function of the time elapsed.
+Examples of application of mass-spring systems are simulation of cloth or
+hair. When performing simulations we want to avoid a switch of position of
+spring-connected vertices caused by the timestep size of the simulation, since
+this invalidates our discretization of the domain. Furthermore we want to reach
+the final simulation state having performed as few computations as possible.
+This is done by having a good balance between the computational load of each
+step and having as large steps as possible without affecting the simulation
+outcome.
 
-When simulating the mass-spring system we perform the following steps:
-\begin{enumerate} \item Initialize system \item Calculate forces \item
-Calculate derrivatives \item Update positions and velocities \item Update time
-\item Go to step 2 if the simulation has not reached it's maximum simulation
-time. \end{enumerate} We now describe the details of each of the steps of
-the simulation.
+When simulating the mass-spring system we perform the following six steps:
+\begin{enumerate}
+    \item Initialize system
+    \item Calculate forces
+    \item Calculate derrivatives
+    \item Update positions and velocities
+    \item Update time
+    \item Go to step 2 if the simulation has not reached it's maximum
+simulation time.
+\end{enumerate}
+We now go through the details of each of the steps of the simulation.
 
@@
 $S$ of springs. Each vertex $v \in V$ has a specified position $p_v$,
-mass $m_v$, force $F_v$ and velocity $u_v$, and each spring $s \in S$
+mass $m_v$, force $F_v$ and velocity $u_v$. Each spring $s \in S$
 has two endpoints vertices $p_{f,s} \in V$ and $p_{t,s} \in V \setminus
-p_{f,s}$, a damping factor $d_s$, a stiffness constant $K_s$ and a rest length
-$l_{0,s}$. All these variables and constants combined describe the state
-of the mass-spring system at each time $t$ during a simulation. All these
-variables are initialized in step 1.
+p_{f,s}$, a damping factor $C_s$, a stiffness constant $K_s$ and a rest length
+$l_{0,s}$. The set of all these variables and constants for both vertices and
+springs describes the state
+of the mass-spring system at each time $t$ during a simulation. The entire set
+of variables and constants is initialized in step 1 as well as a variable
+$t_{stop}$ defining the duration/stopping time of the simulation.
 
-The total force of a spring is the sum of the spring force $F_{h,d}$
+The total force of a spring $s \in S$ is the sum of the spring force $F_{h,s}$
 calculated by using Hooke's Law (see \cite[p.~439]{Young:2010})
-and the damping force $F_{C,d}$ using a simple damping method (see
-\cite[p.~457]{Young:2010}): \begin{align} \nonumber l_d &= p_{l,d} -
-p_{r,d} \\ \label{eq:hooke} F_{h,d} &= -k_d x_d = -k_d (l_d - l_{0,d}) \\
-\label{eq:damping} F_{C,d} &= -C_d \frac{\partial x_d}{ \partial t} = - C_d
-\frac{\left(v_{l,d} - v_{r,d}\right) l_d}{\abs{l_d}} \\ F_d &= F_{h,d} +
-F_{C,d} \end{align} \todo{Maybe add direction correction: $F_d \cdot (-l_d
-/ \abs{l_d})$} in which $p_{l,d}$ and $p_{r,d}$ are the positions of the
-left and right endpoints of the spring $d$ respectively, and $v_{l,d}$ and
-$v_{r,d}$ are the velocities of the left and right endpoint vertices of the
-spring $d$ respectively, and $F_d$ is the total force of spring $d$.
+and the damping force $F_{C,s}$ using a simple damping method (see
+\cite[p.~457]{Young:2010}):
+\begin{align}
+    \nonumber l_d &= p_{l,s} - p_{r,s} \\
+    \label{eq:hooke}
+    F_{h,s} &= -k_s x_s = -k_s (l_s - l_{0,s}) \\
+    \label{eq:damping}
+    F_{C,s} &= -C_s \frac{\partial x_s}{ \partial t} = - C_s
+        \frac{\left(v_{l,s} - v_{r,s}\right) l_s}{\abs{l_d}} \\
+    F_s &= \left( F_{h,s} + F_{C,s} \right) \frac{-l_s}{\abs{l_s}}
+\end{align}
+In which $p_{l,s}$ and $p_{r,s}$ are the positions of the
+left and right endpoints of the spring $s$ respectively, $v_{l,s}$ and
+$v_{r,s}$ are the velocities of the left and right endpoint vertices of the
+spring $s$ respectively, and $F_s$ is the total force of spring $s$. Multiplying
+the total spring force by $\frac{-l_s}{\abs{l_s}}$ allows us to correct for the
+direction of the spring getting the final direction-oriented total spring force
+$F_s$.
 
-When calculating the forces applied by each spring, we add the total spring
-force $F_d$ to the left endpoint of the spring and subtract $F_d$ from the
+When calculating the forces applied by each spring, we subtract the total spring
+force $F_d$ from the left endpoint of the spring and add $F_d$ to the
 right endpoint since both endpoints are affected equally in opposite direction
-by the spring $d$ connecting them. These spring forces are calculated in step
+by the spring $s$ connecting them. These spring forces are calculated in step
 2 of the simulation.
 
-We now know how to calculate the force of the spring system and apply it to
-the vertices. In order to calculate the derrivatives of the system (step 3) we
+We now know how to calculate the spring forces of the spring system and apply them to
+the vertices. In order to calculate the first order derrivatives of the system (step 3) we
 use Newton's second law of motion (see \cite[p.~112-115]{Young:2010}) to get
 the difference in speed $\Delta u_v$ and position $\Delta p_v$ of each vertex
-$v$ using a given timestep size $\Delta t$: \begin{align} \label{eq:delta_u}
+$v$ using a given timestep size $\Delta t$:
+\begin{align}
+    \label{eq:delta_u}
     \Delta u_v &= \frac{\partial u_v}{\partial t} \Delta t = \frac{F_v}{m_v}
-\Delta t \\ \label{eq:delta_p} \Delta p_v &= u \Delta t \end{align}
+\Delta t \\
+    \label{eq:delta_p} \Delta p_v &= u \Delta t
+\end{align}
 
 As mentioned in the beginning of this section we have chosen to use the
-explicit Euler method defined below in equation \ref{eq:explicit_euler}
-to calculate the time ``integration'' \cite{Flaherty} when stepping through
+explicit Euler method (see \cite{Flaherty}) defined below in equation \ref{eq:explicit_euler}
+to calculate the time ``integration'' when stepping through
 the simulation. Given a variable $x$ as a function of the time $t$, the
-variable $x$ at the time $t + \Delta t$ is defined as: \begin{align}
-\label{eq:explicit_euler} x(t+\Delta t) &= x(t) + \Delta t \frac{\partial
-x(t)}{\partial t} \end{align} Applying this method for the timestepping
-of our simulation we get the two following position and velocity update
-equations using equations \ref{eq:delta_u} and \ref{eq:delta_p}: \begin{align}
-p_v(t+\Delta t) &= p_v(t) + \Delta p_v \\ u_v(t+\Delta t) &= u_v(t) + \Delta
-u_v \end{align} We use these equations to perform step 4 of the simulation.
+variable $x$ at the time $t + \Delta t$ is defined as:
+\begin{align}
+    \label{eq:explicit_euler}
+    x(t+\Delta t) &= x(t) + \Delta t \frac{\partial x(t)}{\partial t}
+\end{align}
+Applying this method for the timestepping of our simulation we get the
+two following position and velocity update equations using equations
+\ref{eq:delta_u} and \ref{eq:delta_p}:
+\begin{align*}
+    p_v(t+\Delta t) &= p_v(t) + \Delta p_v \\
+    u_v(t+\Delta t) &= u_v(t) + \Delta u_v
+\end{align*}
+We use these equations to perform step 4 of the simulation.
 
-Step 5 is a simple update of the time: $t = t + \Delta t$ and step 6 describes
-itself.
+Step 5 is a simple update of the time $t = t + \Delta t$ and step 6 simply make
+us loop steps 2 to 5 until we reach the stopping time $t_{stop}$ of the
+simulation.
 
 We have now fully described the simple case of simulating a mass-spring system
-using the explicit Euler method and the next subsections describe the three
+using the explicit Euler method. The next subsections describe the three
 different methods/modifications for applying LAOT to the simulation just
-described. The three methods are using different ideas to implement LAOT. They
+described. The three methods all rely on different ideas to implement LAOT. They
 all include using some sort of heuristics in order to perform the adaption of
@@
 methods (chronologically) and hence most of the heuristics could potentially be
-used on all three methods and not only the methods on which they were used. In
-case the methods are working succesfully, they could furthermore be combined to
-generate ie. fully atomic timestepping (both time and spatial adaptive).
+used across all three methods and not only the methods on which they were used.
 
@@
 \label{sec:method_switch}
-The first LAOT method we propose is time-adaptive. We perform a simulation
-with an initial timestep size. In each step we try to predict whether a timestep
-of the same size using the current velocities will result in a switch of the
-left and right vertices of any of the springs. If this is the case, we should
-try to avoid the switch by decreasing the timestep of the current iteration,
-re-calculate the velocity and position changes and resume stepping of the
-simulation using the new timestep size. When switches aren't detected anymore
-we gradually increase the timestep of the simulation again. We call this
-method the adapt-on-switch method.
+The first LAOT method we propose is the time-adaptive ``Adapt-on-switch''
+method. We perform simulation with an initial timestep size. In each step
+we try to predict whether a timestep of the current size using the current
+velocities will result in a switch of the left and right vertices of any
+of the springs of the system. If this is the case, we should try to avoid the switch by
+decreasing the timestep of the current iteration, re-calculate the velocity
+and position changes and resume stepping of the simulation using the new
+timestep size. When switches aren't detected anymore we gradually increase the
+timestep of the simulation again.
 
-The switch of two endspoints connected by a spring $d$ can easily be
-calculated by checking the length ($p_{l,d} - p_{r,d}$) of the spring before
+The switch of two endspoints connected by a spring $s$ can easily be
+calculated by checking the length ($p_{l,s} - p_{r,s}$) of the spring before
 and after a position update. If the sign of the length differs, the two
@@
 
-The heuristics can be chosen to fit the domain of the simulation, but we have
-chosen to use the number of steps since last update, $n_{last}$, and the ratio
-between the current spring length and the spring length at the last update
-$r_{last}$. This should force a bound on the maximal number of timesteps
-before update as well as force frequent updates when the spring is affected by a
-high acceleration in opposite direction of the current velocity/direction of
-change. \todo{Explain length ratio thought}. Further details are explained in
-section \ref{sec:experiments_atomic}.
+The heuristics can be chosen to fit the domain of the simulation, but we
+have chosen to use the number of steps since last update, $n_{last}$, and
+the ratio between the current spring length and the spring length at the
+last update $r_{last}$. This should force a bound on the maximal number of
+timesteps before update as well as force frequent updates when the spring
+is affected by a high acceleration in opposite direction of the current
+velocity/direction of change. Looking at the mechanics of a spring we know
+that a large force/acceleration is the result of a stretched/contracted
+spring. At the ``endpoints'' of the oscillating spring movement the spring
+length change decreases slowly followed by a slow increase in the opposite
+direction until movement has been accelerated up again. The critical
+``endpoints'' of the spring movement should therefore be characterized by a
+low change in spring length. Using these observations we should be able to
+force updates when $r_{last}$ is low. Further details are explained in section
+\ref{sec:experiments_atomic}.
 
 \subsection{Global Estimated Acceleration}
-The third and last LAOT method we propose is time-adaptive. The simulations are
-done using an initial timestep size and define the timestep of each iteration
-based on a heuristic using the acceleration of the springs, and hence we call
-the method Global Estimated Acceleration (GEA). We know that the force and hence
-acceleration of a spring is somewhat linear close to the spring's rest length
-and it ``bends'' off as the spring is streched or contracted further away from
-the rest length. Based on this knowledge we will try to make a heuristic
-telling us about the needed timestep size at any given point during the
-simulation. The smallest needed timestep defines the global timestep for that
-given step.
+The third and last LAOT method we propose is time-adaptive. The simulations
+are done using an initial timestep size and define the timestep of each
+iteration based on a heuristic using the acceleration of the springs, and
+hence we call the method ``Global Estimated Acceleration'' (GEA). We know
+that the force and hence acceleration of a spring is somewhat linear close to
+the spring's rest length. Once the spring is streched or contracted away from
+the rest length the force ``bends'' off.
+Based on this knowledge we will try to make a heuristic telling us about the
+needed timestep size at any given point during the simulation. The smallest
+needed timestep of the system defines the global timestep for that given step.
 
@@
 We have added a small value (0.1) in order to avoid the timestep to increase
-exponentially towards infinity as $\abs{a}$ approach 0. Furthermore a scaling
+exponentially towards infinity as $\abs{a}$ approaches 0. Furthermore a scaling
 of the timestep could be needed depending on the scale of the simulation mesh.
@@
 
-Our though is that for more
-complex simulations we are able to make estimates of this type less costly than
-performing the actual simulations of a small fixed size timestep.
+Our though is that for more complex simulations we are able to make estimates
+of this type of heuristic less costly than performing the actual simulations
+of a small fixed size timestep.
 
@@
 
-\section{Experiments, results and discussion}
+\section{Experiments, results, and discussion hereof}
 \label{sec:experiments}
 
-We first give an example of the actual application of LAOT using the
-standard simulation of mass spring particles. Figure \ref{fig:std} shows two
-simulations of the same set of particles distributed with two different scales
-simulated for 250 seconds each. All the springs $s \in S$ are initiated in a
-stretched state/position with a length 1.25 times longer than the rest length,
-damping and stiffness constants $C_s = 0.5, K_s = 0.5$, and all vertices $v
-\in V$ have a mass of $m_v = 30$. Given this initial state all springs should
-start constracting since they are in a stretched position and eventually
-somewhere further in time stabilize at the rest position since we have added
-the damping force to the system.
+We first give an example of the problem of simulations using the standard
+mass-spring simulator leading to the need for application of LAOT. Figure
+\ref{fig:std} shows two simulations of the same set of 23 particles distributed
+with two different scales simulated for 250 seconds each. All the springs
+$s \in S$ are initiated in a stretched state/position with a length 1.25
+times longer than the rest length, damping constants $C_s = 0.5$, stiffness
+constants $K_s = 0.5$. All vertices $v \in V$ have a mass of $m_v = 30$. Given
+this initial state, all springs should start constracting since they are in a
+stretched position and eventually somewhere further in time stabilize at their
+rest positions since we have added a damping force to the system.
 
-\ref{fig:std_multi_30fps} shows a simulation of the system described above
-using 30 fps and Figure \ref{fig:std_multi_1fps} shows the simulation using 1
+Figure \ref{fig:std_multi_30fps} shows a simulation of the system described above
+using 30 fps and figure \ref{fig:std_multi_1fps} shows the simulation using 1
 fps. We clearly see that using a timestep of 30 fps the simulation is stable
-and able to handle the compression of all the small springs around $t = 110$
+and able to handle the compression of all the small springs around $t = 110$,
 whereas a switch occurs when simulating at 1 fps. This switch causes ripples
@@
 these pose no problem at 1 fps. We will use this simulation for testing all
-three LAOT methods proposed in section \ref{sec:methods}.
+three LAOT methods proposed in section \ref{sec:methods} and denote it the
+multi-scale simulation.
 
@@
 
-%\begin{figure}[H]
-%    \begin{subfigure}[t]{0.5\textwidth}
-%        \includegraphics[width=\textwidth]{../images/standard_uniform_30fps.pdf}
-%        \caption{30 fps}
-%        \label{fig:std_uni_30fps}
-%    \end{subfigure}
-%    \begin{subfigure}[t]{0.5\textwidth}
-%        \includegraphics[width=\textwidth]{../images/standard_uniform_1fps.pdf}
-%        \caption{1 fps}
-%        \label{fig:std_uni_1fps}
-%    \end{subfigure}
-%    \caption{Standard simulations of particles distributed uniformly using
-%    different timestep sizes}
-%\end{figure}
-
 \subsection{Adapt-on-switch}
-As mentioned in section \ref{sec:method_switch} the adapt-on-switch method
+\label{sec:experiment_switch}
+As mentioned in section \ref{sec:method_switch} the Adapt-on-switch method
 can be implemented using either a relative or a absolute adaption of the
@@
 the absolute solution to set the timestep to $\frac{1}{30}$ when switches are
-detected since this generated nice simulations for the standard simulation .
+detected since this generated nice simulations for the standard simulation.
 Furthermore the up-scaling for the relative scheme is set to a factor of 10
@@
 
-Figure \ref{fig:switch} shows the result of running simulations
-described earlier at 1 fps using the adapt-on-switch scheme, where
-the red crosses mark timesteps where a switch is predicted in the
+Figure \ref{fig:switch} shows the result of running the multi-scale
+simulation at 1 fps using the Adapt-on-switch scheme, where the
+red crosses mark timesteps at which a switch is predicted in the
 following step and hence the schemes are ``activated''. Figure
@@
 absolute scheme. As we clearly see in both figures the schemes fail to correct
-the simulations in spite of the switch prediction. Once the first switch
-occurs the simulation will change behavior and hence the first red cross marks
-the point at which the simulation can be rejected from being correct.
-
-After having made these observations we furthermore tested out a third way of
-creating the scheme by predicting how long a timestep we could take before the
-switch would occur and then divide this steplength by 2 to get a new timestep
-length. This generated infeasibly small timesteps forcing us to stop the
-simulation too early.
+the simulations in spite of the switch prediction. Once the first switch of
+each of the simulations occurs the simulations will change behavior and hence
+the first red cross marks the point at which the simulation can be rejected
+from being correct.
 
 Taking these observations into account we can safely say that the
-adapt-on-switch method/scheme is not working for our simulation. This does
+Adapt-on-switch method/scheme is not working for our multi-scale simulation. This does
 however not mean that the method couldn't work for other domains or other
 refinements of the method. The problem with the scheme is that the prediction
-of switches is done too late in the simulation and hence once we detect
-the switches it's too late to do anything about them given our model of
-the mass-spring system. In order for the scheme to work properly it should
-therefore be modified to be able to perform the detection of switches earlier
-in the simulation.
+of switches is done too late in the simulation. Once we detect the switches
+it is too late to do anything about them given our model of the mass-spring
+system. In order for the scheme to work properly it should be modified to be
+able to perform the detection of switches earlier in the simulation if such a
+prediction can be made reliably.
 
 In the real world the spring force would ``bend'' towards $\pm \infty$ once
-the spring is stretched or contracted too much and hence the linear model that
-we use doesn't suffice when using the adapt-on-switch scheme. Having used a
+the spring is stretched or contracted too much resulting in a maximum and
+minimum spring length. In other words the linear model that we use does not
+suffice when using the Adapt-on-switch scheme, since the spring forces aren't
+reacting the way we need them to when a spring is contracted. Having used a
 more realistic choice of model could possibly have had a positive impact of
-the use of this scheme since this would cause the spring force of the spring
-in question to grow in order to avoid the switch. \todo{Maybe the section
-above should be moved to a discussion section.}
+the use of this scheme. This would cause the spring force of the spring
+in question to grow in order to avoid the switch.
+
+As a sidenote it should be mentioned that we very shortly tested out an
+alternative way of calculating the adaptive stepsize to confirm the hypothesis
+above. We predicted how long a timestep we could take before the switch would
+occur and then divided this steplength by 2 to get a new timestep length. This
+generated infeasibly small timesteps forcing us to stop the simulation before
+the time limit $t_{stop}$ was reached. This supports our hypothesis of the
+incompatibility between the domain and the Adapt-on-switch method in the current
+state.
 
@@
 The AUH method concists of two heuristics for determining whether an element
-should be updated or not: steps since last update $l_{last}$, and the ratio
-between the spring length in last update and the current length $r_{last}$.
+should be updated or not: steps since last update $n_{last}$, and the ratio
+$r_{last}$ between the spring length in last update and the current length.
 
-We first test the method using only the first heuristic which will give
-us synchronized updates of all vertices every $r_{last}$ step. Using the
-same setup as the standard simulation we perform simulations using the AUH
-method with the length ratio heuristic ignored by setting the limit of
-the ratio to infinity. We set the timestep to $\Delta t = \frac{1}{10}$
-(10 fps) and vary $l_{last}$. Figure \ref{fig:atomic_multi_sbound} shows
-the resulting simulations where figures \ref{fig:atomic_multi_10sbound},
-\ref{fig:atomic_multi_68sbound}, and \ref{fig:atomic_multi_69sbound} uses
-an $l_{last}$ of 10, 68, and 69 respectively.
+We first test the method using only the first heuristic which will give us
+synchronized updates of all vertices every $r_{last}$. step. Using the same
+setup as the multi-scale simulation we perform simulations using the AUH method
+with the length ratio heuristic ignored, the timestep set to $\Delta t =
+\frac{1}{10}$ (10 fps) and varing $n_{last}$ across three different simulations.
+Figure \ref{fig:atomic_multi_sbound} (\ref{fig:atomic_multi_10sbound} to
+\ref{fig:atomic_multi_69sbound}) shows the resulting simulations where
+figures \ref{fig:atomic_multi_10sbound}, \ref{fig:atomic_multi_68sbound},
+and \ref{fig:atomic_multi_69sbound} uses a $n_{last}$ of 10, 68, and 69
+respectively.
 
-Simulating using 10 fps and $l_{last} = 10$ gives us the same amount
-of velocity updates as the standard simulation using 1 fps (which had
-switches), but gives us a nice and smooth simulation with no problems.
-As can be seen in the figures \ref{fig:atomic_multi_68sbound} and
+Simulating using 10 fps and $n_{last} = 10$ give us the same amount of
+velocity updates as the standard simulation using 1 fps (which had switches).
+Using the AUH scheme we do however get a nice and smooth simulation with no
+problems. As can be seen in the figures \ref{fig:atomic_multi_68sbound} and
 \ref{fig:atomic_multi_69sbound} we are able to perform the simulation using
-$l_{last} = 68$ without having switches occuring whereas using $l_{last} =
+$n_{last} = 68$ without having switches occuring whereas using $n_{last} =
 69$ results in switches occuring. In other words we are able to perform the
@@
 calculations. This is a result of combining the explicit Euler method with
-the AUH scheme since the scheme simply extrapolates the position at the last
+the AUH scheme. The scheme extrapolates the position at the last
 acceleration and velocity update to a position close in time to the next
-update position and then uses this position when updating the acceleration and
-velocity again. Had we chosen an alternative time integration method this
+update position. It then uses this position when updating the acceleration
+and velocity again. Had we chosen an alternative time integration method this
 advantage would most likely have been eliminated and let us perform the larger
-timesteps with the standard simulation.
+timesteps with the standard simulation. This would however had increased the
+computational cost of each timestep.
 
-We now turn our attention towards the second heuristic.
+The second heuristic ($r_{last}$) should be activated when $r_{last}$ is
+small. The bound $r_{b}$ for forcing updates based on $r_{last}$ should
+however be defined according to the simulation in question. Likewise the
+first heuristic should be activated in between the activation of the $r_b$
+bound to force a minimum of updates between the endpoints of the spring
+motion.
+
+In order to test out the heuristic we first perform a simulation using
+only 3 equally distributed particles at 10 fps for 250 seconds. The initial
+state of the simulation is: timestep $\Delta t = 0.1$, damping $C_s = 0.01$,
+stiffness $K_s = 0.5$, initial vertex positions 2, 0 and -2, and spring rest
+length 1.6 for both springs. By performing a series of simulations we find
+the most optimal combination of the two heuristics for the simulation to be
+$r_{last} = 20$, $r_b = 1.002$. Figure \ref{fig:atomic_uniform} shows the
+resulting simulation plots displaying positions as a function of the time
+using the AUH method (\ref{fig:atomic_uniform_10fps_0}) and the standard
+method (\ref{fig:atomic_uniform_10fps_1}). The green dots of the AUH plot
+shows the update positions.
+
+As we see from the AUH plot, the heuristics start out by updating the
+accelerations and velocities at the endpoints of the spring motions as
+intended. Likewise the stepping bound activates updates at lower frequency
+(every 20th step) in between the endpoints as intended. When time passes the
+system is being dampened too much compared to the small damping factor. At the
+end of the simulation the $r_b$ bound activates updates in all steps of the
+simulation. When we compare the resulting simulation of the AUH and standard
+methods we are observing quite large differences. As mentioned the AUH method
+dampens the system too much when the $r_b$ bound is not activated, and hence we
+get the observed difference in simulation result.
+
+Figure \ref{fig:atomic_multi_both} shows the resulting simulation plot
+using the AUH method and setting $n_{last} = 10,~r_b = 1.01$ for the
+multi-scale simulation. As for the previous simulation example we have
+estimated the best values of $n_{last}$ and $r_b$ by varying both and
+inspecting the resulting simulation. A switch occurs at time $t = 104$ which
+corrupts and invalidates the rest of the simulation.
+
+The conclusion of the application of this method is, that the first heuristic
+was very succesful in case we need to use the explicit Euler method. The second
+heuristic was however only affecting the outcome of the simulations negatively,
+and thus it should be rejected.
 
@@
         \includegraphics[width=\textwidth]{../images/atomic_multiscale_10fps_10sbound.pdf}
-        \caption{$l_{last} = 10$, $r_{last}$ ignored}
+        \caption{$n_{last} = 10$, $r_{last}$ ignored}
         \label{fig:atomic_multi_10sbound}
@@
         \includegraphics[width=\textwidth]{../images/atomic_multiscale_10fps_68sbound.pdf}
-        \caption{$l_{last} = 68$, $r_{last}$ ignored}
+        \caption{$n_{last} = 68$, $r_{last}$ ignored}
         \label{fig:atomic_multi_68sbound}
@@
         \includegraphics[width=\textwidth]{../images/atomic_multiscale_10fps_69sbound.pdf}
-        \caption{$l_{last} = 69$, $r_{last}$ ignored}
+        \caption{$n_{last} = 69$, $r_{last}$ ignored}
         \label{fig:atomic_multi_69sbound}
@@
         \includegraphics[width=\textwidth]{../images/atomic_multiscale_10fps_both.pdf}
-        \caption{$l_{last} = 10, r_{last} = 1.01$}
+        \caption{$n_{last} = 10, r_{b} = 1.01$}
         \label{fig:atomic_multi_both}
@@
         \includegraphics[width=\textwidth]{../images/atomic_uniform_10fps_0.pdf}
-        \caption{}
+        \caption{AUH method ($r_b = 1.002,~r_{last} = 20$)}
         \label{fig:atomic_uniform_10fps_0}
@@
         \includegraphics[width=\textwidth]{../images/atomic_uniform_10fps_1.pdf}
-        \caption{}
+        \caption{Standard method}
         \label{fig:atomic_uniform_10fps_1}
     \end{subfigure}
-    \caption{}
+    \caption{Simple 3-vertex simulations at 10 fps}
     \label{fig:atomic_uniform}
@@
 
-if ratio then too much dampening caused by lack of updates: Example: Simple with
-3 springs
+\subsection{Global Estimated Acceleration}
+\label{sec:experiments_inverse}
+In the GEA method we perform adaption of the timestep according to equation
+\ref{eq:inverse}. We therefore first wish to explore what this adaption
+means for a simple simulation, before exploring the effect on the particle
+system used in the previous method experiments. We use the following
+constants: $C_s = 0.5$, $K_s = 0.3$, $l_{0,s} = 1.6$, and the initial vertex
+positions are 2, 0, and -2. Furthermore given our simulation we define our
+timestep to $\Delta t_{GEA} = \Delta t^{0.8}\cdot \frac{1}{250}$ using
+the definition of $\Delta t$ in equation \ref{eq:inverse} in order to get
+steps of proper size. Figure \ref{fig:inverse_uniform_30fps_0} shows the
+resulting simulation displaying the position as a function of the time.
+The red lines show a standard simulation at 5 fps, the black lines show a
+standard simulation at 2.5 fps, and the blue lines show the GEA simulation.
+We observe that the three simulations are coinciding and hence our adaption of
+the timestep has no negative impact of the simulation outcome. Looking at figure
+\ref{fig:inverse_uniform_30fps_steps} we see the timestep sizes during the
+simulations displayed with the same colors as the simulation figure. We see
+that the GEA simulation varies it's timestep in between the two standard
+simulation timesteps and hence in this simulation the varying timestep isn't
+really beneficial as a standard simulation with lower fps is sufficient and
+hence less costly.
 
-\subsection{Global Estimated Acceleration}
-In the GEA method we perform adaption of the timestep according to equation
-\ref{eq:inverse}, and therefore we first wish to explore what this adaption
-means for a simple simulation before exploring the effect on the particle
-system used in the previous method experiments. We use the following
-constants: $d_s = 0.5$, $K_s = 0.3$, and $l_0,s$ is 1.25 times initial
-positions (2, 0, and -2). Furthermore given our simulation we define our
-timestep to $\Delta t_{GEA} = \Delta t^{0.8}\cdot \frac{1}{250}$ using the definition of
-$\Delta t$ in equation \ref{eq:inverse} in order to get steps of proper size.
-Figure \ref{fig:inverse_uniform_30fps_0} shows the resulting simulation
-displaying the position as a function of the time. The red lines show a standard
-simulation at 5 fps, the black lines show a standard simulation at 2.5 fps, and
-the blue lines show the GEA simulation. We observe that the three simulations a
-coinciding. Looking at figure \ref{fig:inverse_uniform_30fps_steps} we see the
-timestep sizes during the simulations displayed with the same colors as the
-simulation figure. We see that the GEA simulation varies it's timestep in
-between the two standard simulation timesteps and hence in this simulation the
-varying timestep isn't really beneficial as a standard simulation with lower fps
-is sufficient and hence less costly.
-
-We now compare the GEA and standard simulation methods using the same simulation
-setup as we did when evaluating the previous methods. Here we have defined the varying
-timestep as $\Delta t_{GEA} = \Delta t^{0.8} \cdot \frac{1}{120}-0.038$ again
+We now compare the GEA and standard simulation methods using the same
+multi-scale simulation setup. Here we have defined the varying timestep
+as $\Delta t_{GEA} = \Delta t^{0.8} \cdot \frac{1}{120}-0.038$ again
 using the definition of $\Delta t$ in equation \ref{eq:inverse}.
-Figure \ref{fig:inverse_multiscale} shows resulting plots of the timestep size
-as a function of the time of the simulation (figure
-\ref{fig:inverse_multiscale_steps}), and the accumulated number of timesteps as a
-function of the time of the simulation (figure
-\ref{fig:inverse_multiscale_cumstep}). The blue lines are the results for the GEA
-method and the red lines are the results for the standard method simulated at
-1.8 fps, which is the lowest fps at which no switches occur.
+Figure \ref{fig:inverse_multiscale} shows resulting plots of the
+timestep size as a function of the time of the simulation (figure
+\ref{fig:inverse_multiscale_steps}), and the accumulated number
+of timesteps as a function of the time of the simulation (figure
+\ref{fig:inverse_multiscale_cumstep}). The blue lines are the results for
+the GEA method and the red lines are the results for the standard method
+simulated at 1.8 fps, which is the lowest fps at which no switches occur.
 Looking solely at the simulation-part of the previous methods (the first 250
-seconds) the GEA simulation used 455 steps while the standard simulation used 451 steps.
-If we however let the
-simulation run for a longer period of time, the GEA method surpasses the
-standard simulation. As we see in figure \ref{fig:inverse_multiscale_cumstep} the
-difference in accumulated number of timesteps rises as time passes and 20,000
-seconds into the simulation the standard simulation has used 36001 steps and GEA
-has used 30145 steps.
-This is caused by the damping of
-the system. As the acceleration of the springs decreases, the timesteps rise as can
-be seen in figure \ref{fig:inverse_uniform_30fps_steps} and
-\ref{fig:inverse_multiscale_steps}. By carefully scaling the
-timestep varying heuristic, the timestep adaption gains benefit over the
-standard simulation. The downside is however, that we first have to put an
-effort into scaling the heuristic to fit our simulation, but for the standard
-simulation we likewise have to experiment with the timestep size in order to
-find the threshold at which switches start occuring.
+seconds) the GEA simulation used 455 steps while the standard simulation
+used 451 steps. If we however let the simulation run for a longer period
+of time, the GEA method surpasses the standard simulation. As we see in
+figure \ref{fig:inverse_multiscale_cumstep} the difference in accumulated
+number of timesteps rises as time passes. 20,000 seconds into the
+simulation the standard simulation has used 36001 steps and GEA has used
+30145 steps. This is caused by the dampening of the system. As the acceleration
+of the springs decreases, the timesteps rise as can be seen in figure
+\ref{fig:inverse_uniform_30fps_steps} and \ref{fig:inverse_multiscale_steps}.
+By carefully scaling the timestep varying heuristic, the timestep adaption
+gains benefit over the standard simulation. The downside is however, that we
+first have to put an effort into scaling the heuristic to fit our simulation,
+but for the standard simulation we likewise have to experiment with the
+timestep size in order to find the threshold at which switches start occuring.
 
@@
 modifying the timestep for the GEA method in the current setup and simulation
-is more expensive than simply performing the standard simulation and hence we
+is more expensive than simply performing the standard simulation. In short we
 have gained no advantage. The execution time for the GEA method could however
@@
 Furthermore the advantage of the adapting is dependent of the simulation in
-question and hence further investigation of the possible use of the GEA method
-should be performed before a final conclusion can be made.
+question. Further investigation of the possible use of the GEA method
+should therefore be performed before a final conclusion can be made.
 
-Another heuristic which potentially could be used for the adaptive timestepping
-would be the acceleration (as opposed to the inverse of the acceleration in
-equation \ref{eq:inverse}). This could be done by having a basis timestep which
-is sufficient when the springs have a low acceleration. When the acceleration
-rises the timestep is decreased in stages instead of a direct one-to-one mapping
-depending on the order of magnitude of the acceleration.
+Another heuristic which potentially could be used for the adaptive
+timestepping would be the acceleration (as opposed to the inverse of the
+acceleration in equation \ref{eq:inverse}). This could be done by having a
+basis timestep which is sufficient when the springs have a low acceleration.
+When the acceleration rises the timestep is decreased in stages instead
+of a direct one-to-one mapping depending on the order of magnitude of the
+acceleration.
 
@@
 
-\section{Future work}
+\section{Further discussion and future work}
+\label{sec:future_work}
+We here present a short discussion of the methods with the purpose of presenting
+possible future work.
 
-Combine atomic approach with inverse acceleration heuristic
+All three methods proposed in this paper could possibly benefit from being
+tested using alternative integration methods. This is due to the fact that
+the explicit Euler method is quite numerically unstable if too large steps
+are taken. Using a more stable integration method could therefore potentially
+change the outcome of the experiments of the methods.
 
-Test methods with other integration schemes
+As mentioned in section \ref{sec:experiment_switch} the Adapt-on-switch method
+could possibly be succesful on a domain modelling the behavior of springs more
+naturally and hence future work could be done in this direction.
 
-Test methods with hyperelastic materials
+The GEA method should be further investigated to see if it could
+give a computational benefit or not as concluded in section
+\ref{sec:experiments_inverse}. Likewise the alternative acceleration heuristic
+proposed at the end of section \ref{sec:experiments_inverse} could potentially
+be implemented, tested and compared to the inverse acceleration heuristic of
+the GEA method to see which of the two methods performs best.
 
-Relate to hyperelastic materials (some linear heuristic = win)
+As we saw in section \ref{sec:experiments} both the AUH method with the
+$n_{last}$ heuristic (only) and the GEA method where (partially) succesful.
+These methods and their applications should therefore be further investigated.
+Combinations of these methods and heuristics could likewise be investigated in
+order to get possibly faster LAOT schemes than the ones proposed, tested and
+discussed in this report. An example of such a LAOT scheme could be a fully
+atomic timestepping scheme where the inverse acceleration heuristic would be
+applied to the atomic updating method. This would result in a LAOT method
+which is both spatial and time adative. It would however require quite a lot
+of bookkeeping and investigation of asynchronous update behavior, which is not
+covered in this report due to the fact that the heuristic for the asynchronous
+behaviour of the AUH method was rejected prior due to other flaws.
+
+Finally the succesful schemes should be applied to hyperelastic material
+simulations, which is the original motivation for the work of this report.
 
@@
 \section{Conclusion}
+In this report we have described a 1-dimensional mass-spring system using the
+explicit Euler. Furthermore we have proposed, described, tested and discussed
+the three LAOT methods Adapt-on-switch, Atomic Updating using Heuristics, and
+Global Estimated Acceleration and the heuristics for each of the methods.
+All these methods (including the standard simulation) and heuristics were
+implemented as described in this report.
+
+The Adapt-on-switch method to predict and correct for spring switches was
+rejected for this domain after testing. The AUH heuristic of using $r_{last}$
+to perform asynchronious Atomic Updating of Accelerations was likewise
+rejected after testing.
+
+The AUH method using only the $n_{last}$ heuristic bound was succesful when
+using the explicit Euler integration, but the advantage of the method is
+likely to disappear when using other integration schemes.
+
+The GEA method was succesful in increasing the timestep size over time and
+hence decrease the number of total steps. It was however too computationally
+heavy in the current implementation and should therefore be investigated further
+before making a final conclusion.
+
+Further future work was proposed for all three methods including amongst
+others: testing all methods by using alternative integration schemes,
+experimenting with combinations of the AUH and GEA methods and heuristics,
+and applying the LAOT methods and heuristics for hyperelastic material
+simulations.
 
